### 2019-04-09 analysis/02_analysis.Rmd
Move analysis to its own directory

---
title: "Results"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    pandoc_args:
      - "--default-image-extension=png"
editor_options: 
  chunk_output_type: console
---

```{r load-libraries-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(ggraph)
library(ggdag)
library(here)

# General settings
source(here("analysis", "options.R"))

# Make all the randomness reproducible
set.seed(1234)

# Load data
# results <- readRDS(here("data", "derived_data", "results.rds"))
```

# Causal pathway

Our theory and hypotheses are laid out in the causal pathway below. Our outcomes (Y: % likely to donate and amount donated) "listen to" or respond to C (crackdown), which is our main treatment or exposure. Funding (F) and issue (I) serve as heuristics for donation and influence a government's decision to crack down on NGOs, hence the dual arrows to C and Y.

```{r causal-dag, fig.width=13/3, fig.height=2}
node_colors <- tribble(
  ~name, ~type,
  "D", "Outcome",
  "I", "Organizational factor",
  "F", "Organizational factor",
  "O", "Organizational factor",
  "G", "Structural factor"
)

# Set up DAG structure
theory_dag <- dagify(D ~ I + F + O + G,
                     G ~ I + F,
                     outcome = "D",
                     exposure = "C") %>% 
  tidy_dagitty(layout = "dh", seed = 1234)

# Add node types/colors to DAG data
theory_dag$data <- theory_dag$data %>% 
  left_join(node_colors, by = "name")

# Make DAG plot
plot_dag <- ggplot(theory_dag, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(size = 6, mapping = aes(color = type)) +
  geom_dag_edges(start_cap = circle(4, "mm"),
                 end_cap = circle(4, "mm")) +
  geom_dag_text(size = pts(6), family = "Roboto Condensed", fontface = "bold") +
  scale_dag() +
  theme_ngo() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank())

# Show and save plot
plot_dag
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.pdf"),
       width = 13/3, height = 2, units = "in", device = cairo_pdf)
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.png"),
       width = 13/3, height = 2, units = "in", type = "cairo", dpi = 300)
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=100}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>