### 2019-04-09 analysis/experiment.Rmd

---
title: "Experiment"
date: "Last run: `r format(Sys.time(), '%F')`"
---

# Survey experiment {.survey}

```{r child = "survey/experiment.md"}
```