### 2019-04-09 README.Rmd
Package-ize this puppy!

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)

Title <- "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
Authors <- "Suparna Chaudhry, Marc Dotson, and Andrew Heiss"
Year <- "2019"
```

# `r Title`

[Suparna Chaudhry](http://www.suparnachaudhry.com/) • Christopher Newport University  
[Marc Dotson](https://marriottschool.byu.edu/directory/details?id=50683) • Brigham Young University  
[Andrew Heiss](https://www.andrewheiss.com) • Brigham Young University

---

## Abstract

As space for civil society has closed around the world, transnational NGOs have faced a crisis of funding. We explore how NGOs have shifted from traditionally Northern funding sources toward grassroots private philanthropic money. How do individual donors in the US feel about donating to legally besieged NGOs abroad? Do legal restrictions on NGOs influence donors’ decision to donate? We use a conjoint survey experiment to argue that domestic political environments of NGO host countries influence preferences of private donors and that legal crackdowns on NGOs serve as a heuristic of organizational deservingness.

---

This repository contains the data and code for our paper. Our pre-print is online here:

> `r Authors`. (`r Year`). *`r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`. Online at `FUTURE_URL`

Our analysis notebook is accessible at <https://stats.andrewheiss.com/why-donors-donate/>.

## How to cite

Please cite this compendium as:

> `r Authors`. (`r format(Sys.Date(), "%Y")`). *Compendium of R code and data for `r Title`*. Accessed `r format(Sys.time(), '%B %e, %Y')`. Online at `FUTURE_URL`

## How to download and install

You can either [download the compendium as a ZIP file](/archive/master.zip) or use GitHub to clone or fork the compendium repository (see the green "Clone or download" button at the top of the GitHub page).

In order to reproduce this project, you'll need to install the compendium as an R package. After downloading the compendium, do the following:

1. Open the compendium as a project in RStudio (double click on `donorheuristics.proj`).
2. Build the compendium package by clicking on the "Install and Restart" button in RStudio's "Build" panel.

You should now be able to run `library(donorheuristics)` and reproduce all the findings in the paper (as well as the paper itself).

To reproduce the analysis, run `make build` from RStudio's "Terminal" panel. Open `analysis/_site/` to see the results. Run `make serve` to serve the site at <http://localhost:7000>.

To reproduce the paper, run `make html` or `make tex` or `make docx` or `make paper` (for all three output formats) from the terminal. Open `manuscript/` (or `manuscript/tex_out/` for PDFs) to see the results.

---

## Project layout

There are several subdirectories with specific purposes:

**Tracked with git**

- `/data`: Data goes here. ***Temporarily* not tracked with git.**
- `/analysis`: All analysis related files go here as standalone R Markdown files dedicated to specific tasks (like `01_clean-data.Rmd` and `02_analysis.Rmd`, etc.). The most current version of this analysis notebook is accessible online at <https://stats.andrewheiss.com/why-donors-donate/>.
- `/R`: Various project-specific R scripts and functions go here. These are built as part of the package and are accessible after loading `library(donorheuristics)`.
- `/manuscript`: The actual writing goes here. A separate Makefile here generates the output as HTML, PDF (through XeTeX), and Word. For now, that Makefile depends on a bunch of helper scripts that only live on Andrew Heiss's computer, so he's the only one that can build the actual paper. Eventually he'll move those scripts (into the `/manuscript/pandoc` folder) here so anyone can build it.

**Not tracked with git**

These folders are ignored with `.gitignore` and instead live in a shared Dropbox folder.

- `/analysis/_site`: This is where the generated website lives. Don't edit anything here—it gets deleted and rewritten all the time.
- `/analysis/output`: All exported tables and figures go here.
- `/admin`: For administrative tasks, like IRB, pre-registration, readings, and other general files.
- `/manuscript/submissions`: For journal and conference sub missions.

## Project workflow

This project is designed to work with [RStudio](https://www.rstudio.com/), and it is built as an [R Markdown website](https://bookdown.org/yihui/rmarkdown/rmarkdown-site.html) which gets [uploaded to the internet](https://stats.andrewheiss.com/why-donors-donate/) via a Makefile. Because of SSH keys and server credentials, only Andrew Heiss can upload the compiled site.

To build the site, click on the "Build Website" button in the Build tab in RStudio. All `.Rmd` files in the root of the project will be knit individually in isolated R sessions. `.Rmd` files not in the project root are not knit when building the site.

We follow a few style and workflow guidelines:

- Write code in R Markdown files. 
- Try to follow the [tidyverse style guide](https://style.tidyverse.org/)
- Use `here::here()` to specify file paths. This makes it easier to move work from the `sandbox` folder to an actual R Markdown file. Here are some examples of using it in scripts:

    ```r
    library(tidyverse)
    library(donorheuristics)
    library(here)

    # Load general global options
    source(here("analysis", "options.R"))

    # Save a plot
    ggsave(plot_name, filename = here("analysis", "output", "figures", "figure1.pdf"), 
           width = 6, height = 4)

    # Save a table
    some_data_frame %>% 
      pandoc.table.return(caption = "Some table {#tbl:table-id}") %>% 
      cat(file = here("analysis", "output", "tables", "tbl-some-table.md"))
    ```

## Collaboration

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.

Because not all collaborators use git, we use Dropbox as the main method for sharing code and data. Contact [Andrew Heiss](mailto:andrew@andrewheiss.com) to get access to the shared Dropbox folder.

At the same time, though, having actual version control with Git is nice. To make Git and Dropbox work correctly together, [we follow this system](http://www.math.cmu.edu/~gautam/sj/blog/20160406-dropbox-git.html). Here's how it works:

- The shared Dropbox folder contains everything for the project—the current git repository + all untracked files.
- **If you *are not* a git user**, make edits in the Dropbox folder and let Andrew know what you've done. He'll commit changes for you.
- **If you *are* a git user**, fork this repository to somewhere on your computer that's *not* in Dropbox. You'll work on your own local copy, pull any upstream changes, make edits, create branches, submit pull requests, and do all regular git things. Do not make any changes to the Dropbox folder itself, unless you're adding files to the folders that aren't tracked by git (e.g. adding a new paper to `/readings`).
- When pull requests are accepted or other commits are made, Andrew will pull those changes into the shared Dropbox folder.

tl;dr version:

- If you don't use git, edit files in the shared Dropbox folder and let Andrew know what you've done.
- If you do use git, fork this repository, edit files wherever you want on your computer (not in Dropbox, though), and make pull requests to the [main repository](https://github.com/andrewheiss/why-donors-donate). Only do stuff with the shared Dropbox folder if you're changing things that aren't tracked with git.