### 2019-05-02 analysis/preregistration.Rmd
Add generic IRB

---
title: "Pre-registration details"
date: "Last run: `r format(Sys.time(), '%F')`"
bibliography: ../manuscript/references.bib
---

# Preregistration

We preregistered this study with the [Center for Open Science's](https://cos.io/) [Open Science Framework](https://osf.io/).

<p><a href="https://osf.io/dx973/register/565fb3678c5e4a66b5582f67" class="btn btn-md btn-success disabled" target="_blank" role="button">
<i class="fa fa-pencil-square" aria-hidden="true"></i>&nbsp;
OSF preregistration protocol &raquo;
</a></p>

\

---

# Human subjects and IRB approval

This project has received IRB approval from the human subjects research committees at BYU, Georgia State, and Christopher Newport. We have included a generic version of our protocol below.

<p><a href="files/byu-irb-approval.pdf" class="btn btn-md btn-success disabled" target="_blank" role="button">
<i class="fa fa-check-circle" aria-hidden="true"></i>&nbsp;
BYU IRB approval
</a> 
&nbsp;
<a href="files/cnu-irb-approval.pdf" class="btn btn-md btn-success disabled" target="_blank">
<i class="fa fa-check-circle" aria-hidden="true" role="button"></i>&nbsp;
Christopher Newport IRB approval
</a>
&nbsp;
<a href="files/gsu-irb-approval.pdf" class="btn btn-info btn-md disabled" target="_blank" role="button">
<i class="fa fa-check-circle" aria-hidden="true"></i>&nbsp;
Georgia State IRB approval
</a></p>

\

```{r child = "survey/irb.md"}
```

# References