### 2019-05-03 analysis/survey/preregistration.Rmd
Add reserach questions

---
title: "Results"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
---

See Chaudhry and Heiss (2019) preregistration for reference: <https://osf.io/dx973>

---

# Study Information

## Title

Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

## Authorship

- Suparna Chaudhry
- Marc Dotson
- Andrew Heiss

## Research Questions

> Please list each research question included in this study.

We use a conjoint survey experiment to examine the impact of organizational and structural features of nongovernmental organizations (NGOs) on their donors’ decisions to engage in philanthropy. We explore three research questions in this study:

1. **Do donors rely on structural characteristics of NGOs as heuristics when deciding to donate? How do structural heuristics compare to organizational heuristics?**

	Donors rely on shortcuts, signals, and heuristics to determine the trustworthiness of NGOs, since seeking out complete information about an organization's deservingness and efficiency is costly and time-consuming. Previous research has found that an NGO's organizational characteristics commonly serve as heuristics for donors. Donors use an organization's overhead costs, the issues it works on, its transparency and accountability practices, and a host of other organizational practices as signals of an organization's efficiency and deservingness, which then influences their decision to make a donation. These kinds of heuristics are attributes that organizations can typically control—NGOs can publish annual reports, restructure their management, and engage in other strategies to appear more worthy of donation.

	Structural characteristics, such as the political and legal environment an NGO faces in its host country, may also serve as signals to donors of NGO deservingness. We are interested in whether the contentiousness of an NGO’s relationship with its host government influences donor decision making. Do donors care if nonprofits they care about are criticized by, persecuted by, or expelled from the countries they work in?

	We are also interested in the effect of organizational characteristics on donor decision making. How do managerial practices (financial transparency and accountability systems), funding sources (private donations and government grants), and issue areas (emergency response, environmental issues, human rights, and refugee relief) compare to structural characteristics when deciding to donate? Which heuristics are more influential?

2. **How do individual-level donor characteristics interact with structural and organizational heuristics? Which kinds of people are more or less likely to consider an NGO's host country political environment, managerial practices, funding sources, or issue area?**

	The decision to donate to an NGO is not determined solely by an organization's characteristics. Donors themselves have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. We are interested in how individual donor characteristics, such as political ideology, political knowledge, religious attendance, involvement in charitable activities, involvement in activism, and demographic attributes interact with organizational- and structural-level factors.

3. **What is the optimal mix of attribute levels for NGOs interested in maximizing donations?**

	Finally, given individual donor characteristics and preferences, we are interested in finding the optimal mix of organizational and structural attributes. What might an NGO try to emphasize in its marketing campaigns? Should it highlight its funding sources, managerial practices, issue area, or relationship with its host governments (even if that relationship is negative)?

## Hypotheses

> For each of the research questions listed in the previous section, provide one or multiple specific and testable hypotheses. Please state if the hypotheses are directional or non-directional. If directional, state the direction. A predicted effect is also appropriate here.

ANDREW TODO: write these

---

# Sampling Plan

## Existing data

Registration prior to creation of data

## Explanation of existing data

We will not use any existing data.

## Data collection procedures

> Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that don’t include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.

Participants of the survey experiment will be recruited through Centiment, a commercial online provider of high quality nonprobability opt-in survey panels. Centiment ensures panel quality by actively recruiting representative samples of the US population and provides monetary incentives and rewards to participants.

To see how varying NGO characteristics influence the decision to donate, our sample will be representative of a population of people who are likely to donate to charity. We ask potential participants a screening question early in the survey (“Q2.5: How often do you donate to charity”). If a participant responds that they give to once every few years or never, they will be disqualified from the study and the survey will end early.

We will provide Centiment with a link to the survey, which is hosted by Qualtrics. Centiment will then distribute the link to their panel. Participants are compensated through Centiment’s internal reward system through cash, points, and other incentives. Centiment does not provide precise details of participant compensation. Centiment states that their compensation is “fair,” and the company’s business model encourages the company to find and maintain high quality panelists. We thus infer that the amount provided is fair and justified. Centiment users receive compensation from the company following the completion of the survey.

## Sample size

> Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis?

Our target sample size is 1,000 participants.

## Sample size rationale

> This could include a power analysis or an arbitrary constraint such as time, money, or personnel.

A sample size of at least 500 respondents is typical for estimating a hierarchical Bayesian model based on conjoint data. We double this amount because we are interested in analyzing subpopulations of respondents, which requires a larger sample, and we had sufficient budget to acquire up to 1,000 respondents.

## Stopping rule

> If your data collection procedures do not give you full control over your exact sample size, specify how you will decide when to terminate your data collection.

Centiment will monitor how many surveys are successfully completed and will solicit responses until our 1,000 target is met.

---

# Variables

## Manipulated variables

> Describe all variables you plan to manipulate and the levels or treatment arms of each variable. For observational studies and meta-analyses, simply state that this is not applicable.

We use a partial fractional factorial design with the following attributes:

**Organizations:**

- *Amnesty International*: A London-based non-governmental organization known for its focus on human rights.
- *Greenpeace*: An independent, nonprofit, global campaigning organization known for using non-violent, creative confrontation to expose global environmental problems and their causes.
- *Oxfam*: An international group known for providing help to poor countries and disaster areas, with a focus on helping create lasting solutions to the injustice of poverty. 
- *Red Cross*: An international organization known for caring for the wounded, sick, and homeless in wartime and following natural disasters.

**Issue areas:**

- *Emergency response*: Respond to situations that pose an immediate risk to health, life, property, or the environment.
- *Environment*: Protecting the natural world and the impact of human activity on its condition.
- *Human rights*: Protect the inalienable fundamental rights to which a person is inherently entitled simply because he or she is a human being.
- *Refugee relief*: Provide relief for those who have been forced to flee his or her country because of persecution, war or violence.

**Organizational practices:**

- *Financial transparency*: Timely and reliable disclosure regarding donations and financial allocations. 
- *Accountability*: An obligation or willingness to accept responsibility or to account for one's actions.

**Funding sources:**

- *Funded primarily by many small private donations*: Funds by individuals who make small independent contributions.
- *Funded primarily by a handful of wealthy private donors*: Funds by wealthy individuals or families, who receive tax deductions for donations.
- *Funded primarily by government grants*: Non-repayable funds gifted by a government department.

**Relationship with host government:**

- *Friendly relationship with government*: Organization has a friendly relationship with its host government.
- *Criticized by government*: Organization faces public condemnation from the government.
- *Under government crackdown*: Host government has undertaken official action to limit or stop the organization.

Participants are then presented with random combinations of these attributes and asked to select which of three hypothetical organizations they'd be willing to donate to.

This results in five manipulated variables:

- Organization
- Issue area
- Organizational practices
- Funding source
- Relationship with host government

## Measured variables

> Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.

We collect one outcome measure, which we use as a dependent variable in our analysis. After displaying a table of randomized organizational attributes, we ask respondents "Imagine you are selecting an organization you will donate to and that each of the listed organizations exists. Which of the following organizations would you donate to?". Respondents are shown 12 of these sets, and their responses constitute our outcome.

We measure several other covariates that inform and predict an individual's choice of donating to a hypothetical organization (see survey instrument for complete question and response wording):

**Public affairs knowledge**

- *Q2.1*: How often do you follow national news?
- *Q2.2*: How often do you follow international news?
- *Q2.3*: Which mediums do you use to follow news?
- *Q2.4*: How often would you say you follow what's going on in government and public affairs?
- *Q5.7*: Have you ever traveled to a developing country?

**Public affairs trust**

- *Q5.6*: Here is a 7-point scale on which you trust political institutions and the state. The scale is arranged from no trust (left) to complete trust (right). Where would you place yourself on this scale?

**Public affairs activity**

- *Q5.1*: Did you vote in the last election?

**Political ideology**

- *Q5.2*: 7-point scale on which the political views that people might hold are arranged from extremely liberal (left) to extremely conservative (right).

**Social ideology**

- *Q5.11*: Here is a 7-point scale of your prosocial views. The scale is arranged from equality (left) to altruism (right). Where would you place yourself on this scale?

**Charity trust**

- *Q2.7*: How important is it that you trust a charity before giving to it?
- *Q2.8*: How much do you trust charities?

**Charity activity**

- *Q2.5*: How often do you donate to charity?
- *Q2.6*: How much did you donate to charity last year?

**Volunteer activity**

- *Q2.9*: Have you volunteered in the past 12 months?
- *Q2.10*: How often do you volunteer?

**Activism activity**

- *Q5.4*: Historically, how involved have you been in activist causes?
- *Q5.5*: Historically, how involved has your family been in activist causes?

**Voluntary association membership**

- *Q5.3*: Here is a list of different types of voluntary organizations. For each organization, indicate whether you are an active member, an inactive member, or not a member of that type of organization:

    - Church or religious organization
    - Sport or recreational organization
    - Art, music, or educational organization
    - Labor union
    - Political party
    - Environmental organization
    - Professional association
    - Humanitarian or charitable organization
    - Consumer organization
    - Other organization

**Religiosity**

- *Q5.8*: How often do you attend religious or worship services, not including weddings and funerals?
- *Q5.9*: How important is religion in your life?
- *Q5.10*: What is your current religion, if any?

**Demographics**

- *Q5.12*: What is your gender?
- *Q5.13*: Are you now married, widowed, divorced, separated, or never married?
- *Q5.14*: What is the highest degree or level of school you have completed?
- *Q5.15*: What is your annual household income before taxes?
- *Q5.16*: Choose one or more races that you consider yourself to be:
- *Q5.17*: How old are you?
- *Q5.18*: What is your ZIP code?

## Indices

> If any measurements are  going to be combined into an index (or even a mean), what measures will you use and how will they be combined? Include either a formula or a precise description of your method. If your are using a more complicated statistical method to combine measures (e.g. a factor analysis), you can note that here but describe the exact method in the analysis plan section.

TODO: The paper at <https://osf.io/u43yg> uses Bayesian hierarchical stuff and they talk a lot about index stuff. I'm not familiar with the model, though, so I don't know what applies here.

COMMENT: I don't see a need to index (i.e., use dimension reduction). We should just be collapsing covariates via Transformations in a way motivated by your experience and theory with respect to the "competing sets of covariates in the distribution of heterogeneity" that we'd like to use to compare effects on the marginal posteriors of the preference parameters.

---

# Design plan

## Study type

Experiment - A researcher randomly assigns treatments to study subjects, this includes field or lab experiments. This is also known as an intervention experiment and includes randomized controlled trials.

## Blinding

For studies that involve human subjects, they will not know the treatment group to which they have been assigned.

## Study design

> Describe your study design. Examples include two-group, factorial, randomized block, and repeated measures. Is it a between (unpaired), within-subject (paired), or mixed design? Describe any counterbalancing required. Typical study designs for observation studies include cohort, cross sectional, and case-control studies.

We use a fractional factorial design. Since no single respondent can possibly see all possible combinations of the attribute levels, we create a number of different versions of the experimental design. We utilize a hierarchical Bayesian model in part to allow for information sharing across like respondents when estimating individual-level preferences for the attribute levels.

## Randomization

> If you are doing a randomized study, how will you randomize, and at what level?

Every respondent will be randomly assigned a version of the fractional factorial experimental design.

---

# Analysis Plan

## Statistical models

> What statistical model will you use to test each hypothesis? Please include the type of model (e.g. ANOVA, multiple regression, SEM, etc) and the specification of the model (this includes each variable that will be included as predictors, outcomes, or covariates). Please specify any interactions that will be tested and remember that any test not included here must be noted as an exploratory test in your final article.

We will use a hierarchical Bayesian multinomial logit model with conjugate or otherwise typical priors. The individual-level model is the multinomial logit and the upper-level model of heterogeneity is multivariate normal,

$$
\begin{aligned}
\beta &\sim \text{MVN}(Z \Gamma, \xi) \\
y &\sim \text{MNL}(X \beta, \varepsilon)
\end{aligned}
$$

where $y$ = which alternative the respondent chooses to donate, $X$ = design matrix of attribute levels, $\beta$ = latent individual preferences for the attribute levels, $Z$ = matrix of individual-level covariates, $\Gamma$ = matrix of coefficients mapping individual-level covariates onto the latent individual-level preferences, and $\varepsilon$ and $\xi$ = errors.

## Transformations

> If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process.

TODO: Are there any others we should collapse?

COMMENT: See comment for Indices.

We collapse several of our variables into dichotomous and trichotomous versions, following this system:

*Q2.1*: How often do you follow national news?

- Multiple times a day → At least once a day
- Every day → At least once a day
- Once a week → Once a week
- Hardly ever → Rarely (base case)
- Never → Rarely

*Q2.4*: How often would you say you follow what's going on in government and public affairs?

- Most of the time → Often
- Some of the time → Often
- Only now and then → Not often (base case)
- Hardly at all → Not often (base case)

*Q2.5*: How often do you donate to charity (with either cash or in-kind)?

- Once a week → At least once a month
- Once a month → At least once a month
- Once every three months → More than once a month, less than once a year
- Once every six months → More than once a month, less than once a year
- Once a year → More than once a month, less than once a year
- Once every few years → Rarely (base case)
- Never → Rarely

*Q2.10*: How often do you volunteer?

- Once a week → At least once a month
- Once a month → At least once a month
- Once every three months → More than once a month, less than once a year
- Once every six months → More than once a month, less than once a year
- Once a year → More than once a month, less than once a year
- Once every few years → Rarely (base case)
- Never → Rarely

*Q5.4*: Historically, how involved have you been in activist causes?

- Extremely involved → Involved
- Very involved → Involved
- Moderately involved → Involved
- Slightly involved → Not involved (base case)
- Never involved → Not involved

*Q5.5*: Historically, how involved has your family been in activist causes?

- Extremely involved → Involved
- Very involved → Involved
- Moderately involved → Involved
- Slightly involved → Not involved (base case)
- Never involved → Not involved

*Q5.8*: How often do you attend religious or worship services, not including weddings and funerals?

- More than once a week → At least once a month
- Once a week → At least once a month
- Once or twice a month → At least once a month
- A few times a year → Rarely (base case)
- Seldom → Rarely
- Never → Rarely
- Don't know → NULL

*Q5.9*: How important is religion in your life?

- Extremely important → Important
- Very important → Important
- Moderately important → Important
- Slightly important → Not important (base case)
- Not at all important → Not important

*Q5.15*: What is your annual household income before taxes?

- Less than $10,000 → More/less than median (base case = less than median)
- \$10,000 to \$19,999 → More/less than median
- \$20,000 to \$29,999 → More/less than median
- \$30,000 to \$39,999 → More/less than median
- \$40,000 to \$49,999 → More/less than median
- \$50,000 to \$59,999 → More/less than median
- \$60,000 to \$69,999 → More/less than median
- \$70,000 to \$79,999 → More/less than median
- \$80,000 to \$89,999 → More/less than median
- \$90,000 to \$99,999 → More/less than median
- \$100,000 to \$149,999 → More/less than median
- \$150,000 to \$199,999 → More/less than median
- \$200,000 to \$299,999 → More/less than median
- \$300,000 or more → More/less than median

*Q5.17*: How old are you?

- Under 18 → More/less than median (base case = less than median)
- 18 - 24 → More/less than median
- 25 - 34 → More/less than median
- 35 - 44 → More/less than median
- 45 - 54 → More/less than median
- 55 - 64 → More/less than median
- 65 - 74 → More/less than median
- 75 - 84 → More/less than median
- 85 or older → More/less than median

## Follow-up analyses

> If not specified previously, will you be conducting any confirmatory analyses to follow up on effects in your statistical model, such as subgroup analyses, pairwise or complex contrasts, or follow-up tests from interactions? Remember that any analyses not specified in this research plan must be noted as exploratory.

TODO: Do we do this?

COMMENT: No, it's really just looking at/comparing the posterior marginals on their own vs. the posterior marginals with different covariates includes, plus the market simulation.

## Inference criteria

> What criteria will you use to make inferences? Please describe the information you’ll use (e.g. specify the p-values, Bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?

NOTE: Here's a preregistration from another Bayesian hierarchical multinomial logit for reference: <https://osf.io/u43yg>

COMMENT: Specify what the "competing sets of covariates" are that we'll include in the distribution of heterogeneity. This should be the motivation of the collapses specified in Transformations.

We will examine the aggregate marginal posterior distributions of the attribute levels and utilize 95% credible intervals to establish "significance." Effects are "significant" if the 95% credible intervals don't include 0. Similarly, marginal posterior distributions are "significantly" different if the 95% credible intervals don't overlap.

We will examine the marginal posterior distributions of the following models:

- Organizational and structural attribute levels with an intercept-only distribution of heterogeneity
- Organizational and structural attribute levels with competing sets of covariates in the distribution of heterogeneity

Finally, we will employ the posterior distribution of model parameters to conduct counterfactual analyses via a market simulator to determine optimal policies.

## Data exclusion

> How will you determine which data points or samples (if any) to exclude from your analyses? How will outliers be handled?

We ask potential participants a screening question early in the survey ("Q2.5: How often do you donate to charity"). If a participant responds that they give to once every few years or never, they will be disqualified from the study and the survey will end early.

We include one question ("*Q2.11*: Please select blue from the following list:") to monitor respondent attention. In our analysis we will exclude respondents who fail this question.

## Missing data

> How will you deal with incomplete or missing data?

Because all survey questions are required, we do not anticipate issues with incomplete or missing data.

## Exploratory analysis (optional)

> If you plan to explore your data set to look for unexpected differences or relationships, you may describe those tests here. An exploratory test is any test where a prediction is not made up front, or there are multiple possible tests that you are going to use. A statistically significant finding in an exploratory test is a great way to form a new confirmatory hypothesis, which could be registered at a later time.

NA.

---

# Script (optional)

> Upload an analysis script with clear comments. This optional step is helpful in order to create a process that is completely transparent and increase the likelihood that your analysis can be replicated. We recommend that you run the code on a simulated dataset in order to check that it will run without errors.

Nothing.

---

# Other

> If there is any additional information that you feel needs to be included in your preregistration, please enter it here.

Nothing.