### 2019-05-13 analysis/survey/preregistration.Rmd
Flesh out hypotheses

---
title: "Preregistration"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
---

# Study Information

## Title

Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy

## Authorship

- Suparna Chaudhry
- Marc Dotson
- Andrew Heiss

## Research Questions

> Please list each research question included in this study.

We use a conjoint survey experiment to examine the impact of organizational features of nongovernmental organizations (NGOs) and the structural factors in target countries in which they operate on donors’ decisions to engage in philanthropy. We explore three research questions in this study:

1. **Do donors rely on structural characteristics of NGOs as heuristics when deciding to donate? How do structural heuristics compare to organizational heuristics?**

    Donors rely on shortcuts, signals, and heuristics to determine the trustworthiness of NGOs, since seeking out complete information about an organization's deservingness and efficiency is costly and time-consuming. Previous research has found that an NGO's organizational characteristics commonly serve as heuristics for donors. Donors use an organization's overhead costs, the issues it works on, its transparency and accountability practices, and a host of other organizational practices as signals of an organization's efficiency and deservingness, which then influences their decision to make a donation. These kinds of heuristics are attributes that organizations can typically control—NGOs can publish annual reports, restructure their management, and engage in other strategies to appear more worthy of donation.

    Structural characteristics, such as the political and legal environment an NGO faces in its host country, may also serve as signals to donors of NGO deservingness. We are interested in whether the contentiousness of an NGO’s relationship with its host government influences donor decision making. Do donors care if nonprofits they care about are criticized by, persecuted by, or expelled from the countries they work in?

    We are also interested in the effect of organizational characteristics on donor decision making. How do managerial practices (financial transparency and accountability systems), funding sources (private donations and government grants), and issue areas (emergency response, environmental issues, human rights, and refugee relief) compare to structural characteristics when deciding to donate? Which heuristics are more influential?

2. **How do individual-level donor characteristics interact with structural and organizational heuristics? Which kinds of people are more or less likely to consider an NGO's host country political environment, managerial practices, funding sources, or issue area?**

    The decision to donate to an NGO is not determined solely by an organization's characteristics. Donors themselves have personality traits, preferences, and experiences that make them more or less likely to engage in philanthropy. We are interested in how individual donor characteristics, such as political ideology, political knowledge, religious attendance, involvement in charitable activities, involvement in activism, and demographic attributes interact with organizational- and structural-level factors.

3. **What is the optimal mix of attribute levels for NGOs interested in maximizing donations?**

    Finally, given individual donor characteristics and preferences, we are interested in finding the optimal mix of organizational and structural attributes. What might an NGO try to emphasize in its marketing campaigns? Should it highlight its funding sources, managerial practices, issue area, or relationship with its host governments (even if that relationship is negative)?

## Hypotheses

> For each of the research questions listed in the previous section, provide one or multiple specific and testable hypotheses. Please state if the hypotheses are directional or non-directional. If directional, state the direction. A predicted effect is also appropriate here.

For our first set of questions, we predict that:

1. **Branding**

    - Donors will be more likely to donate to Oxfam and Red Cross compared to Amnesty International and Greenpeace [Mechanism: awareness of need and contentiousness of issue area]

2. **Government crackdown**

    - Donors will show increased willingness to donate to NGOs that are facing government crackdown or criticism [Mechanism: Governments wouldn't be cracking down on them if they didn't perceive a threat from them which means organizations implementing their missions effectively. This perception of efficacy leads to increased donations.]
    - Donors will show increased willingness to donate to Oxfam and Red Cross when they are facing government crackdown or criticism compared to when Amnesty or Greenpeace is facing crackdown.

3. **Issue area**

    - Donors will show increased willingness to donate to NGOs working in less contentious issue areas (emergency response and refugee relief) over more contentious issue areas (environment and human rights) 
    - Donors will show increased willingness to donate to NGOs facing government crackdown/criticism working in less contentious issue areas (emergency response and refugee relief) over more contentious issue areas (environment and human rights) [Mechanisms: Perceptions of deservingness of NGOs dealing with emergency response and refugee relief. Donors are also more likely to donate to programs that are compatible with government preferences and have easily measurable outputs, which environment and human rights programs often lack. NGOs working on more contentious issue areas may be expelled or shut down, which would be a waste of donor resources, make it less likely that they donate to these groups.]

4. **Funding sources**

    - Donors will show increased willingness to donate to NGOs that are funded primarily by numerous small private donors compared to NGOs that are funded by a handful of wealhty private donors and government grants [Mechanism: Perception of efficacy - your contribution matters as a small donor. Government funding may also imply lack of independence of government which can reduce the efficiency of an organization.]
    - Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors
    - Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors and work in less contentious areas (emergency response and refugee relief)

5. **Organizational practices**

    - Donors will show increased willingness to donate to NGOs that are financially transparent [Mechanism: Perception of efficacy]
    - Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent
    - Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and are funded primarily by numerous small private donors
    - Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief) 
    - Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief) and are funded by numerous small donors
    - Donors should be no more or less likely to donate to NGOs that are accountable and hold regular third party audits [Mechanism: Donors don't necessarily seek assurance through third-party programs/audits and charity watchdogs, but rather through word of mouth, personal scrutiny and local networks]

Because of the nature of our statistical methods, we do not have exact hypotheses for the second and third set of questions. We describe how we answer these questions in the "Follow-up analyses" and "Exploratory analysis" sections below.

# Sampling Plan

## Existing data

Registration prior to creation of data

## Explanation of existing data

We will not use any existing data.

## Data collection procedures

> Please describe the process by which you will collect your data. If you are using human subjects, this should include the population from which you obtain subjects, recruitment efforts, payment for participation, how subjects will be selected for eligibility from the initial pool (e.g. inclusion and exclusion rules), and your study timeline. For studies that don’t include human subjects, include information about how you will collect samples, duration of data gathering efforts, source or location of samples, or batch numbers you will use.

Participants of the survey experiment will be recruited through Centiment, a commercial online provider of high quality nonprobability opt-in survey panels. Centiment ensures panel quality by actively recruiting representative samples of the US population and provides monetary incentives and rewards to participants.

To see how varying NGO characteristics influence the decision to donate, our sample will be representative of a population of people who are likely to donate to charity. We ask potential participants a screening question early in the survey (“Q2.5: How often do you donate to charity”). If a participant responds that they give to once every few years or never, they will be disqualified from the study and the survey will end early.

We will provide Centiment with a link to the survey, which is hosted by Qualtrics. Centiment will then distribute the link to their panel. Participants are compensated through Centiment’s internal reward system through cash, points, and other incentives. Centiment does not provide precise details of participant compensation. Centiment states that their compensation is “fair,” and the company’s business model encourages the company to find and maintain high quality panelists. We thus infer that the amount provided is fair and justified. Centiment users receive compensation from the company following the completion of the survey.

## Sample size

> Describe the sample size of your study. How many units will be analyzed in the study? This could be the number of people, birds, classrooms, plots, interactions, or countries included. If the units are not individuals, then describe the size requirements for each unit. If you are using a clustered or multilevel design, how many units are you collecting at each level of the analysis?

Our target sample size is 1,000 participants.

## Sample size rationale

> This could include a power analysis or an arbitrary constraint such as time, money, or personnel.

A sample size of at least 500 respondents is typical for estimating a hierarchical Bayesian model based on conjoint data. We double this amount because we are interested in analyzing subpopulations of respondents, which requires a larger sample, and we had sufficient budget to acquire up to 1,000 respondents.

## Stopping rule

> If your data collection procedures do not give you full control over your exact sample size, specify how you will decide when to terminate your data collection.

Centiment will monitor how many surveys are successfully completed and will solicit responses until our 1,000 target is met.

---

# Variables

## Manipulated variables

> Describe all variables you plan to manipulate and the levels or treatment arms of each variable. For observational studies and meta-analyses, simply state that this is not applicable.

We use a partial fractional factorial design with the following attributes:

**Organizations:**

- *Amnesty International*: A London-based non-governmental organization known for its focus on human rights.
- *Greenpeace*: An independent, nonprofit, global campaigning organization known for using non-violent, creative confrontation to expose global environmental problems and their causes.
- *Oxfam*: An international group known for providing help to poor countries and disaster areas, with a focus on helping create lasting solutions to the injustice of poverty. 
- *Red Cross*: An international organization known for caring for the wounded, sick, and homeless in wartime and following natural disasters.

**Issue areas:**

- *Emergency response*: Respond to situations that pose an immediate risk to health, life, property, or the environment.
- *Environment*: Protecting the natural world and the impact of human activity on its condition.
- *Human rights*: Protect the inalienable fundamental rights to which a person is inherently entitled simply because he or she is a human being.
- *Refugee relief*: Provide relief for those who have been forced to flee his or her country because of persecution, war or violence.

**Organizational practices:**

- *Financial transparency*: Organization discloses information regarding its donations and financial allocations in a timely and reliable manner.
- *Accountability*: Organization undergoes a regular third-party audit to ensure that it is meeting its program goals and obligations.

**Funding sources:**

- *Funded primarily by many small private donations*: Funds by individuals who make small independent contributions.
- *Funded primarily by a handful of wealthy private donors*: Funds by wealthy individuals or families, who receive tax deductions for donations.
- *Funded primarily by government grants*: Non-repayable funds gifted by a government department.

**Relationship with host government:**

- *Friendly relationship with government*: Organization has a friendly relationship with its host government.
- *Criticized by government*: Organization faces public condemnation from the government.
- *Under government crackdown*: Host government has undertaken official action to limit or stop the organization.

Participants are then presented with random combinations of these attributes and asked to select which of three hypothetical organizations they'd be willing to donate to.

This results in five manipulated variables:

- Organization
- Issue area
- Organizational practices
- Funding source
- Relationship with host government

## Measured variables

> Describe each variable that you will measure. This will include outcome measures, as well as any predictors or covariates that you will measure. You do not need to include any variables that you plan on collecting if they are not going to be included in the confirmatory analyses of this study.

We collect one outcome measure, which we use as a dependent variable in our analysis. After displaying a table of randomized organizational attributes, we ask respondents "Imagine you are selecting an organization you will donate to and that each of the listed organizations exists. Which of the following organizations would you donate to?". Respondents are shown 12 of these sets, and their responses constitute our outcome.

We measure several other covariates that inform and predict an individual's choice of donating to a hypothetical organization (see survey instrument for complete question and response wording):

**Public affairs knowledge**

- *Q2.1*: How often do you follow national news?
- *Q2.2*: How often do you follow international news?
- *Q2.3*: Which mediums do you use to follow news?
- *Q2.4*: How often would you say you follow what's going on in government and public affairs?
- *Q5.7*: Have you ever traveled to a developing country?

**Public affairs trust**

- *Q5.6*: Here is a 7-point scale on which you trust political institutions and the state. The scale is arranged from no trust (left) to complete trust (right). Where would you place yourself on this scale?

**Public affairs activity**

- *Q5.1*: Did you vote in the last election?

**Political ideology**

- *Q5.2*: 7-point scale on which the political views that people might hold are arranged from extremely liberal (left) to extremely conservative (right).

**Social ideology**

- *Q5.11*: Here is a 7-point scale of your prosocial views. The scale is arranged from equality (left) to altruism (right). Where would you place yourself on this scale?

**Charity trust**

- *Q2.7*: How important is it that you trust a charity before giving to it?
- *Q2.8*: How much do you trust charities?

**Charity activity**

- *Q2.5*: How often do you donate to charity?
- *Q2.6*: How much did you donate to charity last year?

**Volunteer activity**

- *Q2.9*: Have you volunteered in the past 12 months?
- *Q2.10*: How often do you volunteer?

**Activism activity**

- *Q5.4*: Historically, how involved have you been in activist causes?
- *Q5.5*: Historically, how involved has your family been in activist causes?

**Voluntary association membership**

- *Q5.3*: Here is a list of different types of voluntary organizations. For each organization, indicate whether you are an active member, an inactive member, or not a member of that type of organization:

    - Church or religious organization
    - Sport or recreational organization
    - Art, music, or educational organization
    - Labor union
    - Political party
    - Environmental organization
    - Professional association
    - Humanitarian or charitable organization
    - Consumer organization
    - Other organization

**Religiosity**

- *Q5.8*: How often do you attend religious or worship services, not including weddings and funerals?
- *Q5.9*: How important is religion in your life?
- *Q5.10*: What is your current religion, if any?

**Demographics**

- *Q5.12*: What is your gender?
- *Q5.13*: Are you now married, widowed, divorced, separated, or never married?
- *Q5.14*: What is the highest degree or level of school you have completed?
- *Q5.15*: What is your annual household income before taxes?
- *Q5.16*: Choose one or more races that you consider yourself to be:
- *Q5.17*: How old are you?
- *Q5.18*: What is your ZIP code?

## Indices

> If any measurements are  going to be combined into an index (or even a mean), what measures will you use and how will they be combined? Include either a formula or a precise description of your method. If your are using a more complicated statistical method to combine measures (e.g. a factor analysis), you can note that here but describe the exact method in the analysis plan section.

We will not combine any measures to create indicies.

---

# Design plan

## Study type

Experiment - A researcher randomly assigns treatments to study subjects, this includes field or lab experiments. This is also known as an intervention experiment and includes randomized controlled trials.

## Blinding

For studies that involve human subjects, they will not know the treatment group to which they have been assigned.

## Study design

> Describe your study design. Examples include two-group, factorial, randomized block, and repeated measures. Is it a between (unpaired), within-subject (paired), or mixed design? Describe any counterbalancing required. Typical study designs for observation studies include cohort, cross sectional, and case-control studies.

We use a fractional factorial design. Since no single respondent can possibly see all possible combinations of the attribute levels, we create a number of different versions of the experimental design. We utilize a hierarchical Bayesian model in part to allow for information sharing across like respondents when estimating individual-level preferences for the attribute levels.

## Randomization

> If you are doing a randomized study, how will you randomize, and at what level?

Every respondent will be randomly assigned a version of the fractional factorial experimental design.

---

# Analysis Plan

## Statistical models

> What statistical model will you use to test each hypothesis? Please include the type of model (e.g. ANOVA, multiple regression, SEM, etc) and the specification of the model (this includes each variable that will be included as predictors, outcomes, or covariates). Please specify any interactions that will be tested and remember that any test not included here must be noted as an exploratory test in your final article.

We will use a hierarchical Bayesian multinomial logit model with conjugate or otherwise typical priors. The individual-level model is the multinomial logit and the upper-level model of heterogeneity is multivariate normal,

$$
\begin{aligned}
\beta &\sim \operatorname{Multivariate} \mathcal{N}(Z \Gamma, \xi) \\
y &\sim \operatorname{Multinomial logit}(X \beta, \varepsilon)
\end{aligned}
$$

where $y$ = which alternative the respondent chooses to donate, $X$ = design matrix of attribute levels, $\beta$ = latent individual preferences for the attribute levels, $Z$ = matrix of individual-level covariates, $\Gamma$ = matrix of coefficients mapping individual-level covariates onto the latent individual-level preferences, and $\varepsilon$ and $\xi$ = errors.

## Transformations

> If you plan on transforming, centering, recoding the data, or will require a coding scheme for categorical variables, please describe that process.

We collapse several of our variables into dichotomous and trichotomous versions, following this system:

*Q2.1*: How often do you follow national news?

- Multiple times a day → At least once a day
- Every day → At least once a day
- Once a week → Once a week
- Hardly ever → Rarely (base case)
- Never → Rarely

*Q2.4*: How often would you say you follow what's going on in government and public affairs?

- Most of the time → Often
- Some of the time → Often
- Only now and then → Not often (base case)
- Hardly at all → Not often (base case)

*Q2.5*: How often do you donate to charity (with either cash or in-kind)?

- Once a week → At least once a month
- Once a month → At least once a month
- Once every three months → More than once a month, less than once a year
- Once every six months → More than once a month, less than once a year
- Once a year → More than once a month, less than once a year
- Once every few years → Rarely (base case)
- Never → Rarely

*Q2.10*: How often do you volunteer?

- Once a week → At least once a month
- Once a month → At least once a month
- Once every three months → More than once a month, less than once a year
- Once every six months → More than once a month, less than once a year
- Once a year → More than once a month, less than once a year
- Once every few years → Rarely (base case)
- Never → Rarely

*Q5.4*: Historically, how involved have you been in activist causes?

- Extremely involved → Involved
- Very involved → Involved
- Moderately involved → Involved
- Slightly involved → Not involved (base case)
- Never involved → Not involved

*Q5.5*: Historically, how involved has your family been in activist causes?

- Extremely involved → Involved
- Very involved → Involved
- Moderately involved → Involved
- Slightly involved → Not involved (base case)
- Never involved → Not involved

*Q5.8*: How often do you attend religious or worship services, not including weddings and funerals?

- More than once a week → At least once a month
- Once a week → At least once a month
- Once or twice a month → At least once a month
- A few times a year → Rarely (base case)
- Seldom → Rarely
- Never → Rarely
- Don't know → NULL

*Q5.9*: How important is religion in your life?

- Extremely important → Important
- Very important → Important
- Moderately important → Important
- Slightly important → Not important (base case)
- Not at all important → Not important

*Q5.15*: What is your annual household income before taxes?

- Less than $10,000 → More/less than median (base case = less than median)
- \$10,000 to \$19,999 → More/less than median
- \$20,000 to \$29,999 → More/less than median
- \$30,000 to \$39,999 → More/less than median
- \$40,000 to \$49,999 → More/less than median
- \$50,000 to \$59,999 → More/less than median
- \$60,000 to \$69,999 → More/less than median
- \$70,000 to \$79,999 → More/less than median
- \$80,000 to \$89,999 → More/less than median
- \$90,000 to \$99,999 → More/less than median
- \$100,000 to \$149,999 → More/less than median
- \$150,000 to \$199,999 → More/less than median
- \$200,000 to \$299,999 → More/less than median
- \$300,000 or more → More/less than median

*Q5.17*: How old are you?

- Under 18 → More/less than median (base case = less than median)
- 18 - 24 → More/less than median
- 25 - 34 → More/less than median
- 35 - 44 → More/less than median
- 45 - 54 → More/less than median
- 55 - 64 → More/less than median
- 65 - 74 → More/less than median
- 75 - 84 → More/less than median
- 85 or older → More/less than median

## Follow-up analyses

> If not specified previously, will you be conducting any confirmatory analyses to follow up on effects in your statistical model, such as subgroup analyses, pairwise or complex contrasts, or follow-up tests from interactions? Remember that any analyses not specified in this research plan must be noted as exploratory.

In our second set of research questions, we explore how individual-level donor characteristics interact with structural and organizational heuristics. We are interested in finding which kinds of people are more or less likely to consider an NGO's host country political environment, managerial practices, funding sources, or issue area when deciding to donate.

There is substantial literature regarding individual factors that motivate people to donate to charities, and a full test of each of these factors goes beyond the scope of this research. In general, those who trust charities, are more involved in philanthropy, advocacy, and other voluntary associations, and are more religious tend to be more likely to donate to charity. In this study, however, we are less interested in the effects of single characteristics and more interested in how constellations of characteristics interact with specific heuristics. For instance, we suspect that politically conservative, religiously active donors who have a history of volunteering would be more likely to donate to a disaster relief organization facing government crackdown.

Specifying every possible combination of individual- and organization-level attributes, however, would result in hundreds of hypotheses. The statistical method we are using lends itself to this kind of analysis. Rather than look at the effects of specific coefficients, we measure the difference in the marginal posterior distribution before and after including collections of covariates, which allows us to see which sets of individual characteristics shape the propensity to donate given organizational and structural heuristics.

We will run the upper level of our model using these sets of covariates:

- Public affairs: `Donation preferences ~ Public affairs knowledge + Public affairs activity`
- Political ideology: `Donation preferences ~ Political ideology`
- Social views: `Donation preferences ~ Public affairs trust + Social ideology + Religiosity`
- Charity and voluntarism: `Donation preferences ~ Charity trust + Charity activity + Volunteer activity + Activism activity + Association membership`
- Demographics: `Donation preferences ~ Gender + Marital status + Education + Income + Race + Age`

We will also include these groups of variables simultaneously as we check the performance of different models:

- `Donation preferences ~ (Public affairs) + (Political ideology)`
- `Donation preferences ~ (Public affairs) + (Political ideology) + (Social views)`
- `Donation preferences ~ (Public affairs) + (Political ideology) + (Social views) + (Charity and voluntarism)`
- `Donation preferences ~ (Public affairs) + (Political ideology) + (Social views) + (Charity and voluntarism) + (Demographics)`

## Inference criteria

> What criteria will you use to make inferences? Please describe the information you’ll use (e.g. specify the p-values, Bayes factors, specific model fit indices), as well as cut-off criterion, where appropriate. Will you be using one or two tailed tests for each of your analyses? If you are comparing multiple conditions or testing multiple hypotheses, will you account for this?

We will examine the aggregate marginal posterior distributions of the attribute levels and use 95% credible intervals to establish "significance." Effects are "significant" if the 95% credible intervals don't include 0. Similarly, marginal posterior distributions are "significantly" different if the 95% credible intervals don't overlap.

We will examine the marginal posterior distributions of the following models:

- Organizational and structural attribute levels with an intercept-only distribution of heterogeneity
- Organizational and structural attribute levels with competing sets of covariates in the distribution of heterogeneity

Finally, we will employ the posterior distribution of model parameters to conduct counterfactual analyses via a market simulator to determine optimal policies.

## Data exclusion

> How will you determine which data points or samples (if any) to exclude from your analyses? How will outliers be handled?

We ask potential participants a screening question early in the survey ("Q2.5: How often do you donate to charity"). If a participant responds that they give to once every few years or never, they will be disqualified from the study and the survey will end early.

We include one question ("*Q2.11*: Please select blue from the following list:") to monitor respondent attention. In our analysis we will exclude respondents who fail this question.

## Missing data

> How will you deal with incomplete or missing data?

Because all survey questions are required, we do not anticipate issues with incomplete or missing data.

## Exploratory analysis (optional)

> If you plan to explore your data set to look for unexpected differences or relationships, you may describe those tests here. An exploratory test is any test where a prediction is not made up front, or there are multiple possible tests that you are going to use. A statistically significant finding in an exploratory test is a great way to form a new confirmatory hypothesis, which could be registered at a later time.

In our third research question, we are interested in finding the optimal mix of attribute levels for NGOs interested in maximizing donations. This is an exercise of prediction, based on the organizational- and individual-level results of our first two sets of research questions. We answer this question using marketing simulations and suggesting direct policies for specific segments of the population (hypothetical example: NGOs should signal that they engage in financial transparency when they face government criticism if they want to reach religious liberals with a history of activism). We do not know *a priori* what the results of these simulations might be, so we do not state precise hypotheses.

---

# Script (optional)

> Upload an analysis script with clear comments. This optional step is helpful in order to create a process that is completely transparent and increase the likelihood that your analysis can be replicated. We recommend that you run the code on a simulated dataset in order to check that it will run without errors.

Nothing.

---

# Other

> If there is any additional information that you feel needs to be included in your preregistration, please enter it here.

Nothing.