### 2019-05-22 analysis/01_check-design.Rmd
Cleaned up experiment, added a check for the design and dummy coding and created Rmd for model runs

---
title: "Check experimental design"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Load the design and check level balance

```{r clean-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(here)

# Load design
design <- read_csv(here("data", "design.csv"))

# Check to make sure there is a balance across levels
as_tibble(design) %>%
  select(Organization:`Relationship with host government`) %>%
  gather(key = atts, value = lvls) %>%
  ggplot(aes(x = lvls)) +
  geom_bar() +
  facet_wrap(~ atts, scales = "free", nrow = 2)
```

# Dummy-code the design

```{r}
# Convert the attributes in design into factors
dummy_design <- as.data.frame(design)
for(i in 4:ncol(dummy_design)) {
  dummy_design[,i] <- as.factor(dummy_design[,i])
}

# Use the output of lm to get a dummy-coded version of the design
dummy_design$y <- matrix(rnorm(nrow(dummy_design)), ncol=1)
out <- lm(y ~ ., dummy_design, x = TRUE)
dummy_design <- out$x[,-1]
rownames(dummy_design) <- NULL

# Save dummy-coded design output
write.csv(dummy_design, here("data", "dummy_design.csv"))
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=90}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>