### 2019-06-20 analysis/02_clean-data.Rmd
Finished checking and cleaning data, exporting the cleaned final data

---
title: "Clean data"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Check data

## Load data

```{r clean-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(here)

col_names <- names(read_csv(here::here("data", "conjointqsf_final.csv")))
data <- read_csv(
  here::here("data", "conjointqsf_final.csv"),
  skip = 2,
  col_names = col_names
)
```

## Respondents who finished

```{r clean-data, warning=FALSE, message=FALSE}
data %>%
  summarize(
    total = n(),
    finished = sum(V10)
  )
```

Out of 1508 respondents, 1437 "finished" the survey and 71 didn't.

## Consent

```{r clean-data, warning=FALSE, message=FALSE}
data %>%
  count(Q1.1)

data %>%
  count(Q1.1, Q2.1)
```

30 respondents didn't give consent while 3 respondents gave consent and didn't respond starting in Q2.1.

## Screening

```{r clean-data, warning=FALSE, message=FALSE}
data %>%
  count(Q2.5)

data %>%
  count(Q2.5, Q2.6) %>% 
  filter(is.na(Q2.6) == TRUE)
```

An additional 12 respondents drop out by Q2.5 (a total of 45). 385 respondents are screened at Q2.5. 2 more respondents drop off by Q2.6 (a total of 47).

## Check display logic

```{r clean-data, warning=FALSE, message=FALSE}
data %>%
  count(Q2.9)

data %>%
  count(Q2.9, Q2.10)
```

8 more respondents have dropped out (a total of 55). The 445 people who don't volunteer in the past 12 months properly skip Q2.10.

## Attention check

```{r clean-data, warning=FALSE, message=FALSE}
data %>%
  count(Q2.11)
```

6 respondents don't make the attention check.

## Confirm ages

```{r clean-data, warning=FALSE, message=FALSE}
data %>%
  count(Q5.17)
```

By the end of the survey, 44 more respondents have dropped out (a total of 99) but no one is under 18, as they indicated as part of the consent.

# Clean data

## Clean based on initial criteria

```{r clean-data, warning=FALSE, message=FALSE}
data_clean <- data %>%
  filter(V10 == 1) %>%         # Keep respondents who completed the entire survey
  filter(Q1.1 == 1) %>%        # Keep respondents who gave consent
  filter(Q2.5 %in% c(1:5)) %>% # Keep respondents who passed the screener
  filter(Q2.11 == 3)           # Keep respondents who passed the attention check
```

## Confirm the data is clean

```{r clean-data, warning=FALSE, message=FALSE}
# Finished respondents
data_clean %>% 
  summarize(
    total = n(),
    finished = sum(V10)
  )

# Consent
data_clean %>%
  count(Q1.1)

# Screening
data_clean %>%
  count(Q2.5)

# Attention check
data_clean %>% 
  count(Q2.11)
```

We have 1016 total respondents who have finished the entire survey, given consent, and passed the screening and attention check.

## Confirm the choice data is clean

```{r clean-data, warning=FALSE, message=FALSE}
data_clean %>%
  count(vers_CBCONJOINT) %>%
  count(n)
```

The design version appears to be properly assigned at random.

```{r clean-data, warning=FALSE, message=FALSE}
data_clean %>%
  select(Q4.1:Q4.12) %>%
  gather(key = "task", value = "choice") %>%
  count(choice)
```

Everyone in the clean data responds to every choice task and the "none of the above" option doesn't appear to be abused.

```{r clean-data, warning=FALSE, message=FALSE}
design_orig <- read_csv(here("data", "design.csv")) %>% 
  rename(
    version = question_set,
    task = question,
    alt = choice
  ) %>% 
  gather(key = "atts", value = "levels", -c(version:alt))

design_surv <- data_clean %>% 
  select(contains("CBCONJOINT"), -revision_CBCONJOINT) %>% 
  gather(key = "temp", value = "level_names", -vers_CBCONJOINT) %>% 
  separate(temp, into = c("atts", "task", "alt"), sep = "\\.") %>% 
  separate(alt, into = c("alt", "temp")) %>% 
  select(vers_CBCONJOINT, task, alt, atts, level_names) %>% 
  rename(version = vers_CBCONJOINT) %>% 
  mutate(
    task = as.numeric(task),
    alt = as.numeric(alt),
    atts = recode(
      atts,
      "732d72a6-25a3-4194-8416-461acae30837" = "Organization",
      "bd54b90a-efb3-4ac8-b65a-1ebeefcaf703" = "Issue area",
      "a6836e28-fd7a-4b71-b65e-f6269e5c883f" = "Financial transparency",
      "c2249c1a-11fa-4395-a0a3-61d9648a2e55" = "Accountability",
      "db44c406-390e-4fc4-875b-65796aa95ffa" = "Funding sources",
      "85f14c50-61cc-461c-8e77-19453b83d76d" = "Relationship with host government"
    )
  ) %>%
  distinct() %>% 
  arrange(version, task, alt)

design_orig %>% 
  count(atts, levels) %>% 
  mutate(prop = n/sum(n))

design_surv %>% 
  count(atts, level_names) %>% 
  mutate(prop = n/sum(n))

design_surv %>% 
  left_join(
    design_orig,
    by = c("version", "task", "alt", "atts")
  ) %>% 
  count(atts, levels, level_names)
```

The design matrix as reported in the survey is correct. We can use the separate design file matched by design version for estimation.

## Save clean data

```{r clean-data, warning=FALSE, message=FALSE}
data <- data_clean

write_csv(data, here("data", "final_data.csv"))
```

<!-- select(-c("732d72a6-25a3-4194-8416-461acae30837.1.1_CBCONJOINT":"revision_CBCONJOINT")) -->
<!-- design <- read_csv( -->
<!--   here::here("data", "conjointqsf_final.csv"), -->
<!--   skip = 2, -->
<!--   col_names = col_names -->
<!-- ) %>%  -->
<!--   select(contains("CBCONJOINT")) -->

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=90}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>