### 2019-06-20 analysis/03_run-models.Rmd

---
title: "Run models"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    pandoc_args:
      - "--default-image-extension=png"
editor_options: 
  chunk_output_type: console
---

```{r run-models warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(ggraph)
library(ggdag)
library(here)

# General settings
source(here("analysis", "options.R"))

# Make all the randomness reproducible
set.seed(1234)
```

# Run models