### 2019-10-29 manuscript/manuscript.Rmd
Added interpretations to second set of questions

---
title: "Why Donors Donate: Disentangling Organizational and Structural Heuristics for International Philanthropy"
author:
- name: Suparna Chaudhry
  affiliation: Christopher Newport University
  url: http://www.suparnachaudhry.com/
  email: suparna.chaudhry@cnu.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Brigham Young University
  url: https://www.andrewheiss.com/
  email: andrew_heiss@byu.edu
date: March 24, 2019
git-repo: https://github.com/andrewheiss/why-donors-donate
reference-section-title: References
toc: false
---

# Brief note on our overarching project

We are interested in the organizational and structural factors that influence donors to contribute money to international nongovernmental organizations (INGOs) and other nonprofit organizations. Existing research has found that donors respond to a host of heuristics and signals when deciding whether to donate to an NGO, since seeking out complete information about an organization's deservingness and efficiency is costly and time-consuming.

One common signal for donors is organizational efficiency, such as the amount an organization spends on overhead costs, whether an organization makes its finances public, or whether an organization participates in voluntary accountability practices, such as listing organizational details on websites like GuideStar and Charity Navigator. Another heuristic for deservingness is the issue area an NGO engages in. Organizations that address humanitarian issues such as disaster relief and international aid and development have broader appeal to donors who might shy away from NGOs dealing with more political causes like human rights. Finally, the source of NGO funding can also serve as a helpful signal of how effective and useful an individual donor's action might be. Donors are likely to not contribute to NGOs that rely heavily on government grants since their donation might feel more impersonal and less needed, while donors are likely to give to organizations that rely primarily on private donations given that they may feel personally invested in the organization. Each of these signals is *organizational*. NGOs typically have direct control over its organizational practices, mission and issue area, and funding sources.

NGOs have less control over *structural* issues, and the role of political and legal institutions in donor decision making has been understudied. However, over the past decade international NGOs have become increasingly politicized and legally restricted around the world. Dozens of countries have imposed legal restrictions on international NGO funding, missions, and programs, and in many countries, NGOs have been forced to close down operations and leave. How do individual donors respond when NGOs are kicked out of their host countries? How does legal crackdown abroad affect perceptions of an NGO's deservingness or efficiency?

In our larger project (which has unfortunately faced some substantial delays), we will use a conjoint survey experiment to disentangle these different questions. We theorize that organizational characteristics like managerial practices, issue area, and funding sources all influence the decision to donate to an NGO (see @fig:big-dag). At the same time, however, we propose that structural characteristics like an NGO's host-country political and legal environment can also serve as donor heuristics. Donors who see that an NGO has been kicked out of a country or that an NGO faces general legal trouble abroad will use that knowledge to decide whether or not to donate.

![Causal diagram showing causes of donor behavior. D = donation to NGO (outcome); G = NGO-government relationship, I = NGO issue area, F = NGO funding source, O = NGO organizational and managerial practices.](output/figures/causal-path.pdf){#fig:big-dag}

Identifying the causal link between structural factors and donor behavior is trickier than measuring the link between organizational factors and donor behavior, however, because (1) NGOs have less direct control over their political environments, and (2) organizational characteristics like issue area and funding can both sour the NGO-government relationship (as we explain in further detail below, governments are more likely to crack down on NGOs that address human rights and that receive government funding). @fig:big-dag helps disentangle this complex relationship. Organizational factors (**O**, or organizational practices like financial disclosure and accountability practices; **I**, or issue area; and **F**, or funding) all influence the decision to donate (**D**). Structural factors like an NGO's relationship with its host government (**G**) also influence donation behavior, but the relationship itself is shaped by both funding and issue area.

Through our conjoint survey experiment, we will be able to measure how strongly each of these factors shapes the propensity to donate and discover how much donors are swayed by the political and legal environments of NGO host countries.

What follows here is a smaller foray into one aspect of this causal diagram. We use a vignette-based survey experiment to measure donor response to just crackdown, issue area, and funding.

We will field our larger conjoint survey experiment very soon and we look forward to your comments and suggestions before we launch it!

# Methods

Conjoint analysis forces a consumer to choose between alternatives, revealing which components of a choice are the most important to that consumer. Conjoint analysis has been widely used within social sciences to explain causality and has recently begun to be used in political and public policy fields. Helfer (2016) and Mummolo (2016) both used conjoint analysis to assess consumer perceptions of political communications. Conjoint is key in political research because it captures multidimensional causal relations.

Knudsen and Johannesson (2018) argue that political communication research requires conjoint analysis because it accounts for the masking effects of factors as well as the complexity of a respondent's judgment of credibility. Because we also want to capture the complexity of a respondent's judgment of whether they should donate to an NGO or not, we used conjoint analysis to draw our conclusions.

# Results

### First Set of Questions
The first set of questions (hypotheses) can be answered using the results of the intercept-only model.

We hypothesized that each of these structural and organizational characteristics would have a significant effect on donor behavior, though we were unsure of the direction of that effect.

![](output/figures/combined-results.png)

The following plots zoom in on the strength and direction of the effects of these characteristics.

Compared to when an NGO has a friendly relationship with the host government, donors who see that an NGO faces general legal trouble abroad (under crackdown or criticized) are less likely to donate.

![](output/figures/relationship-with-gov.png)

Donors are more likely to donate when organizations are accountable and financially transparent.

![](output/figures/managerial-practices.png)

Donors are most likely to donate when an NGO is funded primarily by many small private donations. Donors are less likely to donate when the organization is subsidized by government grants, and even less likely when it is funded by a few private donors.

![](output/figures/funding-sources.png)

Of all issue areas, donations to emergency response were most likely, followed by human rights and then environment. Refugee relief was the least likely issue area to be donated to.

![](output/figures/issue-areas.png)

All organizations were positively percieved by respondents. The Red Cross was significantly more favored.

![](output/figures/organization.png)

### Second Set of Questions
The second set of questions (confirmatory) can be answered using the following upper-level models:
  - Public affairs: `Donation preferences ~ Public affairs knowledge + Public affairs activity`
    ![](output/figures/public-affairs.png)

Those that haven't traveled internationally are significantly less likely to donate to NGOs than those who have.
![](output/figures/traveled.png)

Of all media sources, those that follow the news through an app or email newsletter are more likely to donate in general. Interestingly, those that read newsletters are the most likely to donate overall. They feel most comfortable donating to environmental causes. They care significantly less about managerial practices and funding sources than other groups. 
      ![](output/figures/following-news.png)

When it comes to donating to organizations, those whose top news medium is online sources (excluding social media) are least likely to donate.
![](output/figures/following-news2.png)

  - Political ideology: `Donation preferences ~ Political ideology`

Those that are conservative are significantly more likely to donate to the Red Cross.
![](output/figures/political-ideology.png)

  - Social views: `Donation preferences ~ Public affairs trust + Social ideology + Religiosity`

![](output/figures/social-views.png)

Hindus are very unlikely to donate to human rights or refugee relief, whereas Muslims are very likely to donate to human rights and refugee relief. Muslims are also very willing to donate to organizations that are facing government criticism or crackdown in the country. 
  ![](output/figures/religion.png)

  - Charity and voluntarism: `Donation preferences ~ Charity trust + Charity activity + Volunteer activity + Activism activity + Association membership`

Those that have a history of either family or personal activism are most likely to donate to organizations. Those that haven't volunteered in the past 12 months are also more likely to donate.   
![](output/figures/charity-and-voluntarism.png)  
Those that are in a professional or humanitarian organization are most likely to donate. It's odd that those that belong to an environmental organization are less likely to donate to environmental NGOs.
  ![](output/figures/other-organizations.png)

  - Demographics: `Donation preferences ~ Gender + Marital status + Education + Income + Race + Age`

Demographics are broken down further and analyzed below.
  ![](output/figures/demographics.png)
Note: not sure if "other" indicates "male" or if male is the intercept everything else is compared to.
  ![](output/figures/gender.png)  
Overall, those that are widowed are less likely to donate to organizations than others. They also don't pay attention to financial transparency and accountability, though those that have never married do.
  ![](output/figures/marital-status.png)  
Financial transparency was important to everyone regardless of education level. 
  ![](output/figures/education.png)
Income doesn't impact the various attributes except when it comes to organization. There is a lot more spread when it comes to choosing an organization.
  ![](output/figures/income.png)
Not seeing obvious patterns across races.
    ![](output/figures/race.png)
  Across all ages, people are most likely to donate to emergency response efforts more than any other issue area. 
  ![](output/figures/age.png)

  - `Donation preferences ~ (Public affairs) + (Political ideology)`

Those that read printed news are least likely to donate to any of these organizations. Those that follow international news multiple times are day are more likely to donate and are less concerned when an organization is criticized by the government.  
![](output/figures/public-political.png)

  - `Donation preferences ~ (Public affairs) + (Political ideology) + (Social views)`

For the most part, these plots have too much information to gain useful insight.
![](output/figures/public-political-social.png)

  - `Donation preferences ~ (Public affairs) + (Political ideology) + (Social views) + (Charity and voluntarism)`

![](output/figures/public-political-social-charity.png)

  - `Donation preferences ~ (Public affairs) + (Political ideology) + (Social views) + (Charity and voluntarism) + (Demographics)`

![](output/figures/public-political-social-charity-demo.png)

### Third Set of Questions
The third set of questions (exploratory) can be answered using a market simulator.