### 2020-01-14 analysis/04_figs_tables.Rmd

---
title: "Figures and tables for paper"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    pandoc_args:
      - "--default-image-extension=png"
editor_options: 
  chunk_output_type: console
---

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(donorheuristics)
# library(rstan)
# library(bayesplot)
library(tidybayes)
library(broom)
# library(bayesm)
# library(ggridges)
library(ggraph)
library(ggdag)
library(pander)
library(patchwork)
library(scales)
library(here)

# General settings
source(here("analysis", "options.R"))

# Make all the randomness reproducible
set.seed(1234)

# Load data
results <- read_rds(here("data", "derived_data", "final_data.rds"))

# Name labels
level_labels <- c(
  str_c("Organization: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross")),
  str_c("Issue area: ", c("Environment", "Human rights", "Refugee relief")),
  str_c("Organizational practices: ", c("Financial transparency", "Accountability")),
  str_c("Funding sources: ", c("Handful of wealthy private donors", "Government grants")),
  str_c("Relationship with host government: ", c("Criticized", "Under crackdown"))
)

level_lookup <- tibble(levels = level_labels) %>% 
  separate(levels, c("category", "attribute"), sep = ": ") %>% 
  mutate(attributes_long = paste0(category, ": ", attribute),
         i = 1:n()) %>% 
  mutate_at(vars(category, attribute, attributes_long), list(fct_inorder))

draws <- read_rds(here("analysis", "output", "posterior_draws", 
                       "intercept_noncentered.rds")) %>% 
  mutate(model = "Intercept only") %>% 
  left_join(level_lookup, by = "i")
```

```{r}
diff_cols <- function(col1, col2) {
  posterior_diffs <- draws %>% 
    filter(attribute %in% c({{col1}}, {{col2}})) %>% 
    select(.chain, .iteration, Theta, attribute) %>% 
    pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
    mutate(diff_theta = .[[4]] - .[[3]],
           diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

  diff_theta <- mean(posterior_diffs$diff_theta_e)
  p_greater0 <- mean(posterior_diffs$diff_theta_e > 0)
  p_less0 <- mean(posterior_diffs$diff_theta_e < 0)
  p_not0 <- ifelse(median(posterior_diffs$diff_theta_e) > 0, p_greater0, p_less0)

  phrase <- paste0("∆θ = ", round(diff_theta, 3), "; p(∆ ≠ 0) = ", round(p_not0, 3))

  return(list(posterior_diffs = posterior_diffs, diff_theta = diff_theta,
              p_greater0 = p_greater0, p_less0 = p_less0, p_not0 = p_not0,
              phrase = phrase))
}
```

# Causal pathway

Our theory and hypotheses are laid out in the causal pathway below. Organizational factors (**O**, or organizational practices like financial disclosure and accountability practices; **I**, or issue area; and **F**, or funding) all influence the decision to donate (**D**). Structural factors like an NGO's relationship with its host government (**G**) also influence donation behavior, but the relationship itself is shaped by both funding and issue area.

```{r causal-dag, fig.width=13/3, fig.height=2}
node_colors <- tribble(
  ~name, ~type,
  "D", "Outcome",
  "I", "Organizational factor",
  "F", "Organizational factor",
  "O", "Organizational factor",
  "G", "Structural factor"
) %>% 
  mutate(type = fct_inorder(type))

# Set up DAG structure
theory_dag <- dagify(D ~ I + F + O + G,
                     G ~ I + F,
                     outcome = "D",
                     exposure = "C") %>% 
  tidy_dagitty(layout = "dh", seed = 1234)

# Add node types/colors to DAG data
theory_dag$data <- theory_dag$data %>% 
  left_join(node_colors, by = "name")

# Make DAG plot
plot_dag <- ggplot(theory_dag, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(size = 6, mapping = aes(color = type)) +
  geom_dag_edges(start_cap = circle(4, "mm"),
                 end_cap = circle(4, "mm")) +
  geom_dag_text(size = pts(6), family = "Roboto Condensed", fontface = "bold") +
  scale_color_manual(values = c(clrs$blue, clrs$red, clrs$orange), name = NULL) +
  theme_ngo() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank())

# Show and save plot
plot_dag
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.pdf"),
       width = 13/3, height = 2, units = "in", device = cairo_pdf)
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.png"),
       width = 13/3, height = 2, units = "in", type = "cairo", dpi = 300)
```

# Sample details

```{r}
vars_to_summarize <- tribble(
  ~category, ~variable, ~clean_name,
  "Demographics", "Q5.12", "Gender",
  "Demographics", "Q5.17", "Age",
  "Demographics", "Q5.13", "Marital status",
  "Demographics", "Q5.14", "Education",
  "Demographics", "Q5.15", "Income",
  "Attitudes toward charity", "Q2.5", "Frequency of donating to charity",
  "Attitudes toward charity", "Q2.6", "Amount of donations to charity last year",
  "Attitudes toward charity", "Q2.7_f", "Importance of trusting charities",
  "Attitudes toward charity", "Q2.8_f", "Level of trust in charities",
  "Attitudes toward charity", "Q2.10", "Frequency of volunteering",
  "Politics, ideology, and religion", "Q2.1", "Frequency of following national news",
  "Politics, ideology, and religion", "Q5.7", "Traveled to a developing country",
  "Politics, ideology, and religion", "Q5.1", "Voted in last election",
  "Politics, ideology, and religion", "Q5.6_f", "Trust in political institutions and the state",
  "Politics, ideology, and religion", "Q5.2_f", "Political ideology",
  "Politics, ideology, and religion", "Q5.4", "Involvement in activist causes",
  "Politics, ideology, and religion", "Q5.8", "Frequency of attending religious services",
  "Politics, ideology, and religion", "Q5.9", "Importance of religion"
)

summarize_factor <- function(x) {
  output <- table(x) %>% 
    as_tibble() %>% 
    magrittr::set_colnames(., c("level", "count")) %>% 
    mutate(level = factor(level, levels = levels(x), ordered = TRUE)) %>%
    mutate(prop = count / sum(count),
           nice_prop = scales::percent(prop))

  return(list(output))
}

participant_summary <- results %>% 
  select(one_of(vars_to_summarize$variable)) %>% 
  summarize_all(summarize_factor) %>% 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "details") %>% 
  left_join(vars_to_summarize, by = "variable") %>% 
  unnest(details) %>% 
  mutate(level = as.character(level)) %>% 
  mutate(level = case_when(
    variable == "Q2.7_f" & level == "1" ~ "1 (not important)",
    variable == "Q2.7_f" & level == "7" ~ "7 (important)",
    variable == "Q2.8_f" & level == "1" ~ "1 (no trust)",
    variable == "Q2.8_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.6_f" & level == "1" ~ "1 (no trust)",
    variable == "Q5.6_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.2_f" & level == "1" ~ "1 (extremely liberal)",
    variable == "Q5.2_f" & level == "7" ~ "7 (extremely conservative)",
    variable == "Q5.15" & level == "Less than median" ~ "Less than 2017 national median ($61,372)",
    variable == "Q5.17" & level == "Less than median" ~ "Less than 2017 national median (36)",
    TRUE ~ level
  )) %>% 
  mutate(clean_name_shrunk = ifelse(clean_name == lag(clean_name), "", clean_name),
         clean_name_shrunk = ifelse(is.na(clean_name_shrunk), 
                                    clean_name[1], 
                                    clean_name_shrunk),
         category_shrunk = ifelse(category == lag(category), "", category),
         category_shrunk = ifelse(is.na(category_shrunk), 
                                    category[1], 
                                    category_shrunk))
```

```{r results="asis"}
participant_summary %>% 
  select(" " = category_shrunk, "  " = clean_name_shrunk, 
         "Response" = level, "N" = count, "%" = nice_prop) %>% 
  pandoc.table.return(caption = 'Summary of individual respondent characteristics {#tbl:sample-details}',
                      justify = "lllcc") %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-sample-details.md"))
```

# Results

## Baseline probabilities

```{r probs-baseline, fig.width=4.5, fig.height=4}
orgs <- draws %>% 
  filter(category == "Organization") %>% 
  mutate(prob = plogis(Theta))

fig_baseline <- ggplot(orgs, aes(x = prob, y = attribute)) +
  geom_halfeyeh(.width = 0.95, aes(fill = attribute)) +
  scale_fill_manual(values = c(clrs$yellow, clrs$olive, clrs$lime, clrs$red),
                    guide = FALSE) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Baseline probability of selecting organization", y = NULL) +
  theme_ngo()

fig_baseline
ggsave(fig_baseline, filename = here("analysis", "output", "figures", "probs-org.pdf"),
       width = 13/3, height = 3, units = "in", device = cairo_pdf)
ggsave(fig_baseline, filename = here("analysis", "output", "figures", "probs-org.png"),
       width = 13/3, height = 3, units = "in", type = "cairo", dpi = 300)
```

```{r}
orgs %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(tidy = data %>% map(~ median_hdci(.$prob))) %>% 
  unnest(tidy)
```

## Coefficients and marginal effects

```{r first-hypotheses-coefficients, fig.width=26/3, fig.height=5.5}
diff_refugee_hr <- diff_cols("Refugee relief", "Human rights")
diff_hr_env <- diff_cols("Human rights", "Environment")
diff_env_refugee <- diff_cols("Environment", "Refugee relief")

plot_issue <- ggplot(filter(draws, category == "Issue area"), 
                     aes(x = exp(Theta), y = attribute)) + 
  stat_halfeyeh(.width = 0.95, fill = clrs$blue) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.7, ymin = 2.1, ymax = 2.9, size = 0.25) +
  # annotate(geom = "label", x = 0.7, y = 2.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_refugee_hr$phrase) +
  # annotate(geom = "linerange", x = 0.8, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.8, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_hr_env$phrase) +
  # annotate(geom = "linerange", x = 0.585, ymin = 1.1, ymax = 2.9, size = 0.25) +
  # annotate(geom = "label", x = 0.585, y = 2.05, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_env_refugee$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Issue area",
       subtitle = "Reference group = emergency response") +
  theme_ngo()

diff_govt_richpeople <- diff_cols("Government grants", 
                                  "Handful of wealthy private donors")

plot_funding <- ggplot(filter(draws, category == "Funding sources"), 
                       aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$maroon) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.825, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.825, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_govt_richpeople$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Funding sources",
       subtitle = "Reference group = many small private donations") +
  theme_ngo()

diff_account_trans <- diff_cols("Accountability", "Financial transparency")

plot_org <- ggplot(filter(draws, category == "Organizational practices"), 
                   aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$orange) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 1.85, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 1.85, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_account_trans$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Organizational practices",
       subtitle = "Reference groups = no accountability; no transparency") +
  theme_ngo()

diff_govt <- diff_cols("Under crackdown", "Criticized")

plot_government <- ggplot(filter(draws, category == "Relationship with host government"), 
                          aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$teal) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.75, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.75, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_govt$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Relationship with host government",
       subtitle = "Reference group = friendly relationship with government") +
  theme_ngo()

plot_first_hypotheses <- (plot_org + plot_issue + plot_funding + plot_government)

plot_first_hypotheses
ggsave(plot_first_hypotheses, filename = here("analysis", "output", "figures", "results-h1-4.pdf"),
       width = 26/3, height = 5.5, units = "in", device = cairo_pdf)
ggsave(plot_first_hypotheses, filename = here("analysis", "output", "figures", "results-h1-4.png"),
       width = 26/3, height = 5.5, units = "in", type = "cairo", dpi = 300)
```

```{r}
draws %>% 
  filter(category != "Organization") %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(tidy = data %>% map(~ median_hdci(.$Theta))) %>% 
  unnest(tidy) %>% 
  mutate(exp_y = exp(y))
```

## Differences in means

According to Kruschke 8.4, we can look at the posterior distribution of the difference in thetas, then calculate the proportion of those differences that are greater/less than 0

$$
\operatorname{mean}(\theta_\text{Government grants} - \theta_\text{Private donors}) = 
$$

$$
\Delta \theta = 4; p(\Delta > 0) = 5
$$

```{r eval=FALSE}
col1 <- "Government grants"
col2 <- "Handful of wealthy private donors"

posterior_diffs <- draws %>% 
  filter(attribute %in% c({{col1}}, {{col2}})) %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]],
         diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

mean(posterior_diffs$diff_theta_e)

asdf <- draws %>% 
  filter(category == "Funding sources") %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]]) %>% 
  mutate(diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  summarize(bloop = exp(mean(Theta)),
            bloop1 = mean(exp(Theta)))

mean(asdf$diff_theta_e)
tidybayes::hdi(asdf$diff_theta_e)

mean(asdf$diff_theta > 0)
mean(asdf$diff_theta_e > 0)

asdf <- draws %>% 
  filter(category == "Organizational practices") %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]]) %>% 
  mutate(diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

mean(asdf$diff_theta > 0)
mean(asdf$diff_theta_e > 0)

ggplot(asdf, aes(x = diff_theta_e, y = 1)) +
  geom_halfeyeh(.width = 0.95)

draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  summarize(bloop = mean(Theta),
            bloop1 = exp(mean(Theta)))

qwer <- draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(hdi = data %>% map(~ exp(tidybayes::hdi(.$Theta, .width = 0.95))))
asdf %>% 
  unnest(hdi)
mean(asdf$Theta)
tidybayes::hdi(asdf$Theta, .width = 0.95)
```