### 2020-06-19 analysis/shiny_walkthrough.Rmd
Added a Shiny walk through document

---
title: "Shiny Walkthrough"
output: html_notebook
---

## Shiny

Shiny dashboard structure (UI - user interface)

3 Main parts:

* Header - dashboardHeader()
* Sidebar - dashboardSidebar()
* Body - dashboardBody()

## How to build a Sidebar Dashboard

sidebar <- dashboardSidebar(
  sidebarMenu(
    menuItem("Data",
      tabName = "data") #Used for linking in the body
  )
)

You can have users input information - 
Different input functions:
selectInput() # Allows user to select a singular item from a list
textInput() # Allows user to input text
dateInput() # Allows user to input a date
checkboxInput() # Allows user to select multiple items

How to build user input

selectInput(
  inputId = "object", # This is the name you described the object elsewhere in the application
  label = "Favorite Character", # The label you want to show to the user
  choices = c(...) # List the choices the user has to pick from
)

### Render Functions 
Functions that you use in your application's server side code, assigning them to outputs that appear in your user interface.

renderPrint()
renderText()
renderPlot()
renderUI() # HTML or a shiny tag object

## Example Using Star Wars df
```{r}
library(shiny)
library(tidyverse)
library(flexdashboard)
library(shinydashboard)

sidebar <- dashboardSidebar(
  selectInput(
    inputId = "name",
    label = "Favorite Character",
    choices = c(starwars$name)
  )
)

body <- dashboardBody(
  textOutput("name")
)

ui <- dashboardPage(header = dashboardHeader(),
                    sidebar = sidebar,
                    body = body
                    )

server <- function(input, output) {
  output$name <- renderText({
      input$name
    })
}

shinyApp(ui, server)

```