### 2020-06-24 analysis/flexdashboard_walkthrough.Rmd
Create a flexdashboard walkthrough

---
title: "Untitled"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
```

Column {data-width=650}
-----------------------------------------------------------------------
### Chart A

```{r}
starwars10 <- starwars %>% 
  filter(height >= 180)

starwars10 %>% 
  ggplot(aes(x = name, y = height)) +
    geom_col() +
    coord_flip()
```

Column {data-width=350}
-----------------------------------------------------------------------
### Chart B

```{r}

```

### Chart C

```{r}

```