### 2020-06-24 analysis/flexdashboard_walkthrough.Rmd
Input a description about Flexdashboards.

---
title: "Untitled"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

Flexdashboards are often used along with shiny to build interactive dashboards. Dashboards are divided into columns and rows, with output components delineated using level 3 markdown headers (###). By default, dashboards are laid out within a single column, with charts stacked vertically within a column and sized to fill available browser height.

To lay out charts using multiple columns you introduce a level 2 markdown header (--------------) for each column.

You can also choose to orient dashboards row-wise rather than column-wise by specifying the orientation: rows option.

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
```

Column {data-width=650}
-----------------------------------------------------------------------
### Chart A

```{r}
starwars10 <- starwars %>% 
  filter(height >= 180)

starwars10 %>% 
  ggplot(aes(x = name, y = height)) +
    geom_col() +
    coord_flip()
```

Column {data-width=350}
-----------------------------------------------------------------------
### Chart B

```{r}

```

### Chart C

```{r}

```