### 2020-10-28 analysis/05_market_simulator.Rmd
Cleaned up code for all the new rendered objects for uploadable configurations

---
title: "Market simulator"
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
runtime: shiny
---

```{r load-packages, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(flexdashboard)
library(plotly)
library(shiny)
```

```{r load-data, message=FALSE}
# Gammas of the "everything" model
gammas <- read_rds(
  here::here('analysis', 'output', 'posterior_draws', 'public_political_social_charity_demo.rds')
  ) %>% 
  group_by(i, j) %>% 
  summarize(gamma_mean = mean(Gamma)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = j, values_from = gamma_mean) %>% 
  select(-i) %>% 
  t()

# Organization configuration
organization_data <- reactive({
  if (is.null(input$org_update$datapath)) {
    organization_data <- read_csv(
      here::here('analysis', 'output', 'configurations', 'organization_config01.csv')
    )
  } else {
    organization_data <- read_csv(input$org_update$datapath)
  }
})

# Persona configuration
persona_data <- reactive({
  if (is.null(input$per_update$datapath)) {
    persona_data <- read_csv(
      here::here('analysis', 'output', 'configurations', 'persona_config01.csv')
    )
  } else {
    persona_data <- read_csv(input$per_update$datapath)
  }
})
```

```{r organization-reactive-table}
# Assign level labels
org_table <- data.frame(
  levels = c(
    # Organization
    "Amnesty International",
    "Greenpeace", 
    "Oxfam", 
    "Red Cross",
    # Issue area
    "Environment", 
    "Human rights",
    "Refugee relief",
    # Financial transparency
    "Engages in financial transparency",
    # Financial accountability
    "Engages in financial accountability",
    # Funding sources
    "Handful of wealthy private donors", 
    "Government grants",
    # Relationship with host government
    "Criticized by government",
    "Under government crackdown"
  )
)

# Organization configuration reactive element
org_reactive <- reactive({
  # Organization 1
  org_table$org1_value <- 0
  org_table$org1_value[org_table$levels %in% input$org1_org] <- 1
  org_table$org1_value[org_table$levels %in% input$org1_issue] <- 1
  org_table$org1_value[org_table$levels %in% input$org1_fin_trans] <- 1
  org_table$org1_value[org_table$levels %in% input$org1_fin_account] <- 1
  org_table$org1_value[org_table$levels %in% input$org1_funding] <- 1
  org_table$org1_value[org_table$levels %in% input$org1_host_gov] <- 1

  # Organization 2
  org_table$org2_value <- 0
  org_table$org2_value[org_table$levels %in% input$org2_org] <- 1
  org_table$org2_value[org_table$levels %in% input$org2_issue] <- 1
  org_table$org2_value[org_table$levels %in% input$org2_fin_trans] <- 1
  org_table$org2_value[org_table$levels %in% input$org2_fin_account] <- 1
  org_table$org2_value[org_table$levels %in% input$org2_funding] <- 1
  org_table$org2_value[org_table$levels %in% input$org2_host_gov] <- 1

  # Organization 3
  org_table$org3_value <- 0
  org_table$org3_value[org_table$levels %in% input$org3_org] <- 1
  org_table$org3_value[org_table$levels %in% input$org3_issue] <- 1
  org_table$org3_value[org_table$levels %in% input$org3_fin_trans] <- 1
  org_table$org3_value[org_table$levels %in% input$org3_fin_account] <- 1
  org_table$org3_value[org_table$levels %in% input$org3_funding] <- 1
  org_table$org3_value[org_table$levels %in% input$org3_host_gov] <- 1

  # Organization 4
  org_table$org4_value <- 0
  org_table$org4_value[org_table$levels %in% input$org4_org] <- 1
  org_table$org4_value[org_table$levels %in% input$org4_issue] <- 1
  org_table$org4_value[org_table$levels %in% input$org4_fin_trans] <- 1
  org_table$org4_value[org_table$levels %in% input$org4_fin_account] <- 1
  org_table$org4_value[org_table$levels %in% input$org4_funding] <- 1
  org_table$org4_value[org_table$levels %in% input$org4_host_gov] <- 1
  org_table
})
```

```{r persona-reactive-table}
# Assign covariate labels
per_table <- data.frame(
  levels = c(
    # Intercept
    "Intercept",
    # Follow national news
    "Once a week",
    "At least once a day",
    # Follow international news
    "Sometimes",
    "Always",
    # Mediums to follow news
    "TV",
    "Print",
    "Online (not social media)",
    "Social media",
    "Radio",
    "Email newsletter",
    "News app",
    # Follow government and public affairs
    "Often", 
    # Traveled to a developing country
    "Traveled to a developing country", 
    # Liberal to Conservative
    "Somewhat liberal",
    "Slightly liberal",
    "Moderate",
    "Slightly conservative",
    "Somewhat conservative",
    "Extremely conservative",
    # Trust in political institutions
    "Very little trust in political institutions",
    "Little trust in political institutions",
    "Neutral trust in political institutions",
    "Some trust in political institutions",
    "A lot of trust in political institutions",
    "Complete trust in political institutions",
    # People should be more charitable
    "Disagree",
    "Somewhat disagree",
    "Neutral",
    "Somewhat agree",
    "Agree",
    "Strongly agree",
    # Religious attendance
    "Rarely attend",
    "Attend at least once a month",
    # Importance of religion
    "Important",
    # Religion
    "Protestant",
    "Christian Orthodox",
    "Jewish",
    "Muslim",
    "Sikh",
    "Hindu",
    "Buddhist",
    "Atheist",
    "Other religion",
    # Importance of trusting charities
    "Very unimportant",
    "Somewhat unimportant",
    "Neutral importance",
    "Somewhat important",
    "Very important",
    "Essential",
    # Trust in charities
    "Very little trust in charities",
    "Little trust in charities",
    "Neutral trust in charities",
    "Some trust in charities",
    "A lot of trust in charities",
    "Complete trust in charities",
    # Donation frequency
    "Donate at least once a month", 
    # Donated last year
    "$50-$99",
    "$100-$499",
    "$500-$999",
    "$1,000-$4,999",
    "$5,000-$9,999",
    "$10,000+",
    # Volunteered in past 12 months
    "Volunteered in past 12 months",
    # Volunteer frequency
    "Rarely volunteer",
    "Volunteer more than once a month, less than once a year",
    "Volunteer at least once a month",
    # History of activism
    "Personally involved",
    # History of family activism
    "Family involved",
    # Church or religious organization
    "Inactive church member",
    "Active church member",
    # Sport or recreational organization
    "Inactive sports or rec organization member",
    "Active sports or rec organization member", 
    # Art, music, or educational organization
    "Inactive art, music, or educational organization member",
    "Active art, music, or educational organization member",
    # Labor union
    "Inactive union member",
    "Active union member",
    # Political party
    "Inactive political party member",
    "Active political party member",
    # Environmental organization
    "Inactive environmental organization member",
    "Active environmental organization member",
    # Professional association
    "Inactive professional association member",
    "Active professional association member",
    # Humanitarian or charitable organization
    "Inactive humanitarian or charitable organization member",
    "Active humanitarian or charitable organization member",
    # Consumer organization
    "Inactive consumer organization member",
    "Active consumer organization member",
    # Other organization
    "Inactive member in another organization",
    "Active member in another organization",
    # Gender
    "Female",
    "Transgender",
    "Prefer not to say",
    "Other gender",
    # Marital status
    "Widowed",
    "Divorced",
    "Separated",
    "Never married",
    # Education
    "High school graduate", 
    "Some college",
    "2 year degree",
    "4 year degree",
    "Graduate or professional degree",
    "Doctorate",
    # Income
    "More than median of $61,372", 
    # Race
    "White", 
    "Black or African American", 
    "American Indian or Alaskan Native", 
    "Asian", 
    "Native Hawaiian or Pacific Islander",
    "Other race",
    # Age
    "More than median age of 36"
  )
)

# Persona configuration reactive element
per_reactive <- reactive({
  # Persona 1
  per_table$per1_value <- 0
  per_table$per1_value[per_table$levels == 'Intercept'] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_national_news] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_international_news] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_news_medium] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_follow_gov] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_travel_developing] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_political_pref] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_institution_trust] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_others_charitable] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_worship_attendance] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_religion_importance] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_religion] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_charity_importance] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_trust_in_charity] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_donation_frequency] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_donation_amount] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_volunteer_12] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_volunteer_frequency] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_personal_activism] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_family_activism] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_religious_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_sport_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_art_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_labor_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_political_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_environmental_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_professional_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_humanitarian_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_consumer_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_other_member] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_gender] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_marital_status] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_education] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_income] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_race] <- 1
  per_table$per1_value[per_table$levels %in% input$per1_age] <- 1

  # Persona 2
  per_table$per2_value <- 0
  per_table$per2_value[per_table$levels == 'Intercept'] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_national_news] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_international_news] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_news_medium] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_follow_gov] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_travel_developing] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_political_pref] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_institution_trust] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_others_charitable] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_worship_attendance] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_religion_importance] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_religion] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_charity_importance] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_trust_in_charity] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_donation_frequency] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_donation_amount] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_volunteer_12] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_volunteer_frequency] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_personal_activism] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_family_activism] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_religious_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_sport_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_art_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_labor_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_political_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_environmental_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_professional_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_humanitarian_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_consumer_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_other_member] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_gender] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_marital_status] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_education] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_income] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_race] <- 1
  per_table$per2_value[per_table$levels %in% input$per2_age] <- 1

  # Persona 3
  per_table$per3_value <- 0
  per_table$per3_value[per_table$levels == 'Intercept'] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_national_news] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_international_news] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_news_medium] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_follow_gov] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_travel_developing] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_political_pref] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_institution_trust] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_others_charitable] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_worship_attendance] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_religion_importance] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_religion] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_charity_importance] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_trust_in_charity] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_donation_frequency] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_donation_amount] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_volunteer_12] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_volunteer_frequency] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_personal_activism] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_family_activism] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_religious_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_sport_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_art_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_labor_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_political_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_environmental_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_professional_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_humanitarian_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_consumer_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_other_member] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_gender] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_marital_status] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_education] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_income] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_race] <- 1
  per_table$per3_value[per_table$levels %in% input$per3_age] <- 1
  per_table
})
```

Market share
--------------------------------------

```{r compute-and-plot}
# Compute and render the market share plot
renderPlotly({
  org_choice = org_reactive()
  per_choice = per_reactive()

  # Persona 1 predicted betas
  per1_betas <- as.matrix(gammas) %*% as.matrix(per_choice[,2])

  # Persona 1 expected utilities
  per1_org1_exp <- exp(t(org_choice[,2]) %*% as.matrix(per1_betas))
  per1_org2_exp <- exp(t(org_choice[,3]) %*% as.matrix(per1_betas))
  per1_org3_exp <- exp(t(org_choice[,4]) %*% as.matrix(per1_betas))
  per1_org4_exp <- exp(t(org_choice[,5]) %*% as.matrix(per1_betas))

  # Persona 1 market shares
  per1_share_denom <- sum(per1_org1_exp, per1_org2_exp, per1_org3_exp, per1_org4_exp)
  per1_org1_share <- 100 * (per1_org1_exp / per1_share_denom)
  per1_org2_share <- 100 * (per1_org2_exp / per1_share_denom)
  per1_org3_share <- 100 * (per1_org3_exp / per1_share_denom)
  per1_org4_share <- 100 * (per1_org4_exp / per1_share_denom)

  # Persona 2 predicted betas
  per2_betas <- as.matrix(gammas) %*% as.matrix(per_choice[,3])

  # Persona 2 expected utilities
  per2_org1_exp <- exp(t(org_choice[,2]) %*% as.matrix(per2_betas))
  per2_org2_exp <- exp(t(org_choice[,3]) %*% as.matrix(per2_betas))
  per2_org3_exp <- exp(t(org_choice[,4]) %*% as.matrix(per2_betas))
  per2_org4_exp <- exp(t(org_choice[,5]) %*% as.matrix(per2_betas))

  # Persona 2 market shares
  per2_share_denom <- sum(per2_org1_exp, per2_org2_exp, per2_org3_exp, per2_org4_exp)
  per2_org1_share <- 100 * (per2_org1_exp / per2_share_denom)
  per2_org2_share <- 100 * (per2_org2_exp / per2_share_denom)
  per2_org3_share <- 100 * (per2_org3_exp / per2_share_denom)
  per2_org4_share <- 100 * (per2_org4_exp / per2_share_denom)

  # Persona 3 predicted betas
  per3_betas <- as.matrix(gammas) %*% as.matrix(per_choice[,4])

  # Persona 3 expected utilities
  per3_org1_exp <- exp(t(org_choice[,2]) %*% as.matrix(per3_betas))
  per3_org2_exp <- exp(t(org_choice[,3]) %*% as.matrix(per3_betas))
  per3_org3_exp <- exp(t(org_choice[,4]) %*% as.matrix(per3_betas))
  per3_org4_exp <- exp(t(org_choice[,5]) %*% as.matrix(per3_betas))

  # Persona 3 market shares
  per3_share_denom <- sum(per3_org1_exp, per3_org2_exp, per3_org3_exp, per3_org4_exp)
  per3_org1_share <- 100 * (per3_org1_exp / per3_share_denom)
  per3_org2_share <- 100 * (per3_org2_exp / per3_share_denom)
  per3_org3_share <- 100 * (per3_org3_exp / per3_share_denom)
  per3_org4_share <- 100 * (per3_org4_exp / per3_share_denom)

  # Data for plotly
  plot_data <- data.frame(
    Value = c(
      # Persona 1
      per1_org1_share, per1_org2_share, per1_org3_share, per1_org4_share, 
      # Persona 2
      per2_org1_share, per2_org2_share, per2_org3_share, per2_org4_share,
      # Persona 3
      per3_org1_share, per3_org2_share, per3_org3_share, per3_org4_share
    ),
    Organization = c(input$org1_name, input$org2_name, input$org3_name, input$org4_name),
    Persona = c(
      rep(input$per1_name, 4), 
      rep(input$per2_name, 4), 
      rep(input$per3_name, 4)
    )
  )

  ggplotly(
    plot_data %>% 
      # Factor Organization and Persona with ordered=TRUE
      mutate(
        Organization = factor(
          Organization, 
          levels = c(input$org1_name, input$org2_name, input$org3_name, input$org4_name), 
          ordered = TRUE
        ),
        Persona = factor(
          Persona, 
          levels = c(input$per1_name, input$per2_name, input$per3_name), 
          ordered = TRUE
        )
      ) %>% 
      ggplot(aes(y = Value, x = Organization, fill = Persona)) +
      geom_bar(position = "dodge", stat = "identity") +
      labs(x="", y = "Market Share") + 
      scale_fill_manual(values = c("#FFC30B", "#28a153", "#c23838")) +
      theme(
        legend.position = "right",
        panel.background = element_blank(),
        panel.grid.major.y = element_line(color = "#D3D3D3")
      )
    )
})

# Compute and render the market share as a table
renderTable({
  org_choice = org_reactive()
  per_choice = per_reactive()

  # Persona 1 predicted betas
  per1_betas <- as.matrix(gammas) %*% as.matrix(per_choice[,2])

  # Persona 1 expected utilities
  per1_org1_exp <- exp(t(org_choice[,2]) %*% as.matrix(per1_betas))
  per1_org2_exp <- exp(t(org_choice[,3]) %*% as.matrix(per1_betas))
  per1_org3_exp <- exp(t(org_choice[,4]) %*% as.matrix(per1_betas))
  per1_org4_exp <- exp(t(org_choice[,5]) %*% as.matrix(per1_betas))

  # Persona 1 market shares
  per1_share_denom <- sum(per1_org1_exp, per1_org2_exp, per1_org3_exp, per1_org4_exp)
  per1_org1_share <- 100 * (per1_org1_exp / per1_share_denom)
  per1_org2_share <- 100 * (per1_org2_exp / per1_share_denom)
  per1_org3_share <- 100 * (per1_org3_exp / per1_share_denom)
  per1_org4_share <- 100 * (per1_org4_exp / per1_share_denom)

  # Persona 2 predicted betas
  per2_betas <- as.matrix(gammas) %*% as.matrix(per_choice[,3])

  # Persona 2 expected utilities
  per2_org1_exp <- exp(t(org_choice[,2]) %*% as.matrix(per2_betas))
  per2_org2_exp <- exp(t(org_choice[,3]) %*% as.matrix(per2_betas))
  per2_org3_exp <- exp(t(org_choice[,4]) %*% as.matrix(per2_betas))
  per2_org4_exp <- exp(t(org_choice[,5]) %*% as.matrix(per2_betas))

  # Persona 2 market shares
  per2_share_denom <- sum(per2_org1_exp, per2_org2_exp, per2_org3_exp, per2_org4_exp)
  per2_org1_share <- 100 * (per2_org1_exp / per2_share_denom)
  per2_org2_share <- 100 * (per2_org2_exp / per2_share_denom)
  per2_org3_share <- 100 * (per2_org3_exp / per2_share_denom)
  per2_org4_share <- 100 * (per2_org4_exp / per2_share_denom)

  # Persona 3 predicted betas
  per3_betas <- as.matrix(gammas) %*% as.matrix(per_choice[,4])

  # Persona 3 expected utilities
  per3_org1_exp <- exp(t(org_choice[,2]) %*% as.matrix(per3_betas))
  per3_org2_exp <- exp(t(org_choice[,3]) %*% as.matrix(per3_betas))
  per3_org3_exp <- exp(t(org_choice[,4]) %*% as.matrix(per3_betas))
  per3_org4_exp <- exp(t(org_choice[,5]) %*% as.matrix(per3_betas))

  # Persona 3 market shares
  per3_share_denom <- sum(per3_org1_exp, per3_org2_exp, per3_org3_exp, per3_org4_exp)
  per3_org1_share <- 100 * (per3_org1_exp / per3_share_denom)
  per3_org2_share <- 100 * (per3_org2_exp / per3_share_denom)
  per3_org3_share <- 100 * (per3_org3_exp / per3_share_denom)
  per3_org4_share <- 100 * (per3_org4_exp / per3_share_denom)

  # Table
  data <- data.frame(
    Shares = c(
      # Persona 1
      per1_org1_share, per1_org2_share, per1_org3_share, per1_org4_share, 
      # Persona 2
      per2_org1_share, per2_org2_share, per2_org3_share, per2_org4_share,
      # Persona 3
      per3_org1_share, per3_org2_share, per3_org3_share, per3_org4_share
    ),
    Organization = c(input$org1_name, input$org2_name, input$org3_name, input$org4_name),
    Persona = c(
      rep(input$per1_name, 4), 
      rep(input$per2_name, 4), 
      rep(input$per3_name, 4)
    )
  ) %>% 
    pivot_wider(names_from = Persona, values_from = Shares)

  data
})
```

Configurations {.tabset .tabset-fade}
--------------------------------------------------

### Organization 1

```{r org1-inputs}
# Naming
textInput(
  "org1_name",
  label = "Organization Name:",
  value = "Organization 1",
  placeholder = "Enter Organization Name",
  width = "450px"
)

# Organization
renderUI({
  org_data = organization_data()
  selectInput(
    "org1_org",
    label = "Organization:",
    choice = c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"),
    # selected = "Red Cross",
    selected = as.character(org_data[1:4,][which(org_data[1:4,2] == 1), 1]),
    width = "450px"
  )
})

# Issue area
renderUI({
  org_data = organization_data()
  selectInput(
    "org1_issue", 
    label = "Issue Area:",
    choices = c("Emergency response", "Environment", "Human rights", "Refugee relief"),
    # selected = "Emergency response",
    selected = ifelse(
      sum(org_data[5:7,2]) == 0, "Emergency response",
      as.character(org_data[5:7,][which(org_data[5:7,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial transparency
renderUI({
  org_data = organization_data()
  selectInput(
    "org1_fin_trans",
    label = "Financial Transparency:",
    choices = c(
      "Doesn’t engage in financial transparency", 
      "Engages in financial transparency"
    ),
    # selected = "Engages in financial transparency",
    selected = ifelse(
      sum(org_data[8,2]) == 0, "Doesn’t engage in financial transparency",
      as.character(org_data[8,][which(org_data[8,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial accountability
renderUI({
  org_data = organization_data()
  selectInput(
    "org1_fin_account",
    label = "Financial Accountability:",
    choices = c(
      "Doesn’t engage in financial accountability", 
      "Engages in financial accountability"
    ),
    # selected = "Engages in financial accountability",
    selected = ifelse(
      sum(org_data[9,2]) == 0, "Doesn’t engage in financial accountability",
      as.character(org_data[9,][which(org_data[9,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Funding sources
renderUI({
  org_data = organization_data()
  selectInput(
    "org1_funding",
    label = "Funding Source:",
    choices = c(
      "Many small private donations", 
      "Handful of wealthy private donors", 
      "Government grants"
    ),
    # selected = "Many small private donations",
    selected = ifelse(
      sum(org_data[10:11,2]) == 0, "Many small private donations",
      as.character(org_data[10:11,][which(org_data[10:11,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Relationship with host government
renderUI({
  org_data = organization_data()
  selectInput(
    "org1_host_gov",
    label = "Relationship With Host Government",
    choices = c(
      "Friendly relationship with government", 
      "Criticized by government",  
      "Under government crackdown"
    ),
    # selected = "Friendly relationship with government",
    selected = ifelse(
      sum(org_data[12:13,2]) == 0, "Friendly relationship with government",
      as.character(org_data[12:13,][which(org_data[12:13,2] == 1), 1])
    ),
    width = "450px"
  )
})
```

### Organization 2

```{r org2-inputs}
# Naming
textInput(
  "org2_name",
  label = "Organization Name:",
  value = "Organization 2",
  placeholder = "Enter Organization Name",
  width = "450px"
)

# Organization
renderUI({
  org_data = organization_data()
  selectInput(
    "org2_org",
    label = "Organization:",
    choice = c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"),
    # selected = "Red Cross",
    selected = as.character(org_data[1:4,][which(org_data[1:4,3] == 1), 1]),
    width = "450px"
  )
})

# Issue area
renderUI({
  org_data = organization_data()
  selectInput(
    "org2_issue", 
    label = "Issue Area:",
    choices = c("Emergency response", "Environment", "Human rights", "Refugee relief"),
    # selected = "Environment",
    selected = ifelse(
      sum(org_data[5:7,3]) == 0, "Emergency response",
      as.character(org_data[5:7,][which(org_data[5:7,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial transparency
renderUI({
  org_data = organization_data()
  selectInput(
    "org2_fin_trans",
    label = "Financial Transparency:",
    choices = c(
      "Doesn’t engage in financial transparency", 
      "Engages in financial transparency"
    ),
    # selected = "Engages in financial transparency",
    selected = ifelse(
      sum(org_data[8,3]) == 0, "Doesn’t engage in financial transparency",
      as.character(org_data[8,][which(org_data[8,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial accountability
renderUI({
  org_data = organization_data()
  selectInput(
    "org2_fin_account",
    label = "Financial Accountability:",
    choices = c(
      "Doesn’t engage in financial accountability", 
      "Engages in financial accountability"
    ),
    # selected = "Engages in financial accountability",
    selected = ifelse(
      sum(org_data[9,3]) == 0, "Doesn’t engage in financial accountability",
      as.character(org_data[9,][which(org_data[9,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Funding sources
renderUI({
  org_data = organization_data()
  selectInput(
    "org2_funding",
    label = "Funding Source:",
    choices = c(
      "Many small private donations", 
      "Handful of wealthy private donors", 
      "Government grants"
    ),
    # selected = "Many small private donations",
    selected = ifelse(
      sum(org_data[10:11,3]) == 0, "Many small private donations",
      as.character(org_data[10:11,][which(org_data[10:11,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Relationship with host government
renderUI({
  org_data = organization_data()
  selectInput(
    "org2_host_gov",
    label = "Relationship With Host Government",
    choices = c(
      "Friendly relationship with government", 
      "Criticized by government",  
      "Under government crackdown"
    ),
    # selected = "Friendly relationship with government",
    selected = ifelse(
      sum(org_data[12:13,3]) == 0, "Friendly relationship with government",
      as.character(org_data[12:13,][which(org_data[12:13,3] == 1), 1])
    ),
    width = "450px"
  )
})
```

### Organization 3

```{r org3-inputs}
# Naming
textInput(
  "org3_name",
  label = "Organization Name:",
  value = "Organization 3",
  placeholder = "Enter Organization Name",
  width = "450px"
)

# Organization
renderUI({
  org_data = organization_data()
  selectInput(
    "org3_org",
    label = "Organization:",
    choice = c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"),
    # selected = "Red Cross",
    selected = as.character(org_data[1:4,][which(org_data[1:4,4] == 1), 1]),
    width = "450px"
  )
})

# Issue area
renderUI({
  org_data = organization_data()
  selectInput(
    "org3_issue", 
    label = "Issue Area:",
    choices = c("Emergency response", "Environment", "Human rights", "Refugee relief"),
    # selected = "Human rights",
    selected = ifelse(
      sum(org_data[5:7,4]) == 0, "Emergency response",
      as.character(org_data[5:7,][which(org_data[5:7,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial transparency
renderUI({
  org_data = organization_data()
  selectInput(
    "org3_fin_trans",
    label = "Financial Transparency:",
    choices = c(
      "Doesn’t engage in financial transparency", 
      "Engages in financial transparency"
    ),
    # selected = "Engages in financial transparency",
    selected = ifelse(
      sum(org_data[8,4]) == 0, "Doesn’t engage in financial transparency",
      as.character(org_data[8,][which(org_data[8,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial accountability
renderUI({
  org_data = organization_data()
  selectInput(
    "org3_fin_account",
    label = "Financial Accountability:",
    choices = c(
      "Doesn’t engage in financial accountability", 
      "Engages in financial accountability"
    ),
    # selected = "Engages in financial accountability",
    selected = ifelse(
      sum(org_data[9,4]) == 0, "Doesn’t engage in financial accountability",
      as.character(org_data[9,][which(org_data[9,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Funding sources
renderUI({
  org_data = organization_data()
  selectInput(
    "org3_funding",
    label = "Funding Source:",
    choices = c(
      "Many small private donations", 
      "Handful of wealthy private donors", 
      "Government grants"
    ),
    # selected = "Many small private donations",
    selected = ifelse(
      sum(org_data[10:11,4]) == 0, "Many small private donations",
      as.character(org_data[10:11,][which(org_data[10:11,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Relationship with host government
renderUI({
  org_data = organization_data()
  selectInput(
    "org3_host_gov",
    label = "Relationship With Host Government",
    choices = c(
      "Friendly relationship with government", 
      "Criticized by government",  
      "Under government crackdown"
    ),
    # selected = "Friendly relationship with government",
    selected = ifelse(
      sum(org_data[12:13,4]) == 0, "Friendly relationship with government",
      as.character(org_data[12:13,][which(org_data[12:13,4] == 1), 1])
    ),
    width = "450px"
  )
})
```

### Organization 4

```{r org4-inputs}
# Naming
textInput(
  "org4_name",
  label = "Organization Name:",
  value = "Organization 4",
  placeholder = "Enter Organization Name",
  width = "450px"
)

# Organization
renderUI({
  org_data = organization_data()
  selectInput(
    "org4_org",
    label = "Organization:",
    choice = c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"),
    # selected = "Red Cross",
    selected = as.character(org_data[1:4,][which(org_data[1:4,5] == 1), 1]),
    width = "450px"
  )
})

# Issue area
renderUI({
  org_data = organization_data()
  selectInput(
    "org4_issue", 
    label = "Issue Area:",
    choices = c("Emergency response", "Environment", "Human rights", "Refugee relief"),
    # selected = "Refugee relief",
    selected = ifelse(
      sum(org_data[5:7,5]) == 0, "Emergency response",
      as.character(org_data[5:7,][which(org_data[5:7,5] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial transparency
renderUI({
  org_data = organization_data()
  selectInput(
    "org4_fin_trans",
    label = "Financial Transparency:",
    choices = c(
      "Doesn’t engage in financial transparency", 
      "Engages in financial transparency"
    ),
    # selected = "Engages in financial transparency",
    selected = ifelse(
      sum(org_data[8,5]) == 0, "Doesn’t engage in financial transparency",
      as.character(org_data[8,][which(org_data[8,5] == 1), 1])
    ),
    width = "450px"
  )
})

# Financial accountability
renderUI({
  org_data = organization_data()
  selectInput(
    "org4_fin_account",
    label = "Financial Accountability:",
    choices = c(
      "Doesn’t engage in financial accountability", 
      "Engages in financial accountability"
    ),
    # selected = "Engages in financial accountability",
    selected = ifelse(
      sum(org_data[9,5]) == 0, "Doesn’t engage in financial accountability",
      as.character(org_data[9,][which(org_data[9,5] == 1), 1])
    ),
    width = "450px"
  )
})

# Funding sources
renderUI({
  org_data = organization_data()
  selectInput(
    "org4_funding",
    label = "Funding Source:",
    choices = c(
      "Many small private donations", 
      "Handful of wealthy private donors", 
      "Government grants"
    ),
    # selected = "Many small private donations",
    selected = ifelse(
      sum(org_data[10:11,5]) == 0, "Many small private donations",
      as.character(org_data[10:11,][which(org_data[10:11,5] == 1), 1])
    ),
    width = "450px"
  )
})

# Relationship with host government
renderUI({
  org_data = organization_data()
  selectInput(
    "org4_host_gov",
    label = "Relationship With Host Government",
    choices = c(
      "Friendly relationship with government", 
      "Criticized by government",  
      "Under government crackdown"
    ),
    # selected = "Friendly relationship with government",
    selected = ifelse(
      sum(org_data[12:13,5]) == 0, "Friendly relationship with government",
      as.character(org_data[12:13,][which(org_data[12:13,5] == 1), 1])
    ),
    width = "450px"
  )
})
```

### Persona 1

```{r per1-inputs}
# Naming
textInput(
  "per1_name",
  label = "Persona Name:",
  value = "Persona 1",
  placeholder = "Enter Persona Name",
  width = "450px"
)

# Follow national news
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_national_news",
    label = "Follow national news:",
    choices = c("Rarely", "Once a week", "At least once a day"),
    # selected = "At least once a day",
    selected = ifelse(
      sum(per_data[2:3,2]) == 0, "Rarely",
      as.character(per_data[2:3,][which(per_data[2:3,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Follow international news
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_international_news",
    label = "Follow international news:",
    choices = c("Never", "Sometimes", "Always"),
    # selected = "Always",
    selected = ifelse(
      sum(per_data[4:5,2]) == 0, "Never",
      as.character(per_data[4:5,][which(per_data[4:5,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Mediums to follow news
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_news_medium",
    label = "Medium to follow news:",
    choices = c(
      "TV",
      "Print",
      "Online (not social media)",
      "Social media",
      "Radio",
      "Email newsletter",
      "News app"
    ),
    multiple = TRUE,
    # selected = c("TV", "Print", "Social media"),
    selected = deframe(per_data[6:12,][which(per_data[6:12,2] == 1), 1]),
    width = "450px"
  )
})

# Follow government and public affairs
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_follow_gov",
    label = "Follow government & public affairs:",
    choices = c("Not Often", "Often"),
    # selected = "Often",
    selected = ifelse(
      sum(per_data[13,2]) == 0, "Not Often",
      as.character(per_data[13,][which(per_data[13,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Traveled to a developing country
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_travel_developing",
    label = "Travel to developing country:",
    choices = c("Hasn't traveled to a developing country", "Traveled to a developing country"),
    # selected = "Traveled to a developing country",
    selected = ifelse(
      sum(per_data[14,2]) == 0, "Hasn't traveled to a developing country",
      as.character(per_data[14,][which(per_data[14,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Liberal to Conservative
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_political_pref",
    label = "Political views from liberal to conservative:",
    choices = c(
      "Extremely liberal",
      "Somewhat liberal",
      "Slightly liberal",
      "Moderate",
      "Slightly conservative",
      "Somewhat conservative",
      "Extremely conservative"
    ),
    # selected = "Somewhat liberal",
    selected = ifelse(
      sum(per_data[15:20,2]) == 0, "Extremely liberal",
      as.character(per_data[15:20,][which(per_data[15:20,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Trust in political institutions
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_institution_trust",
    label = "Political institution trust from none to complete:",
    choices = c(
      "No trust at all in political institutions",
      "Very little trust in political institutions",
      "Little trust in political institutions",
      "Neutral trust in political institutions",
      "Some trust in political institutions",
      "A lot of trust in political institutions",
      "Complete trust in political institutions"
    ),
    # selected = "Complete trust in political institutions",
    selected = ifelse(
      sum(per_data[21:26,2]) == 0, "No trust at all in political institutions",
      as.character(per_data[21:26,][which(per_data[21:26,2] == 1), 1])
    ),
    width = "450px"
  )
})

# People should be more charitable
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_others_charitable",
    label = "People should be more charitable:",
    choices = c(
      "Strongly disagree",
      "Disagree",
      "Somewhat disagree",
      "Neutral",
      "Somewhat agree",
      "Agree",
      "Strongly agree"
    ),
    # selected = "Strongly agree",
    selected = ifelse(
      sum(per_data[27:32,2]) == 0, "Strongly disagree",
      as.character(per_data[27:32,][which(per_data[27:32,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Religious attendance
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_worship_attendance",
    label = "Worship service attendance:",
    choices = c(
      "Not sure",
      "Rarely attend",
      "Attend at least once a month"
    ),
    # selected = "Attend at least once a month",
    selected = ifelse(
      sum(per_data[33:34,2]) == 0, "Not sure",
      as.character(per_data[33:34,][which(per_data[33:34,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Importance of religion
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_religion_importance",
    label = "Importance of religion:",
    choices = c("Not important", "Important"),
    # selected = "Important",
    selected = ifelse(
      sum(per_data[35,2]) == 0, "Not important",
      as.character(per_data[35,][which(per_data[35,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Religion
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_religion",
    label = "Religion:",
    choices = c(
      "Catholic",
      'Protestant',
      'Christian Orthodox',
      'Jewish',
      'Muslim',
      'Sikh',
      'Hindu',
      'Buddhist',
      'Atheist',
      'Other religion'
    ),
    # selected = 'Protestant',
     selected = ifelse(
      sum(per_data[36:44,2]) == 0, "Catholic",
      as.character(per_data[36:44,][which(per_data[36:44,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Importance of trusting charities
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_charity_importance",
    label = "Charity Importance:",
    choices = c(
      "Not at all important",
      "Very unimportant",
      "Somewhat unimportant",
      "Neutral importance",
      "Somewhat important",
      "Very important",
      "Essential"
    ),
    # selected = "Somewhat unimportant",
    selected = ifelse(
      sum(per_data[45:50,2]) == 0, "Not at all important",
      as.character(per_data[45:50,][which(per_data[45:50,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Trust in charities
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_trust_in_charity",
    label = "Charity Trust:",
    choices = c(
      "No trust at all in charities",
      "Very little trust in charities",
      "Little trust in charities",
      "Neutral trust in charities",
      "Some trust in charities",
      "A lot of trust in charities",
      "Complete trust in charities"
    ),
    # selected = "Complete trust in charities",
     selected = ifelse(
      sum(per_data[51:56,2]) == 0, "No trust at all in charities",
      as.character(per_data[51:56,][which(per_data[51:56,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Donation frequency
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_donation_frequency",
    label = "Donation Frequency:",
    choices = c('Donate more than once a month, less than once a year', 'Donate at least once a month'),
    # selected = "Donate at least once a month",
    selected = ifelse(
      sum(per_data[57,2]) == 0, "Donate more than once a month, less than once a year",
      as.character(per_data[57,][which(per_data[57,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Donated last year
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_donation_amount",
    label = "Donation Amount Last Year:",
    choices = c(
      '$1-$49',
      '$50-$99',
      '$100-$499',
      '$500-$999',
      '$1,000-$4,999',
      '$5,000-$9,999',
      '$10,000+'
    ),
    # selected = "$100-$499",
    selected = ifelse(
      sum(per_data[58:63,2]) == 0, "$1-$49",
      as.character(per_data[58:63,][which(per_data[58:63,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Volunteered in past 12 months
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_volunteer_12",
    label = "Volunteered in last 12 Months:",
    choices = c("Haven't volunteered in past 12 months", "Volunteered in past 12 months"),
    # selected = "Volunteered in past 12 months",
    selected = ifelse(
      sum(per_data[64,2]) == 0, "Haven't volunteered in past 12 months",
      as.character(per_data[64,][which(per_data[64,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Volunteer frequency
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_volunteer_frequency",
    label = "Volunteer Frequency:",
    choices = c(
      "Haven't volunteered in past 12 months",
      "Rarely volunteer",
      "Volunteer more than once a month, less than once a year",
      "Volunteer at least once a month"
    ),
    # selected = "Volunteer at least once a month",
    selected = ifelse(
      sum(per_data[65:67,2]) == 0, "Haven't volunteered in past 12 months",
      as.character(per_data[65:67,][which(per_data[65:67,2] == 1), 1])
    ),
    width = "450px"
  )
})

# History of activism
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_personal_activism",
    label = "Personal Activism:",
    choices = c('Not personally involved', 'Personally involved'),
    # selected = "Personally involved",
    selected = ifelse(
      sum(per_data[68,2]) == 0, "Not personally involved",
      as.character(per_data[68,][which(per_data[68,2] == 1), 1])
    ),
    width = "450px"
  )
})

# History of family activism
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_family_activism",
    label = "Family Activism:",
    choices = c('Family not involved', 'Family involved'),
    # selected = "Family involved",
    selected = ifelse(
      sum(per_data[69,2]) == 0, "Family not involved",
      as.character(per_data[69,][which(per_data[69,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Church or religious organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_religious_member",
    label = "Church Membership:",
    choices = c("Don't belong to church",'Inactive church member', 'Active church member'),
    # selected = "Don't belong to church",
    selected = ifelse(
      sum(per_data[70:71,2]) == 0, "Don't belong to church",
      as.character(per_data[70:71,][which(per_data[70:71,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Sport or recreational organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_sport_member",
    label = "Sport or Recreational Membership:",
    choices = c(
      "Don't belong to sports or rec organization", 
      'Inactive sports or rec organization member',
      'Active sports or rec organization member'
    ),
    # selected = "Don't belong to sports or rec organization",
    selected = ifelse(
      sum(per_data[72:73,2]) == 0, "Don't belong to sports or rec organization",
      as.character(per_data[72:73,][which(per_data[72:73,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Art, music, or educational organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_art_member",
    label = "Art, Music, or Educational Membership:",
    choices = c(
      "Don't belong to educational organization",
      'Inactive art, music, or educational organization member',
      'Active art, music, or educational organization member'
    ),
    # selected = "Don't belong to educational organization",
    selected = ifelse(
      sum(per_data[74:75,2]) == 0, "Don't belong to educational organization",
      as.character(per_data[74:75,][which(per_data[74:75,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Labor union
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_labor_member",
    label = "Labor Union Membership:",
    choices = c(
      "Don't belong to union",
      'Inactive union member', 
      'Active union member'
    ),
    # selected = "Don't belong to union",
    selected = ifelse(
      sum(per_data[76:77,2]) == 0, "Don't belong to union",
      as.character(per_data[76:77,][which(per_data[76:77,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Political party
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_political_member",
    label = "Political Party Membership:",
    choices = c(
      "Don't belong to political party",
      'Inactive political party member',
      'Active political party member'
    ),
    # selected = "Active political party member",
    selected = ifelse(
      sum(per_data[78:79,2]) == 0, "Don't belong to political party",
      as.character(per_data[78:79,][which(per_data[78:79,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Environmental organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_environmental_member",
    label = "Environmental Organization Membership:",
    choices = c(
      "Don't belong to environmental organization",
      'Inactive environmental organization member',
      'Active environmental organization member'
    ),
    # selected = "Don't belong to environmental organization",
    selected = ifelse(
      sum(per_data[80:81,2]) == 0, "Don't belong to environmental organization",
      as.character(per_data[80:81,][which(per_data[80:81,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Professional association
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_professional_member",
    label = "Professional Association Membership:",
    choices = c(
      "Don't belong to professional association",
      'Inactive professional association member',
      'Active professional association member'
    ),
    # selected = "Don't belong to professional association",
    selected = ifelse(
      sum(per_data[82:83,2]) == 0, "Don't belong to professional association",
      as.character(per_data[82:83,][which(per_data[82:83,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Humanitarian or charitable organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_humanitarian_member",
    label = "Humanitarian or Charitable Organization Membership:",
    choices = c(
      "Don't belong to humanitarian or charitable organization",
      'Inactive humanitarian or charitable organization member',
      'Active humanitarian or charitable organization member'
    ),
    # selected = "Active humanitarian or charitable organization member",
    selected = ifelse(
      sum(per_data[84:85,2]) == 0, "Don't belong to humanitarian or charitable organization",
      as.character(per_data[84:85,][which(per_data[84:85,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Consumer organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_consumer_member",
    label = "Consumer Organization Membership:",
    choices = c(
      "Don't belong to consumer organization",
      'Inactive consumer organization member',
      'Active consumer organization member'
    ),
    # selected = "Don't belong to consumer organization",
    selected = ifelse(
      sum(per_data[86:87,2]) == 0, "Don't belong to consumer organization",
      as.character(per_data[86:87,][which(per_data[86:87,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Other organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_other_member",
    label = "Other Organization Membership:",
    choices = c(
      "Don't belong to another organization",
      'Inactive member in another organization',
      'Active member in another organization'
    ),
    # selected = "Don't belong to another organization",
    selected = ifelse(
      sum(per_data[88:89,2]) == 0, "Don't belong to another organization",
      as.character(per_data[88:89,][which(per_data[88:89,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Gender
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_gender",
    label = "Gender:",
    choices = c(
      "Male",
      "Female",
      "Transgender",
      "Prefer not to say",
      "Other gender"
    ),
    # selected = "Male",
    selected = ifelse(
      sum(per_data[90:93,2]) == 0, "Male",
      as.character(per_data[90:93,][which(per_data[90:93,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Marital status
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_marital_status",
    label = "Marital Status:",
    choices = c(
      "Married",
      "Widowed",
      "Divorced",
      "Separated",
      "Never married"
    ),
    # selected = "Married",
    selected = ifelse(
      sum(per_data[94:97,2]) == 0, "Married",
      as.character(per_data[94:97,][which(per_data[94:97,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Education
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_education",
    label = "Education:",
    choices = c(
      "Less than high school",
      "High school graduate",
      "Some college",
      "2 year degree",
      "4 year degree",
      "Graduate or professional degree",
      "Doctorate"
    ),
    # selected = "2 year degree",
    selected = ifelse(
      sum(per_data[98:103,2]) == 0, "Less than high school",
      as.character(per_data[98:103,][which(per_data[98:103,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Income
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_income",
    label = "Income:",
    choices = c('Less than median of $61,372', 'More than median of $61,372'),
    # selected = "More than median of $61,372",
    selected = ifelse(
      sum(per_data[104,2]) == 0, "Less than median of $61,372",
      as.character(per_data[104,][which(per_data[104,2] == 1), 1])
    ),
    width = "450px"
  )
})

# Race
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_race",
    label = "Race:",
    choices = c(
      'White',
      'Black or African American',
      'American Indian or Alaskan Native',
      'Asian',
      'Native Hawaiian or Pacific Islander',
      'Other race'
    ),
    multiple = TRUE,
    # selected = "White",
    selected = deframe(per_data[105:110,][which(per_data[105:110,2] == 1), 1]),
    width = "450px"
  )
})

# Age
renderUI({
  per_data = persona_data()
  selectInput(
    "per1_age",
    label = "Age:",
    choices = c('Less than median age of 36', "More than median age of 36"),
    # selected = "Less than median age of 36",
    selected = ifelse(
      sum(per_data[111,2]) == 0, "Less than median age of 36",
      as.character(per_data[111,][which(per_data[111,2] == 1), 1])
    ),
    width = "450px"
  )
})
```

### Persona 2

```{r per2-inputs}
# Naming
textInput(
  "per2_name",
  label = "Persona Name:",
  value = "Persona 2",
  placeholder = "Enter Persona Name",
  width = "450px"
)

# Follow national news
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_national_news",
    label = "Follow national news:",
    choices = c("Rarely", "Once a week", "At least once a day"),
    # selected = "At least once a day",
    selected = ifelse(
      sum(per_data[2:3,3]) == 0, "Rarely",
      as.character(per_data[2:3,][which(per_data[2:3,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Follow international news
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_international_news",
    label = "Follow international news:",
    choices = c("Never", "Sometimes", "Always"),
    # selected = "Always",
    selected = ifelse(
      sum(per_data[4:5,3]) == 0, "Never",
      as.character(per_data[4:5,][which(per_data[4:5,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Mediums to follow news
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_news_medium",
    label = "Medium to follow news:",
    choices = c(
      "TV",
      "Print",
      "Online (not social media)",
      "Social media",
      "Radio",
      "Email newsletter",
      "News app"
    ),
    multiple = TRUE,
    # selected = c("TV", "Print", "Social media"),
    selected = deframe(per_data[6:12,][which(per_data[6:12,3] == 1), 1]),
    width = "450px"
  )
})

# Follow government and public affairs
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_follow_gov",
    label = "Follow government & public affairs:",
    choices = c("Not Often", "Often"),
    # selected = "Often",
    selected = ifelse(
      sum(per_data[13,3]) == 0, "Not Often",
      as.character(per_data[13,][which(per_data[13,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Traveled to a developing country
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_travel_developing",
    label = "Travel to developing country:",
    choices = c("Hasn't traveled to a developing country", "Traveled to a developing country"),
    # selected = "Traveled to a developing country",
    selected = ifelse(
      sum(per_data[14,3]) == 0, "Hasn't traveled to a developing country",
      as.character(per_data[14,][which(per_data[14,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Liberal to Conservative
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_political_pref",
    label = "Political views from liberal to conservative:",
    choices = c(
      "Extremely liberal",
      "Somewhat liberal",
      "Slightly liberal",
      "Moderate",
      "Slightly conservative",
      "Somewhat conservative",
      "Extremely conservative"
    ),
    # selected = "Somewhat liberal",
    selected = ifelse(
      sum(per_data[15:20,3]) == 0, "Extremely liberal",
      as.character(per_data[15:20,][which(per_data[15:20,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Trust in political institutions
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_institution_trust",
    label = "Political institution trust from none to complete:",
    choices = c(
      "No trust at all in political institutions",
      "Very little trust in political institutions",
      "Little trust in political institutions",
      "Neutral trust in political institutions",
      "Some trust in political institutions",
      "A lot of trust in political institutions",
      "Complete trust in political institutions"
    ),
    # selected = "Complete trust in political institutions",
    selected = ifelse(
      sum(per_data[21:26,3]) == 0, "No trust at all in political institutions",
      as.character(per_data[21:26,][which(per_data[21:26,3] == 1), 1])
    ),
    width = "450px"
  )
})

# People should be more charitable
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_others_charitable",
    label = "People should be more charitable:",
    choices = c(
      "Strongly disagree",
      "Disagree",
      "Somewhat disagree",
      "Neutral",
      "Somewhat agree",
      "Agree",
      "Strongly agree"
    ),
    # selected = "Strongly agree",
    selected = ifelse(
      sum(per_data[27:32,3]) == 0, "Strongly disagree",
      as.character(per_data[27:32,][which(per_data[27:32,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Religious attendance
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_worship_attendance",
    label = "Worship service attendance:",
    choices = c(
      "Not sure",
      "Rarely attend",
      "Attend at least once a month"
    ),
    # selected = "Attend at least once a month",
    selected = ifelse(
      sum(per_data[33:34,3]) == 0, "Not sure",
      as.character(per_data[33:34,][which(per_data[33:34,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Importance of religion
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_religion_importance",
    label = "Importance of religion:",
    choices = c("Not important", "Important"),
    # selected = "Important",
    selected = ifelse(
      sum(per_data[35,3]) == 0, "Not important",
      as.character(per_data[35,][which(per_data[35,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Religion
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_religion",
    label = "Religion:",
    choices = c(
      'Catholic',
      'Protestant',
      'Christian Orthodox',
      'Jewish',
      'Muslim',
      'Sikh',
      'Hindu',
      'Buddhist',
      'Atheist',
      'Other religion'
    ),
    # selected = 'Protestant',
    selected = ifelse(
      sum(per_data[36:44,3]) == 0, "Catholic",
      as.character(per_data[36:44,][which(per_data[36:44,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Importance of trusting charities
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_charity_importance",
    label = "Charity Importance:",
    choices = c(
      "Not at all important",
      "Very unimportant",
      "Somewhat unimportant",
      "Neutral importance",
      "Somewhat important",
      "Very important",
      "Essential"
    ),
    # selected = "Somewhat unimportant",
    selected = ifelse(
      sum(per_data[45:50,3]) == 0, "Not at all important",
      as.character(per_data[45:50,][which(per_data[45:50,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Trust in charities
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_trust_in_charity",
    label = "Charity Trust:",
    choices = c(
      "No trust at all in charities",
      "Very little trust in charities",
      "Little trust in charities",
      "Neutral trust in charities",
      "Some trust in charities",
      "A lot of trust in charities",
      "Complete trust in charities"
    ),
    # selected = "Complete trust in charities",
    selected = ifelse(
      sum(per_data[51:56,3]) == 0, "No trust at all in charities",
      as.character(per_data[51:56,][which(per_data[51:56,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Donation frequency
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_donation_frequency",
    label = "Donation Frequency:",
    choices = c('Donate more than once a month, less than once a year', 'Donate at least once a month'),
    # selected = "Donate at least once a month",
    selected = ifelse(
      sum(per_data[57,3]) == 0, "Donate more than once a month, less than once a year",
      as.character(per_data[57,][which(per_data[57,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Donated last year
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_donation_amount",
    label = "Donation Amount Last Year:",
    choices = c(
      '$1-$49',
      '$50-$99',
      '$100-$499',
      '$500-$999',
      '$1,000-$4,999',
      '$5,000-$9,999',
      '$10,000+'
    ),
    # selected = "$100-$499",
    selected = ifelse(
      sum(per_data[58:63,3]) == 0, "$1-$49",
      as.character(per_data[58:63,][which(per_data[58:63,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Volunteered in past 12 months
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_volunteer_12",
    label = "Volunteered in last 12 Months:",
    choices = c("Haven't volunteered in past 12 months", "Volunteered in past 12 months"),
    # selected = "Volunteered in past 12 months",
    selected = ifelse(
      sum(per_data[64,3]) == 0, "Haven't volunteered in past 12 months",
      as.character(per_data[64,][which(per_data[64,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Volunteer frequency
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_volunteer_frequency",
    label = "Volunteer Frequency:",
    choices = c(
      "Haven't volunteered in past 12 months",
      "Rarely volunteer",
      "Volunteer more than once a month, less than once a year",
      "Volunteer at least once a month"
    ),
    # selected = "Volunteer at least once a month",
    selected = ifelse(
      sum(per_data[65:67,3]) == 0, "Haven't volunteered in past 12 months",
      as.character(per_data[65:67,][which(per_data[65:67,3] == 1), 1])
    ),
    width = "450px"
  )
})

# History of activism
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_personal_activism",
    label = "Personal Activism:",
    choices = c('Not personally involved', 'Personally involved'),
    # selected = "Personally involved",
    selected = ifelse(
      sum(per_data[68,3]) == 0, "Not personally involved",
      as.character(per_data[68,][which(per_data[68,3] == 1), 1])
    ),
    width = "450px"
  )
})

# History of family activism
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_family_activism",
    label = "Family Activism:",
    choices = c('Family not involved', 'Family involved'),
    # selected = "Family involved",
    selected = ifelse(
      sum(per_data[69,3]) == 0, "Family not involved",
      as.character(per_data[69,][which(per_data[69,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Church or religious organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_religious_member",
    label = "Church Membership:",
    choices = c("Don't belong to church",'Inactive church member', 'Active church member'),
    # selected = "Don't belong to church",
    selected = ifelse(
      sum(per_data[70:71,3]) == 0, "Don't belong to church",
      as.character(per_data[70:71,][which(per_data[70:71,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Sport or recreational organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_sport_member",
    label = "Sport or Recreational Membership:",
    choices = c(
      "Don't belong to sports or rec organization", 
      'Inactive sports or rec organization member',
      'Active sports or rec organization member'
    ),
    # selected = "Don't belong to sports or rec organization",
    selected = ifelse(
      sum(per_data[72:73,3]) == 0, "Don't belong to sports or rec organization",
      as.character(per_data[72:73,][which(per_data[72:73,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Art, music, or educational organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_art_member",
    label = "Art, Music, or Educational Membership:",
    choices = c(
      "Don't belong to educational organization",
      'Inactive art, music, or educational organization member',
      'Active art, music, or educational organization member'
    ),
    # selected = "Don't belong to educational organization",
    selected = ifelse(
      sum(per_data[74:75,3]) == 0, "Don't belong to educational organization",
      as.character(per_data[74:75,][which(per_data[74:75,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Labor union
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_labor_member",
    label = "Labor Union Membership:",
    choices = c("Don't belong to union", 'Inactive union member', 'Active union member'),
    # selected = "Don't belong to union",
    selected = ifelse(
      sum(per_data[76:77,3]) == 0, "Don't belong to union",
      as.character(per_data[76:77,][which(per_data[76:77,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Political party
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_political_member",
    label = "Political Party Membership:",
    choices = c(
      "Don't belong to political party",
      'Inactive political party member',
      'Active political party member'
    ),
    # selected = "Active political party member",
    selected = ifelse(
      sum(per_data[78:79,3]) == 0, "Don't belong to political party",
      as.character(per_data[78:79,][which(per_data[78:79,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Environmental organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_environmental_member",
    label = "Environmental Organization Membership:",
    choices = c(
      "Don't belong to environmental organization",
      'Inactive environmental organization member',
      'Active environmental organization member'
    ),
    # selected = "Don't belong to environmental organization",
    selected = ifelse(
      sum(per_data[80:81,3]) == 0, "Don't belong to environmental organization",
      as.character(per_data[80:81,][which(per_data[80:81,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Professional association
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_professional_member",
    label = "Professional Association Membership:",
    choices = c(
      "Don't belong to professional association",
      'Inactive professional association member',
      'Active professional association member'
    ),
    # selected = "Don't belong to professional association",
    selected = ifelse(
      sum(per_data[82:83,3]) == 0, "Don't belong to professional association",
      as.character(per_data[82:83,][which(per_data[82:83,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Humanitarian or charitable organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_humanitarian_member",
    label = "Humanitarian or Charitable Organization Membership:",
    choices = c(
      "Don't belong to humanitarian or charitable organization",
      'Inactive humanitarian or charitable organization member',
      'Active humanitarian or charitable organization member'
    ),
    # selected = "Active humanitarian or charitable organization member",
    selected = ifelse(
      sum(per_data[84:85,3]) == 0, "Don't belong to humanitarian or charitable organization",
      as.character(per_data[84:85,][which(per_data[84:85,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Consumer organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_consumer_member",
    label = "Consumer Organization Membership:",
    choices = c(
      "Don't belong to consumer organization",
      'Inactive consumer organization member',
      'Active consumer organization member'
    ),
    # selected = "Don't belong to consumer organization",
    selected = ifelse(
      sum(per_data[86:87,3]) == 0, "Don't belong to consumer organization",
      as.character(per_data[86:87,][which(per_data[86:87,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Other organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_other_member",
    label = "Other Organization Membership:",
    choices = c(
      "Don't belong to another organization",
      'Inactive member in another organization',
      'Active member in another organization'
    ),
    # selected = "Don't belong to another organization",
    selected = ifelse(
      sum(per_data[88:89,3]) == 0, "Don't belong to another organization",
      as.character(per_data[88:89,][which(per_data[88:89,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Gender
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_gender",
    label = "Gender:",
    choices = c(
      "Male",
      "Female",
      "Transgender",
      "Prefer not to say",
      "Other gender"
    ),
    # selected = "Male",
    selected = ifelse(
      sum(per_data[90:93,3]) == 0, "Male",
      as.character(per_data[90:93,][which(per_data[90:93,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Marital status
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_marital_status",
    label = "Marital Status:",
    choices = c(
      "Married",
      "Widowed",
      "Divorced",
      "Separated",
      "Never married"
    ),
    # selected = "Married",
    selected = ifelse(
      sum(per_data[94:97,3]) == 0, "Married",
      as.character(per_data[94:97,][which(per_data[94:97,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Education
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_education",
    label = "Education:",
    choices = c(
      "Less than high school",
      "High school graduate",
      "Some college",
      "2 year degree",
      "4 year degree",
      "Graduate or professional degree",
      "Doctorate"
    ),
    # selected = "Some college",
    selected = ifelse(
      sum(per_data[98:103,3]) == 0, "Less than high school",
      as.character(per_data[98:103,][which(per_data[98:103,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Income
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_income",
    label = "Income:",
    choices = c('Less than median of $61,372', 'More than median of $61,372'),
    # selected = "More than median of $61,372",
    selected = ifelse(
      sum(per_data[104,3]) == 0, "Less than median of $61,372",
      as.character(per_data[104,][which(per_data[104,3] == 1), 1])
    ),
    width = "450px"
  )
})

# Race
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_race",
    label = "Race:",
    choices = c(
      'White',
      'Black or African American',
      'American Indian or Alaskan Native',
      'Asian',
      'Native Hawaiian or Pacific Islander',
      'Other race'
    ),
    multiple = TRUE,
    # selected = "American Indian or Alaskan Native",
    selected = deframe(per_data[105:110,][which(per_data[105:110,3] == 1), 1]),
    width = "450px"
  )
})

# Age
renderUI({
  per_data = persona_data()
  selectInput(
    "per2_age",
    label = "Age:",
    choices = c('Less than median age of 36', "More than median age of 36"),
    # selected = "Less than median age of 36",
    selected = ifelse(
      sum(per_data[111,3]) == 0, "Less than median age of 36",
      as.character(per_data[111,][which(per_data[111,3] == 1), 1])
    ),
    width = "450px"
  )
})
```

### Persona 3

```{r per3-inputs}
# Naming
textInput(
  "per3_name",
  label = "Persona Name:",
  value = "Persona 3",
  placeholder = "Enter Persona Name",
  width = "450px"
)

# Follow national news
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_national_news",
    label = "Follow national news:",
    choices = c("Rarely", "Once a week", "At least once a day"),
    # selected = "At least once a day",
    selected = ifelse(
      sum(per_data[2:3,4]) == 0, "Rarely",
      as.character(per_data[2:3,][which(per_data[2:3,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Follow international news
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_international_news",
    label = "Follow international news:",
    choices = c("Never", "Sometimes", "Always"),
    # selected = "Always",
    selected = ifelse(
      sum(per_data[4:5,4]) == 0, "Never",
      as.character(per_data[4:5,][which(per_data[4:5,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Mediums to follow news
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_news_medium",
    label = "Medium to follow news:",
    choices = c(
      "TV",
      "Print",
      "Online (not social media)",
      "Social media",
      "Radio",
      "Email newsletter",
      "News app"
    ),
    multiple = TRUE,
    # selected = c("TV", "Print", "Social media"),
    selected = deframe(per_data[6:12,][which(per_data[6:12,4] == 1), 1]),
    width = "450px"
  )
})

# Follow government and public affairs
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_follow_gov",
    label = "Follow government & public affairs:",
    choices = c("Not Often", "Often"),
    # selected = "Often",
    selected = ifelse(
      sum(per_data[13,4]) == 0, "Not Often",
      as.character(per_data[13,][which(per_data[13,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Traveled to a developing country
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_travel_developing",
    label = "Travel to developing country:",
    choices = c("Hasn't traveled to a developing country", "Traveled to a developing country"),
    # selected = "Traveled to a developing country",
    selected = ifelse(
      sum(per_data[14,4]) == 0, "Hasn't traveled to a developing country",
      as.character(per_data[14,][which(per_data[14,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Liberal to Conservative
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_political_pref",
    label = "Political views from liberal to conservative:",
    choices = c(
      "Extremely liberal",
      "Somewhat liberal",
      "Slightly liberal",
      "Moderate",
      "Slightly conservative",
      "Somewhat conservative",
      "Extremely conservative"
    ),
    # selected = "Somewhat liberal",
    selected = ifelse(
      sum(per_data[15:20,4]) == 0, "Extremely liberal",
      as.character(per_data[15:20,][which(per_data[15:20,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Trust in political institutions
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_institution_trust",
    label = "Political institution trust from none to complete:",
    choices = c(
      "No trust at all in political institutions",
      "Very little trust in political institutions",
      "Little trust in political institutions",
      "Neutral trust in political institutions",
      "Some trust in political institutions",
      "A lot of trust in political institutions",
      "Complete trust in political institutions"
    ),
    # selected = "Complete trust in political institutions",
    selected = ifelse(
      sum(per_data[21:26,4]) == 0, "No trust at all in political institutions",
      as.character(per_data[21:26,][which(per_data[21:26,4] == 1), 1])
    ),
    width = "450px"
  )
})

# People should be more charitable
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_others_charitable",
    label = "People should be more charitable:",
    choices = c(
      "Strongly disagree",
      "Disagree",
      "Somewhat disagree",
      "Neutral",
      "Somewhat agree",
      "Agree",
      "Strongly agree"
    ),
    # selected = "Strongly agree",
    selected = ifelse(
      sum(per_data[27:32,4]) == 0, "Strongly disagree",
      as.character(per_data[27:32,][which(per_data[27:32,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Religious attendance
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_worship_attendance",
    label = "Worship service attendance:",
    choices = c(
      "Not sure",
      "Rarely attend",
      "Attend at least once a month"
    ),
    # selected = "Attend at least once a month",
    selected = ifelse(
      sum(per_data[33:34,4]) == 0, "Not sure",
      as.character(per_data[33:34,][which(per_data[33:34,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Importance of religion
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_religion_importance",
    label = "Importance of religion:",
    choices = c("Not important", "Important"),
    # selected = "Important",
    selected = ifelse(
      sum(per_data[35,4]) == 0, "Not important",
      as.character(per_data[35,][which(per_data[35,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Religion
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_religion",
    label = "Religion:",
    choices = c(
      'Catholic',
      'Protestant',
      'Christian Orthodox',
      'Jewish',
      'Muslim',
      'Sikh',
      'Hindu',
      'Buddhist',
      'Atheist',
      'Other religion'
    ),
    # selected = 'Protestant',
    selected = ifelse(
      sum(per_data[36:44,4]) == 0, "Catholic",
      as.character(per_data[36:44,][which(per_data[36:44,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Importance of trusting charities
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_charity_importance",
    label = "Charity Importance:",
    choices = c(
      "Not at all important",
      "Very unimportant",
      "Somewhat unimportant",
      "Neutral importance",
      "Somewhat important",
      "Very important",
      "Essential"
    ),
    # selected = "Somewhat unimportant",
    selected = ifelse(
      sum(per_data[45:50,4]) == 0, "Not at all important",
      as.character(per_data[45:50,][which(per_data[45:50,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Trust in charities
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_trust_in_charity",
    label = "Charity Trust:",
    choices = c(
      "No trust at all in charities",
      "Very little trust in charities",
      "Little trust in charities",
      "Neutral trust in charities",
      "Some trust in charities",
      "A lot of trust in charities",
      "Complete trust in charities"
    ),
    # selected = "Complete trust in charities",
    selected = ifelse(
      sum(per_data[51:56,4]) == 0, "No trust at all in charities",
      as.character(per_data[51:56,][which(per_data[51:56,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Donation frequency
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_donation_frequency",
    label = "Donation Frequency:",
    choices = c('Donate more than once a month, less than once a year', 'Donate at least once a month'),
    # selected = "Donate at least once a month",
    selected = ifelse(
      sum(per_data[57,4]) == 0, "Donate more than once a month, less than once a year",
      as.character(per_data[57,][which(per_data[57,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Donated last year
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_donation_amount",
    label = "Donation Amount Last Year:",
    choices = c(
      '$1-$49',
      '$50-$99',
      '$100-$499',
      '$500-$999',
      '$1,000-$4,999',
      '$5,000-$9,999',
      '$10,000+'
    ),
    # selected = "$100-$499",
    selected = ifelse(
      sum(per_data[58:63,4]) == 0, "$1-$49",
      as.character(per_data[58:63,][which(per_data[58:63,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Volunteered in past 12 months
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_volunteer_12",
    label = "Volunteered in last 12 Months:",
    choices = c("Haven't volunteered in past 12 months", "Volunteered in past 12 months"),
    # selected = "Volunteered in past 12 months",
    selected = ifelse(
      sum(per_data[64,4]) == 0, "Haven't volunteered in past 12 months",
      as.character(per_data[64,][which(per_data[64,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Volunteer frequency
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_volunteer_frequency",
    label = "Volunteer Frequency:",
    choices = c(
      "Haven't volunteered in past 12 months",
      "Rarely volunteer",
      "Volunteer more than once a month, less than once a year",
      "Volunteer at least once a month"
    ),
    # selected = "Volunteer at least once a month",
    selected = ifelse(
      sum(per_data[65:67,4]) == 0, "Haven't volunteered in past 12 months",
      as.character(per_data[65:67,][which(per_data[65:67,4] == 1), 1])
    ),
    width = "450px"
  )
})

# History of activism
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_personal_activism",
    label = "Personal Activism:",
    choices = c('Not personally involved', 'Personally involved'),
    # selected = "Personally involved",
    selected = ifelse(
      sum(per_data[68,4]) == 0, "Not personally involved",
      as.character(per_data[68,][which(per_data[68,4] == 1), 1])
    ),
    width = "450px"
  )
})

# History of family activism
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_family_activism",
    label = "Family Activism:",
    choices = c('Family not involved', 'Family involved'),
    # selected = "Family involved",
    selected = ifelse(
      sum(per_data[69,4]) == 0, "Family not involved",
      as.character(per_data[69,][which(per_data[69,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Church or religious organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_religious_member",
    label = "Church Membership:",
    choices = c("Don't belong to church",'Inactive church member', 'Active church member'),
    # selected = "Don't belong to church",
    selected = ifelse(
      sum(per_data[70:71,4]) == 0, "Don't belong to church",
      as.character(per_data[70:71,][which(per_data[70:71,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Sport or recreational organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_sport_member",
    label = "Sport or Recreational Membership:",
    choices = c(
      "Don't belong to sports or rec organization", 
      'Inactive sports or rec organization member',
      'Active sports or rec organization member'
    ),
    # selected = "Don't belong to sports or rec organization",
    selected = ifelse(
      sum(per_data[72:73,4]) == 0, "Don't belong to sports or rec organization",
      as.character(per_data[72:73,][which(per_data[72:73,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Art, music, or educational organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_art_member",
    label = "Art, Music, or Educational Membership:",
    choices = c(
      "Don't belong to educational organization",
      'Inactive art, music, or educational organization member',
      'Active art, music, or educational organization member'
    ),
    # selected = "Don't belong to educational organization",
    selected = ifelse(
      sum(per_data[74:75,4]) == 0, "Don't belong to educational organization",
      as.character(per_data[74:75,][which(per_data[74:75,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Labor union
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_labor_member",
    label = "Labor Union Membership:",
    choices = c("Don't belong to union", 'Inactive union member', 'Active union member'),
    # selected = "Don't belong to union",
    selected = ifelse(
      sum(per_data[76:77,4]) == 0, "Don't belong to union",
      as.character(per_data[76:77,][which(per_data[76:77,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Political party
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_political_member",
    label = "Political Party Membership:",
    choices = c(
      "Don't belong to political party",
      'Inactive political party member',
      'Active political party member'
    ),
    # selected = "Active political party member",
    selected = ifelse(
      sum(per_data[78:79,4]) == 0, "Don't belong to political party",
      as.character(per_data[78:79,][which(per_data[78:79,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Environmental organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_environmental_member",
    label = "Environmental Organization Membership:",
    choices = c(
      "Don't belong to environmental organization",
      'Inactive environmental organization member',
      'Active environmental organization member'
    ),
    # selected = "Don't belong to environmental organization",
    selected = ifelse(
      sum(per_data[80:81,4]) == 0, "Don't belong to environmental organization",
      as.character(per_data[80:81,][which(per_data[80:81,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Professional association
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_professional_member",
    label = "Professional Association Membership:",
    choices = c(
      "Don't belong to professional association",
      'Inactive professional association member',
      'Active professional association member'
    ),
    # selected = "Don't belong to professional association",
    selected = ifelse(
      sum(per_data[82:83,4]) == 0, "Don't belong to professional association",
      as.character(per_data[82:83,][which(per_data[82:83,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Humanitarian or charitable organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_humanitarian_member",
    label = "Humanitarian or Charitable Organization Membership:",
    choices = c(
      "Don't belong to humanitarian or charitable organization",
      'Inactive humanitarian or charitable organization member',
      'Active humanitarian or charitable organization member'
    ),
    # selected = "Active humanitarian or charitable organization member",
    selected = ifelse(
      sum(per_data[84:85,4]) == 0, "Don't belong to humanitarian or charitable organization",
      as.character(per_data[84:85,][which(per_data[84:85,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Consumer organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_consumer_member",
    label = "Consumer Organization Membership:",
    choices = c(
      "Don't belong to consumer organization",
      'Inactive consumer organization member',
      'Active consumer organization member'
    ),
    # selected = "Don't belong to consumer organization",
    selected = ifelse(
      sum(per_data[86:87,4]) == 0, "Don't belong to consumer organization",
      as.character(per_data[86:87,][which(per_data[86:87,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Other organization
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_other_member",
    label = "Other Organization Membership:",
    choices = c(
      "Don't belong to another organization",
      'Inactive member in another organization',
      'Active member in another organization'
    ),
    # selected = "Don't belong to another organization",
    selected = ifelse(
      sum(per_data[88:89,4]) == 0, "Don't belong to another organization",
      as.character(per_data[88:89,][which(per_data[88:89,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Gender
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_gender",
    label = "Gender:",
    choices = c(
      "Male",
      "Female",
      "Transgender",
      "Prefer not to say",
      "Other gender"
    ),
    # selected = "Female",
    selected = ifelse(
      sum(per_data[90:93,4]) == 0, "Male",
      as.character(per_data[90:93,][which(per_data[90:93,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Marital status
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_marital_status",
    label = "Marital Status:",
    choices = c(
      "Married",
      "Widowed",
      "Divorced",
      "Separated",
      "Never married"
    ),
    # selected = "Married",
    selected = ifelse(
      sum(per_data[94:97,4]) == 0, "Married",
      as.character(per_data[94:97,][which(per_data[94:97,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Education
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_education",
    label = "Education:",
    choices = c(
      "Less than high school",
      "High school graduate",
      "Some college",
      "2 year degree",
      "4 year degree",
      "Graduate or professional degree",
      "Doctorate"
    ),
    # selected = "Graduate or professional degree",
    selected = ifelse(
      sum(per_data[98:103,4]) == 0, "Less than high school",
      as.character(per_data[98:103,][which(per_data[98:103,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Income
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_income",
    label = "Income:",
    choices = c('Less than median of $61,372', 'More than median of $61,372'),
    # selected = "More than median of $61,372",
    selected = ifelse(
      sum(per_data[104,4]) == 0, "Less than median of $61,372",
      as.character(per_data[104,][which(per_data[104,4] == 1), 1])
    ),
    width = "450px"
  )
})

# Race
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_race",
    label = "Race:",
    choices = c(
      'White',
      'Black or African American',
      'American Indian or Alaskan Native',
      'Asian',
      'Native Hawaiian or Pacific Islander',
      'Other race'
    ),
    multiple = TRUE,
    # selected = c("Black or African American", "American Indian or Alaskan Native"),
    selected = deframe(per_data[105:110,][which(per_data[105:110,4] == 1), 1]),
    width = "450px"
  )
})

# Age
renderUI({
  per_data = persona_data()
  selectInput(
    "per3_age",
    label = "Age:",
    choices = c('Less than median age of 36', "More than median age of 36"),
    # selected = "Less than median age of 36",
    selected = ifelse(
      sum(per_data[111,4]) == 0, "Less than median age of 36",
      as.character(per_data[111,][which(per_data[111,4] == 1), 1])
    ),
    width = "450px"
  )
})
```

### Export

```{r export}
# Render the organization configurations as a table
renderTable({
  org_choice = org_reactive()
  org_choice
})

# Download Button
downloadButton("download_org", label = "Download Organization Data")

downloadHandler(
  filename = function() {
    paste0("org_data_",Sys.time(),".csv")
  },
  content = function(file) {
    write_csv(org_reactive(), file)
  }
)

# Render the persona configurations as a table
renderTable({
  per_choice = per_reactive()
  per_choice
})

# Download Button
downloadButton("download_per", label = "Download Persona Data")

downloadHandler(
  filename = function() {
    paste0("per_data_",Sys.time(),".csv")
  },
  content = function(file) {
    write_csv(per_reactive(), file)
  }
)
```

### Imports

```{r import}
# Organization file input
fileInput(
  "org_update",
  "Import Organization Data",
  accept = c("text/csv",
             "text/comma-separated-values,text/plain",
             ".csv")
)

# actionButton(
#   "org_button",
#   label = "Update Organization Data"
# )

# Persona file input
fileInput(
  "per_update",
  "Import Persona Data",
  accept = c("text/csv",
             "text/comma-separated-values,text/plain",
             ".csv")
)

# Render the organization configurations as a table
renderTable({
  org_import = read_csv(input$org_update$datapath)

  org_import
})

# Render the persona configurations as a table
renderTable({
  per_import = read_csv(input$per_update$datapath)

  per_import
})
```