### 2020-11-29 analysis/02_clean-data.Rmd
Modified the labeling in 02_clean-data (again) so that it all lives there and no additional clean-up is needed in 04_figs-tables

---
title: "Clean data"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Check data

## Load data

We downloaded the `conjointqsf_final.csv` file directly from Qualtrics and ran this code to remove identifying information. We then saved that CSV file as `conjointqsf_final.csv` and include it in `data/raw_data` for distribution. We use the de-identified file in the rest of our data cleaning and analysis.

```r
library(tidyverse)
library(donorheuristics)
library(here)

col_names <- names(read_csv(here::here("data", "private_data", "conjointqsf_final.csv")))
data <- read_csv(
  here::here("data", "private_data", "conjointqsf_final.csv"),
  skip = 3,
  col_names = col_names,
  guess_max = 1500
)

# Remove IP address and location information
data %>% 
  select(-V6, -LocationLatitude, -LocationLongitude, -LocationAccuracy) %>% 
  write_csv(here::here("data", "raw_data", "conjointqsf_final.csv"))
```

```{r load-raw-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(here)

data <- read_csv(
  here::here("data", "raw_data", "conjointqsf_final.csv"),
  guess_max = 1500
)
```

## Respondents who finished

```{r number-finished}
data %>%
  summarize(
    total = n(),
    finished = sum(V10)
  )
```

Out of 1508 respondents, 1437 "finished" the survey and 71 didn't.

## Consent

```{r consent}
data %>%
  count(Q1.1)

data %>%
  count(Q1.1, Q2.1)
```

30 respondents didn't give consent while 3 respondents gave consent and didn't respond starting in Q2.1.

## Screening

```{r screening}
data %>%
  count(Q2.5)

data %>%
  count(Q2.5, Q2.6) %>% 
  filter(is.na(Q2.6) == TRUE)
```

An additional 12 respondents drop out by Q2.5 (a total of 45). 385 respondents are screened at Q2.5. 2 more respondents drop off by Q2.6 (a total of 47).

## Check display logic

```{r display-logic}
data %>%
  count(Q2.9)

data %>%
  count(Q2.9, Q2.10)
```

8 more respondents have dropped out (a total of 55). The 445 people who don't volunteer in the past 12 months properly skip Q2.10.

## Attention check

```{r attention-check}
data %>%
  count(Q2.11)
```

6 respondents don't make the attention check.

## Confirm ages

```{r confirm-ages}
data %>%
  count(Q5.17)
```

By the end of the survey, 44 more respondents have dropped out (a total of 99) but no one is under 18, as they indicated as part of the consent.

# Clean data

## Clean based on initial criteria

```{r clean-data}
data_clean <- data %>%
  filter(V10 == 1) %>%         # Keep respondents who completed the entire survey
  filter(Q1.1 == 1) %>%        # Keep respondents who gave consent
  filter(Q2.5 %in% c(1:5)) %>% # Keep respondents who passed the screener
  filter(Q2.11 == 3)           # Keep respondents who passed the attention check
```

## Confirm the data is clean

```{r clean-data-confirmation}
# Finished respondents
data_clean %>% 
  summarize(
    total = n(),
    finished = sum(V10)
  )

# Consent
data_clean %>%
  count(Q1.1)

# Screening
data_clean %>%
  count(Q2.5)

# Attention check
data_clean %>% 
  count(Q2.11)
```

We have 1016 total respondents who have finished the entire survey, given consent, and passed the screening and attention check.

## Confirm the choice data is clean

```{r choice-clean}
data_clean %>%
  count(vers_CBCONJOINT) %>%
  count(n)
```

The design version appears to be properly assigned at random.

```{r design-random-assignment}
data_clean %>%
  select(Q4.1:Q4.12) %>%
  gather(key = "task", value = "choice") %>%
  count(choice)
```

Everyone in the clean data responds to every choice task and the "none of the above" option doesn't appear to be abused.

```{r check-design, warning=FALSE, message=FALSE}
design_orig <- read_csv(here("data", "derived_data", "design.csv")) %>% 
  rename(
    version = question_set,
    task = question,
    alt = choice
  ) %>% 
  gather(key = "atts", value = "levels", -c(version:alt))

design_surv <- data_clean %>% 
  select(contains("CBCONJOINT"), -revision_CBCONJOINT) %>% 
  gather(key = "temp", value = "level_names", -vers_CBCONJOINT) %>% 
  separate(temp, into = c("atts", "task", "alt"), sep = "\\.") %>% 
  separate(alt, into = c("alt", "temp")) %>% 
  select(vers_CBCONJOINT, task, alt, atts, level_names) %>% 
  rename(version = vers_CBCONJOINT) %>% 
  mutate(
    task = as.numeric(task),
    alt = as.numeric(alt),
    atts = recode(
      atts,
      "732d72a6-25a3-4194-8416-461acae30837" = "Organization",
      "bd54b90a-efb3-4ac8-b65a-1ebeefcaf703" = "Issue area",
      "a6836e28-fd7a-4b71-b65e-f6269e5c883f" = "Financial transparency",
      "c2249c1a-11fa-4395-a0a3-61d9648a2e55" = "Accountability",
      "db44c406-390e-4fc4-875b-65796aa95ffa" = "Funding sources",
      "85f14c50-61cc-461c-8e77-19453b83d76d" = "Relationship with host government"
    )
  ) %>%
  distinct() %>% 
  arrange(version, task, alt)

design_orig %>% 
  count(atts, levels) %>% 
  mutate(prop = n/sum(n))

design_surv %>% 
  count(atts, level_names) %>% 
  mutate(prop = n/sum(n))

design_surv %>% 
  left_join(
    design_orig,
    by = c("version", "task", "alt", "atts")
  ) %>% 
  count(atts, levels, level_names)
```

The design matrix as reported in the survey is correct. We can use the separate design file matched by design version for estimation.

## Remove extraneous information

```{r shrink-data}
data <- data_clean %>% 
  mutate(id = row_number()) %>% 
  rename(version = vers_CBCONJOINT) %>% 
  select(-contains("_TEXT")) %>% 
  select(id, version, Q1.1:Q5.18)
```

# Transform data

## Covariates

```{r transform-covariates}
data <- data %>% 
  mutate(
    # *Q2.1*: How often do you follow national news? 
    # 
    # - Multiple times a day → At least once a day
    # - Every day → At least once a day
    # - Once a week → Once a week
    # - Hardly ever → Rarely (base case)
    # - Never → Rarely
    Q2.1 = factor(
      case_when(
        Q2.1 == 1 ~ "At least once a day",
        Q2.1 == 2 ~ "At least once a day",
        Q2.1 == 3 ~ "Once a week",
        Q2.1 == 4 ~ "Rarely",
        Q2.1 == 5 ~ "Rarely"
      ),
      levels = c("Rarely", "Once a week", "At least once a day"),
      labels = str_c(
        "Follow national news: ",
        c("Rarely", "Once a week", "At least once a day")
      )
    ),
    # *Q2.2*: How often do you follow international news? 
    # 
    # - Always
    # - Sometimes
    # - Never
    Q2.2 = fct_relevel(
      factor(
        Q2.2, 
        labels = str_c(
          "Follow international news: ", 
          c("Always", "Sometimes", "Never")
        ),
      ),
      str_c(
        "Follow international news: ", 
        c("Never", "Sometimes", "Always")
      )
    )
  ) %>% 
  mutate_at(
    # *Q2.3*: Which mediums do you use to follow news? (Select all that apply.)
    # 
    # - TV
    # - Print 
    # - Online (excluding social media) 
    # - Social media
    # - Radio
    # - Email newsletters
    # - News app
    vars(contains("Q2.3")), coalesce, 0
  ) %>%
  mutate(
    Q2.3_1 = factor(Q2.3_1, labels = str_c("Medium to follow news: ", c("No TV", "TV"))),
    Q2.3_2 = factor(Q2.3_2, labels = str_c("Medium to follow news: ", c("No Print", "Print"))),
    Q2.3_3 = factor(Q2.3_3, labels = str_c(
      "Medium to follow news: ", 
      c("No Online (excluding social media)", "Online (excluding social media)")
    )),
    Q2.3_4 = factor(Q2.3_4, labels = str_c(
      "Medium to follow news: ", 
      c("No Social media", "Social media")
    )),
    Q2.3_5 = factor(Q2.3_5, labels = str_c("Medium to follow news: ", c("No Radio", "Radio"))),
    Q2.3_6 = factor(Q2.3_6, labels = str_c(
      "Medium to follow news: ", 
      c("No Email newsletters", "Email newsletters")
    )),
    Q2.3_7 = factor(Q2.3_7, labels = str_c(
      "Medium to follow news: ", 
      c("No News app", "News app")
    )),
    # *Q2.4*: How often would you say you follow what's going on in government and public affairs?
    # 
    # - Most of the time → Often
    # - Some of the time → Often
    # - Only now and then → Not often (base case)
    # - Hardly at all → Not often (base case)
    Q2.4 = factor(
      case_when(
        Q2.4 == 1 ~ "Often",
        Q2.4 == 2 ~ "Often",
        Q2.4 == 3 ~ "Not often",
        Q2.4 == 4 ~ "Not often"
      ),
      levels = c("Not often", "Often"),
      labels = str_c(
        "Follow government and public affairs: ",
        c("Not often", "Often")
      )
    ),
    # *Q2.5*: How often do you donate to charity (with either cash or in-kind)?
    # 
    # - Once a week → At least once a month
    # - Once a month → At least once a month
    # - Once every three months → More than once a month, less than once a year
    # - Once every six months → More than once a month, less than once a year
    # - Once a year → More than once a month, less than once a year (base case)
    Q2.5 = factor(
      case_when(
        Q2.5 == 1 ~ "At least once a month",
        Q2.5 == 2 ~ "At least once a month",
        Q2.5 == 3 ~ "More than once a month, less than once a year",
        Q2.5 == 4 ~ "More than once a month, less than once a year",
        Q2.5 == 5 ~ "More than once a month, less than once a year"
      ),
      levels = c("More than once a month, less than once a year", "At least once a month"),
      labels = str_c(
        "Donation frequency: ",
        c("More than once a month, less than once a year", "At least once a month")
      )
    ),
    # *Q2.6*: How much did you donate to charity last year?
    # 
    # - \$1 to \$49
    # - \$50 to \$99
    # - \$100 to \$499
    # - \$500 to \$999
    # - \$1000 to \$4,999
    # - \$5000 to \$9,999
    # - \$10,000 or more
    Q2.6 = factor(
      Q2.6, 
      labels = str_c("Donated last year: ", c(
        "$1-$49", "$50-$99", "$100-$499", "$500-$999", 
        "$1000-$4,999", "$5000-$9,999", "$10,000+")),
      ordered = TRUE
    ),
    # *Q2.7*: On a scale of not at all important (1) to essential (7), how important is it for you to trust charities?
    # 
    # - 1 (Not at all important)
    # - 2
    # - 3
    # - 4
    # - 5
    # - 6
    # - 7 (Essential)
    Q2.7 = factor(
      Q2.7,
      labels = str_c("Importance of trusing charities: ", c(
        "Not at all important", "Very unimportant", "Somewhat unimportant", 
        "Neutral", "Somewhat important", "Very important", "Essential"
      )),
      ordered = TRUE
    ),
    # *Q2.8*: On a scale of no trust at all (1) to complete trust (7), how much do you trust charities?
    # 
    # - 1 (No trust at all)
    # - 2
    # - 3
    # - 4
    # - 5
    # - 6
    # - 7 (Complete trust)
    Q2.8 = factor(
      Q2.8,
      labels = str_c("Trust in charities: ", c(
        "No trust at all", "Very little trust", "Little trust", 
        "Neutral", "Some trust", "A lot of trust", "Complete trust"
      )),
      ordered = TRUE
    ),
    # *Q2.9*: Have you volunteered in the past 12 months?
    # 
    # - Yes
    # - No
    Q2.9 = fct_relevel(
      factor(Q2.9, labels = str_c("Volunteered in past 12 months: ", c("Yes", "No"))), 
      str_c("Volunteered in past 12 months: ", c("No", "Yes"))
    ),
    # *Q2.10*: How often do you volunteer?
    # 
    # - NA → Haven't volunteered in past 12 months (base case)
    # - Once a week → At least once a month
    # - Once a month → At least once a month
    # - Once every three months → More than once a month, less than once a year
    # - Once every six months → More than once a month, less than once a year
    # - Once a year → More than once a month, less than once a year
    # - Once every few years → Rarely
    Q2.10 = factor(
      case_when(
        is.na(Q2.10) == TRUE ~ "Haven't volunteered in past 12 months",
        Q2.10 == 1 ~ "At least once a month",
        Q2.10 == 2 ~ "At least once a month",
        Q2.10 == 3 ~ "More than once a month, less than once a year",
        Q2.10 == 4 ~ "More than once a month, less than once a year",
        Q2.10 == 5 ~ "More than once a month, less than once a year",
        Q2.10 == 6 ~ "Rarely"
      ),
      levels = c("Haven't volunteered in past 12 months", "Rarely", "More than once a month, less than once a year", "At least once a month"),
      labels = str_c(
        "Volunteer frequency: ",
        c("Haven't volunteered in past 12 months", "Rarely", "More than once a month, less than once a year", "At least once a month")
      )
    ),
    # *Q5.1*: Did you vote in the last election?
    # 
    # - Yes
    # - No
    Q5.1 = fct_relevel(
      factor(Q5.1, labels = str_c("Voted in last election: ", c("Yes", "No"))), 
      str_c("Voted in last election: ", c("No", "Yes"))
    ),
    # *Q5.2*: On a scale of extremely liberal (1) to extremely conservative (7), how would you describe your political views?
    # 
    # - 1 (Extremely liberal)
    # - 2
    # - 3
    # - 4
    # - 5
    # - 6
    # - 7 (Extremely conservative)
    Q5.2 = factor(
      Q5.2,
      labels = str_c("Liberal to Conservative: ", c(
        "Extremely liberal", "Somewhat liberal", "Slightly liberal", "Moderate",
        "Slightly conservative", "Somewhat conservative", "Extremely conservative"
      )),
      ordered = TRUE
    ),
    # *Q5.3*: Here is a list of different types of voluntary organizations. For each organization, indicate whether you are an active member, an inactive member, or not a member of that type of organization:
    # 
    # |                                         | Active member | Inactive member | Don't belong |
    # | --------------------------------------- | :-----------: | :-------------: | :----------: |
    # | Church or religious organization        |       •       |        •        |      •       |
    # | Sport or recreational organization      |       •       |        •        |      •       |
    # | Art, music, or educational organization |       •       |        •        |      •       |
    # | Labor union                             |       •       |        •        |      •       |
    # | Political party                         |       •       |        •        |      •       |
    # | Environmental organization              |       •       |        •        |      •       |
    # | Professional association                |       •       |        •        |      •       |
    # | Humanitarian or charitable organization |       •       |        •        |      •       |
    # | Consumer organization                   |       •       |        •        |      •       |
    # | Other organization                      |       •       |        •        |      •       |
    Q5.3_1 = fct_relevel(
      factor(Q5.3_1, labels = str_c(
        "Church or religious organization: ", 
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Church or religious organization: ", 
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_2 = fct_relevel(
      factor(Q5.3_2, labels = str_c(
        "Sport or recreational organization: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Sport or recreational organization: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_3 = fct_relevel(
      factor(Q5.3_3, labels = str_c(
        "Art, music, or educational organization: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Art, music, or educational organization: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_4 = fct_relevel(
      factor(Q5.3_4, labels = str_c(
        "Labor union: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Labor union: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_5 = fct_relevel(
      factor(Q5.3_5, labels = str_c(
        "Political party: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Political party: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_6 = fct_relevel(
      factor(Q5.3_6, labels = str_c(
        "Environmental organization: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Environmental organization: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_7 = fct_relevel(
      factor(Q5.3_7, labels = str_c(
        "Professional association: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Professional association: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_8 = fct_relevel(
      factor(Q5.3_8, labels = str_c(
        "Humanitarian or charitable organization: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Humanitarian or charitable organization: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_9 = fct_relevel(
      factor(Q5.3_9, labels = str_c(
        "Consumer organization: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Consumer organization: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    Q5.3_10 = fct_relevel(
      factor(Q5.3_10, labels = str_c(
        "Other organization: ",
        c("Active member", "Inactive member", "Don't belong"))
      ), 
      str_c(
        "Other organization: ",
        c("Don't belong", "Inactive member", "Active member")
      )
    ),
    # *Q5.4*: Historically, how involved have you been in activist causes?
    # 
    # - Extremely involved → Involved
    # - Very involved → Involved
    # - Moderately involved → Involved
    # - Slightly involved → Not involved (base case)
    # - Never involved → Not involved
    Q5.4 = factor(
      case_when(
        Q5.4 == 1 ~ "Involved",
        Q5.4 == 2 ~ "Involved",
        Q5.4 == 3 ~ "Involved",
        Q5.4 == 4 ~ "Not involved",
        Q5.4 == 5 ~ "Not involved",
      ),
      levels = c("Not involved", "Involved"),
      labels = str_c(
        "History of activism: ",
        c("Not involved", "Involved")
      )
    ),
    # *Q5.5*: Historically, how involved has your family been in activist causes?
    # 
    # - Extremely involved → Involved
    # - Very involved → Involved
    # - Moderately involved → Involved
    # - Slightly involved → Not involved (base case)
    # - Never involved → Not involved
    Q5.5 = factor(
      case_when(
        Q5.5 == 1 ~ "Involved",
        Q5.5 == 2 ~ "Involved",
        Q5.5 == 3 ~ "Involved",
        Q5.5 == 4 ~ "Not involved",
        Q5.5 == 5 ~ "Not involved"
      ),
      levels = c("Not involved", "Involved"),
      labels = str_c(
        "History of family activism: ",
        c("Not involved", "Involved")
      )
    ),
    # *Q5.6*: On a scale of no trust (1) to complete trust (7), how much do you trust political institutions and the state?
    # 
    # - 1 (No trust)
    # - 2
    # - 3
    # - 4
    # - 5
    # - 6
    # - 7 (Complete trust)
    Q5.6 = factor(
      Q5.6,
      labels = str_c("Trust in political institutions: ",c(
        "No trust at all", "Very little trust", "Little trust", 
        "Neutral", "Some trust", "A lot of trust", "Complete trust"
      )),
      ordered = TRUE
    ),
    # *Q5.7*: Have you ever traveled to a developing country? 
    # 
    # - Yes
    # - No
    Q5.7 = fct_relevel(
      factor(Q5.7, labels = str_c("Traveled to a developing country: ", c("Yes", "No"))), 
      str_c("Traveled to a developing country: ", c("No", "Yes"))
    ),
    # *Q5.8*: How often do you attend religious or worship services, not including weddings and funerals?
    # 
    # - More than once a week → At least once a month
    # - Once a week → At least once a month
    # - Once or twice a month → At least once a month
    # - A few times a year → Rarely (base case)
    # - Seldom → Rarely
    # - Never → Rarely
    # - Don't know → Not sure
    Q5.8 = factor(
      case_when(
        Q5.8 == 1 ~ "At least once a month",
        Q5.8 == 2 ~ "At least once a month",
        Q5.8 == 3 ~ "At least once a month",
        Q5.8 == 4 ~ "Rarely",
        Q5.8 == 5 ~ "Rarely",
        Q5.8 == 6 ~ "Rarely",
        Q5.8 == 7 ~ "Not sure"
      ),
      levels = c("Not sure", "Rarely", "At least once a month"),
      labels = str_c(
        "Religious and worship service attendance: ",
        c("Not sure", "Rarely", "At least once a month")
      )
    ),
    # *Q5.9*: How important is religion in your life?
    # 
    # - Extremely important → Important
    # - Very important → Important
    # - Moderately important → Important
    # - Slightly important → Not important (base case)
    # - Not at all important → Not important
    Q5.9 = factor(
      case_when(
        Q5.9 == 1 ~ "Important",
        Q5.9 == 2 ~ "Important",
        Q5.9 == 3 ~ "Important",
        Q5.9 == 4 ~ "Not important",
        Q5.9 == 5 ~ "Not important"
      ),
      levels = c("Not important", "Important"),
      labels = str_c(
        "Importance of religion: ",
        c("Not important", "Important")
      )
    ),
    # *Q5.10*: What is your current religion, if any?
    # 
    # - Catholic (including Roman Catholic and Orthodox)
    # - Protestant (United Church of Canada, Anglican, Orthodox, Baptist, Lutheran)
    # - Christian Orthodox
    # - Jewish
    # - Muslim
    # - Sikh
    # - Hindu
    # - Buddhist
    # - Atheist (do not believe in God)
    # - Other: _________
    Q5.10 = factor(
      Q5.10, 
      labels = str_c("Religion: ", c(
        "Catholic", "Protestant", "Christian Orthodox", "Jewish", 
        "Muslim", "Sikh", "Hindu", "Buddhist", "Atheist", "Other")), 
      ordered = TRUE
    ),
    # *Q5.11*: On a scale of strongly agree (1) to strongly disagree (7), rate your response to the following statement: People should be more charitable towards others in society.
    # 
    # - 1 (Strongly agree)
    # - 2
    # - 3
    # - 4
    # - 5
    # - 6
    # - 7 (Strongly disagree)
    Q5.11 = fct_relevel(
      factor(
        Q5.11,
        labels = str_c("People should be more charitable: ", c(
          "Strongly agree", "Agree", "Somewhat agree", 
          "Neutral", "Somewhat disagree", "Disagree", "Strongly disagree"
        ))
      ),
      str_c("People should be more charitable: ", c(
        "Strongly disagree", "Disagree", "Somewhat disagree",
        "Neutral", "Somewhat agree", "Agree", "Strongly agree"
      ))
    ),
    # *Q5.12*: What is your gender?
    # 
    # - Male
    # - Female
    # - Transgender
    # - Prefer not to say
    # - Other: _________
    Q5.12 = factor(
      Q5.12, 
      labels = str_c("Gender: ", c(
        "Male", "Female", "Transgender", 
        "Prefer not to say", "Other"
      )), 
      ordered = TRUE
    ),
    # *Q5.13*: Are you now married, widowed, divorced, separated, or never married?
    # 
    # - Married
    # - Widowed
    # - Divorced
    # - Separated
    # - Never married
    Q5.13 = factor(
      Q5.13, 
      labels = str_c("Marital status: ", c(
        "Married", "Widowed", "Divorced", 
        "Separated", "Never married"
      )), 
      ordered = TRUE
    ),
    # *Q5.14*: What is the highest degree or level of school you have completed?
    # 
    # - Less than high school
    # - High school graduate
    # - Some college
    # - 2 year degree
    # - 4 year degree 
    # - Graduate or professional degree
    # - Doctorate
    Q5.14 = factor(
      Q5.14, 
      labels = str_c("Education: ", c(
        "Less than high school", "High school graduate", 
        "Some college", "2 year degree", "4 year degree", 
        "Graduate or professional degree", "Doctorate"
      )),
      ordered = TRUE
    ),
    # *Q5.15*: What is your annual household income before taxes?
    # 
    # - Less than $10,000 → More/less than median (base case = less than median)
    # - \$10,000 to \$19,999 → More/less than median
    # - \$20,000 to \$29,999 → More/less than median
    # - \$30,000 to \$39,999 → More/less than median
    # - \$40,000 to \$49,999 → More/less than median
    # - \$50,000 to \$59,999 → More/less than median
    # - \$60,000 to \$69,999 → More/less than median
    # - \$70,000 to \$79,999 → More/less than median
    # - \$80,000 to \$89,999 → More/less than median
    # - \$90,000 to \$99,999 → More/less than median
    # - \$100,000 to \$149,999 → More/less than median
    # - \$150,000 to \$199,999 → More/less than median
    # - \$200,000 to \$299,999 → More/less than median
    # - \$300,000 or more → More/less than median
    Q5.15 = factor(
      case_when(
        Q5.15 > 5 ~ "More than median", # Median income in 2017 was $61,372
        Q5.15 <= 5 ~ "Less than median"
      ),
      levels = c("Less than median", "More than median"),
      labels = str_c(
        "Income: ",
        c("Less than median", "More than median")
      )
    )
  ) %>% 
  mutate_at(
    # *Q5.16*: Choose one or more races that you consider yourself to be:
    # 
    # - White
    # - Black or African American
    # - American Indian or Alaska Native
    # - Asian
    # - Native Hawaiian or Pacific Islander
    # - Other: _________
    vars(contains("Q5.16")), coalesce, 0
  ) %>% 
  mutate(
    Q5.16_1 = factor(Q5.16_1, labels = str_c("Race: ", c("Not White", "White"))),
    Q5.16_2 = factor(Q5.16_2, labels = str_c(
      "Race: ", c("Not Black or African American", "Black or African American"))
    ),
    Q5.16_3 = factor(Q5.16_3, labels = str_c(
      "Race: ", c("Not American Indian or Alaska Native", "American Indian or Alaska Native"))
    ),
    Q5.16_4 = factor(Q5.16_4, labels = str_c("Race: ", c("Not Asian", "Asian"))),
    Q5.16_5 = factor(Q5.16_5, labels = str_c(
      "Race: ", 
      c("Not Native Hawaiian or Pacific Islander", "Native Hawaiian or Pacific Islander")
    )),
    Q5.16_6 = factor(Q5.16_6, labels = str_c("Race: ", c("Not Other", "Other"))),
    # *Q5.17*: How old are you?
    # 
    # - Under 18 → More/less than median (base case = less than median)
    # - 18 - 24 → More/less than median
    # - 25 - 34 → More/less than median
    # - 35 - 44 → More/less than median
    # - 45 - 54 → More/less than median
    # - 55 - 64 → More/less than median
    # - 65 - 74 → More/less than median
    # - 75 - 84 → More/less than median
    # - 85 or older → More/less than median
    Q5.17 = factor(
      case_when(
        Q5.17 >= 3 ~ "More than median", # Median age in 2017 was 36
        Q5.17 < 3 ~ "Less than median"
      ),
      levels = c("Less than median", "More than median"),
      labels = str_c(
        "Age: ", 
        c("Less than median", "More than median")
      )
    )
  )
```

## Design matrix

```{r transform-design}
# Load design
design <- read_csv(here::here("data", "derived_data", "dummy_design.csv")) %>%
  select(-X1) %>%
  rename(
    version = question_set,
    task = question,
    alt = choice
  )

# Create the array for design X.
X <- array(
  data = NA,
  dim = c(
    nrow(data),             # Number of respondents.
    max(design$task),       # Number of choice tasks per respondent.
    (max(design$alt) + 1),  # Number of product alternatives per choice task.
    (ncol(design) - 3 + 1)  # Number of (estimable) attribute levels.
  )
)

# Reintroduce org1 and add the outside option coded as all zeros.
for (n in 1:dim(X)[1]) {
  # Filter for respondent n.
  X_n <- design %>% filter(version == data$version[n])
  for (s in 1:dim(X)[2]) {
    # Filter for task s and include placeholder for org1.
    X_s <- X_n %>%
      filter(task == s) %>%
      mutate(org1 = 0) %>%
      select(org1, org2:gov_rel3)
    for (p in 1:(dim(X)[3]-1)) {
      # Filter for task s and alt p orgs.
      X_p <- X_n %>%
        filter(task == s, alt == p) %>%
        select(org2:org4) %>%
        as_vector()
      # Fill in org1 to complte org intercept.
      X_s[p, "org1"] <- ifelse(sum(X_p) == 1, 0, 1)
    }
    # Save modified design, including outside option.
    X[n, s,,] <- rbind(X_s, rep(0, ncol(X_s))) %>% as.matrix()
  }
}
```

# Save clean data with transformations

```{r save-clean-data}
write_rds(data, here("data", "derived_data", "final_data.rds"))
write_rds(X, here("data", "derived_data", "final_design.rds"))
```

<!-- select(-c("732d72a6-25a3-4194-8416-461acae30837.1.1_CBCONJOINT":"revision_CBCONJOINT")) -->
<!-- design <- read_csv( -->
<!--   here::here("data", "conjointqsf_final.csv"), -->
<!--   skip = 2, -->
<!--   col_names = col_names -->
<!-- ) %>%  -->
<!--   select(contains("CBCONJOINT")) -->

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=90}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>