### 2021-10-15 analysis/01_experimental-design.Rmd
Included an index-coded design for computing contrasts

---
title: "Check experimental design"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Load the design and check level balance

The experimental design was created as part of the conjoint study in Qualtrics. Here we check to make sure that there is a balance across attribute levels (i.e., levels show up a roughly equal number of times for each attribute).

```{r load-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(fastDummies)
library(here)

# Load design
design <- read_csv(here("data", "derived_data", "design.csv"))

# Check to make sure there is a balance across levels
as_tibble(design) %>%
  select(Organization:`Relationship with host government`) %>%
  gather(key = atts, value = lvls) %>%
  ggplot(aes(x = lvls)) +
  geom_bar() +
  facet_wrap(~ atts, scales = "free", nrow = 2)
```

# Dummy-code the design

Now we prepare the experimental design for use in estimating the model by dropping the first level of each attribute, which are all discrete, as the reference level.

```{r dummy-design, warning=FALSE, message=FALSE}
# Convert the attributes in design into factors
dummy_design <- as.data.frame(design)
for(i in 4:ncol(dummy_design)) {
  dummy_design[,i] <- as.factor(dummy_design[,i])
}

# Rename the variables.
colnames(dummy_design)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use the output of lm to get a dummy-coded version of the design
dummy_design$y <- matrix(rnorm(nrow(dummy_design)), ncol=1)
out <- lm(y ~ ., dummy_design, x = TRUE)
dummy_design <- out$x[,-1]
rownames(dummy_design) <- NULL

# Save dummy-coded design output
write.csv(dummy_design, here("data", "derived_data", "dummy_design.csv"))
```

# Index-code the design

We also need to consider how we are computing contrasts to answer our specified hypotheses. To do so, we index-code where every attribute level is included such that every attribute is its own implied intercept. Note that for a non-Bayesian model, index coding in this way would not be identified. However, for a Bayesian model with proper priors, all parameters are identified. Additionally, the contrasts even of unidentified parameters are identified.

```{r index-design, warning=FALSE, message=FALSE}
# Convert the attributes in design into factors
index_design <- as.data.frame(design)
for(i in 4:ncol(index_design)) {
  index_design[,i] <- as.factor(index_design[,i])
}

# Rename the variables.
colnames(index_design)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get a index-coded version of all the attribute levels.
index_design <- dummy_cols(
  index_design, 
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = FALSE,
  remove_selected_columns = TRUE
)

# Rename the index variables.
colnames(index_design)[4:21] <- c(
  str_c("org", 1:4), str_c("issue", 1:4), str_c("transp", 1:2), 
  str_c("account", 1:2), str_c("fund", 1:3), str_c("gov_rel", 1:3)
)

# Save index-coded design output
write.csv(index_design, here("data", "derived_data", "index_design.csv"))
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=90}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>