### 2021-10-18 analysis/01_experimental-design.Rmd
Cleaned up dummy coding, removed index coding, and built out interactions for the design

---
title: "Check experimental design"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Load the design and check level balance

The experimental design was created as part of the conjoint study in Qualtrics. Here we check to make sure that there is a balance across attribute levels (i.e., levels show up a roughly equal number of times for each attribute).

```{r load-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(donorheuristics)
library(fastDummies)
library(here)

# Load design
design <- read_csv(here("data", "derived_data", "design.csv"))

# Check to make sure there is a balance across levels
as_tibble(design) %>%
  select(Organization:`Relationship with host government`) %>%
  gather(key = atts, value = lvls) %>%
  ggplot(aes(x = lvls)) +
  geom_bar() +
  facet_wrap(~ atts, scales = "free", nrow = 2)
```

# Dummy-code the design

Now we prepare the experimental design for use in estimating the model by dropping the first level of each attribute, which are all discrete, as the reference level.

```{r dummy-design, warning=FALSE, message=FALSE}
# Convert the attributes in design into factors
dummy_design <- as.data.frame(design)
for(i in 4:ncol(dummy_design)) {
  dummy_design[,i] <- as.factor(dummy_design[,i])
}

# Rename the variables.
colnames(dummy_design)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get a index-coded version of all the attribute levels.
dummy_design <- dummy_cols(
  dummy_design,
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = TRUE,
  remove_selected_columns = TRUE
)

# Rename the index variables.
colnames(dummy_design)[4:15] <- c(
  str_c("org", 2:4), str_c("issue", 2:4), str_c("transp", 2),
  str_c("account", 2), str_c("fund", 2:3), str_c("gov_rel", 2:3)
)

# Save dummy-coded design output
write.csv(dummy_design, here("data", "derived_data", "dummy_design.csv"))
```

# Interact-code the design

We have a number of hypotheses referencing specific interactions. To answer these hypotheses, we need a specific intercept-only design with each of the needed interactions.

```{r interact-design, warning=FALSE, message=FALSE}
# Convert the attributes in design into factors
interact_design <- as.data.frame(design)
for(i in 4:ncol(interact_design)) {
  interact_design[,i] <- as.factor(interact_design[,i])
}

# Rename the variables.
colnames(interact_design)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get a dummy-coded version of all the attribute levels.
interact_design <- dummy_cols(
  interact_design, 
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = FALSE,
  remove_selected_columns = TRUE
)

# Rename the dummy-coded variables.
colnames(interact_design)[4:21] <- c(
  str_c("org", 1:4), str_c("issue", 1:4), str_c("transp", 1:2), 
  str_c("account", 1:2), str_c("fund", 1:3), str_c("gov_rel", 1:3)
)

# Build out all the necessary interactions.
interact_design %>% 
  as_tibble() %>% 
  mutate(
    # Hypothesis 2b.
    amnesty_criticized = org1 * gov_rel2,
    amnesty_crackdown = org1 * gov_rel3,
    greenpeace_criticized = org2 * gov_rel2,
    greenpeace_crackdown = org2 * gov_rel3,
    oxfam_criticized = org3 * gov_rel2,
    oxfam_crackdown = org3 * gov_rel3,
    redcross_criticized = org4 * gov_rel2,
    redcross_crackdown = org4 * gov_rel3,
    # Hypothesis 3b.
    emergency_criticized = issue1 * gov_rel2,
    emergency_crackdown = issue1 * gov_rel3,
    environment_criticized = issue2 * gov_rel2,
    environment_crackdown = issue2 * gov_rel3,
    humanrights_criticized = issue3 * gov_rel2,
    humanrights_crackdown = issue3 * gov_rel3,
    refugees_criticized = issue4 * gov_rel2,
    refugees_crackdown = issue4 * gov_rel3,
    # Hypothesis 4b.
    smalldonors_criticized = fund1 * gov_rel2,
    smalldonors_crackdown = fund1 * gov_rel3,
    wealthydonors_criticized = fund2 * gov_rel2,
    wealthydonors_crackdown = fund2 * gov_rel3,
    govgrants_criticized = fund3 * gov_rel2,
    govgrants_crackdown = fund3 * gov_rel3,
    # Hypothesis 4c.
    emergency_smalldonors_crackdown = issue1 * fund1 * gov_rel3,
    emergency_wealthydonors_crackdown = issue1 * fund2 * gov_rel3,
    emergency_govgrants_crackdown = issue1 * fund3 * gov_rel3,
    environment_smalldonors_crackdown = issue2 * fund1 * gov_rel3,
    environment_wealthydonors_crackdown = issue2 * fund2 * gov_rel3,
    environment_govgrants_crackdown = issue2 * fund3 * gov_rel3,
    humanrights_smalldonors_crackdown = issue3 * fund1 * gov_rel3,
    humanrights_wealthydonors_crackdown = issue3 * fund2 * gov_rel3,
    humanrights_govgrants_crackdown = issue3 * fund3 * gov_rel3,
    refugees_smalldonors_crackdown = issue4 * fund1 * gov_rel3,
    refugees_wealthydonors_crackdown = issue4 * fund2 * gov_rel3,
    refugees_govgrants_crackdown = issue4 * fund3 * gov_rel3,
    # Hypothesis 5b.
    transp_criticized = transp1 * gov_rel2,
    transp_crackdown = transp1 * gov_rel3,
    nottransp = ifelse(transp1 == 0, 1, 0),
    nottransp_criticized = nottransp * gov_rel2,
    nottransp_crackdown = nottransp * gov_rel3,
    # Hypothesis 5c.
    transp_smalldonors_criticized = transp1 * fund1 * gov_rel2,
    transp_smalldonors_crackdown = transp1 * fund1 * gov_rel3,
    transp_wealthydonors_criticized = transp1 * fund2 * gov_rel2,
    transp_wealthydonors_crackdown = transp1 * fund2 * gov_rel3,
    transp_govgrants_criticized = transp1 * fund3 * gov_rel2,
    transp_govgrants_crackdown = transp1 * fund3 * gov_rel3,
    nottransp_smalldonors_criticized = nottransp * fund1 * gov_rel2,
    nottransp_smalldonors_crackdown = nottransp * fund1 * gov_rel3,
    nottransp_wealthydonors_criticized = nottransp * fund2 * gov_rel2,
    nottransp_wealthydonors_crackdown = nottransp * fund2 * gov_rel3,
    nottransp_govgrants_criticized = nottransp * fund3 * gov_rel2,
    nottransp_govgrants_crackdown = nottransp * fund3 * gov_rel3,
    # Hypothesis 5d.
    transp_emergency_criticized = transp1 * issue1 * gov_rel2,
    transp_emergency_crackdown = transp1 * issue1 * gov_rel3,
    transp_environment_criticized = transp1 * issue2 * gov_rel2,
    transp_environment_crackdown = transp1 * issue2 * gov_rel3,
    transp_humanrights_criticized = transp1 * issue3 * gov_rel2,
    transp_humanrights_crackdown = transp1 * issue3 * gov_rel3,
    transp_refugees_criticized = transp1 * issue4 * gov_rel2,
    transp_refugees_crackdown = transp1 * issue4 * gov_rel3,
    nottransp_emergency_criticized = nottransp * issue1 * gov_rel2,
    nottransp_emergency_crackdown = nottransp * issue1 * gov_rel3,
    nottransp_environment_criticized = nottransp * issue2 * gov_rel2,
    nottransp_environment_crackdown = nottransp * issue2 * gov_rel3,
    nottransp_humanrights_criticized = nottransp * issue3 * gov_rel2,
    nottransp_humanrights_crackdown = nottransp * issue3 * gov_rel3,
    nottransp_refugees_criticized = nottransp * issue4 * gov_rel2,
    nottransp_refugees_crackdown = nottransp * issue4 * gov_rel3,
    # Hypothesis 5e.
    transp_smalldonors_emergency_criticized = transp1 * fund1 * issue1 * gov_rel2,
    transp_smalldonors_emergency_crackdown = transp1 * fund1 * issue1 * gov_rel3,
    transp_smalldonors_environment_criticized = transp1 * fund1 * issue2 * gov_rel2,
    transp_smalldonors_environment_crackdown = transp1 * fund1 * issue2 * gov_rel3,
    transp_smalldonors_humanrights_criticized = transp1 * fund1 * issue3 * gov_rel2,
    transp_smalldonors_humanrights_crackdown = transp1 * fund1 * issue3 * gov_rel3,
    transp_smalldonors_refugees_criticized = transp1 * fund1 * issue4 * gov_rel2,
    transp_smalldonors_refugees_crackdown = transp1 * fund1 * issue4 * gov_rel3,
    nottransp_smalldonors_emergency_criticized = nottransp * fund1 * issue1 * gov_rel2,
    nottransp_smalldonors_emergency_crackdown = nottransp * fund1 * issue1 * gov_rel3,
    nottransp_smalldonors_environment_criticized = nottransp * fund1 * issue2 * gov_rel2,
    nottransp_smalldonors_environment_crackdown = nottransp * fund1 * issue2 * gov_rel3,
    nottransp_smalldonors_humanrights_criticized = nottransp * fund1 * issue3 * gov_rel2,
    nottransp_smalldonors_humanrights_crackdown = nottransp * fund1 * issue3 * gov_rel3,
    nottransp_smalldonors_refugees_criticized = nottransp * fund1 * issue4 * gov_rel2,
    nottransp_smalldonors_refugees_crackdown = nottransp * fund1 * issue4 * gov_rel3
  ) %>% 
  # Drop first levels and the nottransp helper variable.
  select(-c(contains("1"), nottransp))

# Save interact-coded design output
write.csv(interact_design, here("data", "derived_data", "interact_design.csv"))
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=90}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>