### 2021-12-04 analysis/04_figs-tables.Rmd
Attempt #1 at contrasts with main effects. Hypotheses 4c, 5d and 5e need to be fixed.
I forgot to commit the actual coding changes, somehow. This should do it.

---
title: "Figures and tables for paper"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    pandoc_args:
      - "--default-image-extension=png"
editor_options: 
  chunk_output_type: console
---

# Load packages, data, and draws

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(tidybayes)
library(broom)
# library(ggridges)
library(ggraph)
library(ggdag)
library(pander)
library(patchwork)
library(scales)
library(here)

# Load custom plot functions
source(here("R", "graphics.R"))

# General settings
source(here("analysis", "options.R"))

# Make all the randomness reproducible
set.seed(1234)

# Load data
results <- read_rds(here("data", "derived_data", "final_data.rds"))

# Load draws
intercept <- 1                            # Intercept-only
interactions <- 1                         # Include interactions with the intercept-only model?
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 0 # + Demographics

if (intercept == 1 & interactions == 0) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "intercept.rds")
  ) %>% 
    mutate(model = "Intercept only")
}
if (intercept == 1 & interactions == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "interactions.rds")
  ) %>% 
    mutate(model = "Intercept and interactions")
}
if (public_affairs == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "public_affairs.rds")
  ) %>% 
    mutate(model = "Public affairs")
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7 # Public affairs
    )
}
if (political_ideology == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "political_ideology.rds")
  ) %>% 
    mutate(model = "Political ideology")
  covariates <- results %>%
    select(
      Q5.2 # Political ideology
    )
}
if (social_views == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "social_views.rds")
  ) %>% 
    mutate(model = "Social views")
  covariates <- results %>%
    select(
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10 # Social views
    )
}
if (charity_voluntarism == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "charity_voluntarism.rds")
  ) %>% 
    mutate(model = "Charity and voluntarism")
  covariates <- results %>%
    select(
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10 # Charity and voluntarism
    )
}
if (demographics == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "demographics.rds")
  ) %>% 
    mutate(model = "Demographics")
  covariates <- results %>%
    select(
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17 # Demographics
    )
}
if (public_political == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "public_political.rds")
  ) %>% 
    mutate(model = "Public affairs and political ideology")
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2                                   # Political ideology
    )
}
if (public_political_social == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "public_political_social.rds")
  ) %>% 
    mutate(model = "Public affairs, political ideology, and social views")
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2,                                  # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10         # Social views
    )
}
if (public_political_social_charity == 1) {
  draws <- read_rds(
    here::here("analysis", "output", "posterior_draws", "public_political_social_charity.rds")
  ) %>% 
    mutate(
      model = "Public affairs, political ideology, social views, charity and voluntarism"
    )
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10  # Charity and voluntarism
    )
}
if (public_political_social_charity_demo == 1) {
  draws <- read_rds(
    here::here(
      "analysis", 
      "output", 
      "posterior_draws", 
      "public_political_social_charity_demo.rds"
    )
  ) %>% 
    mutate(
      model = "Public affairs, 
      political ideology, 
      social views, 
      charity and voluntarism, 
      and demographics"
    )
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10, # Charity and voluntarism
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17               # Demographics
    )
}

# Name levels labels
level_labels <- c(
  str_c("Organization: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross")),
  str_c("Issue area: ", c("Environment", "Human rights", "Refugee relief")),
  str_c("Organizational practices: ", c("Financial transparency", "Accountability")),
  str_c("Funding sources: ", c("Handful of wealthy private donors", "Government grants")),
  str_c("Relationship with host government: ", c("Criticized", "Under crackdown"))
)

if (intercept == 1 & interactions == 1) {
  interaction_labels <- c(
    # Hypothesis 2b.
    str_c("Interaction: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"), " Criticized"),
    str_c("Interaction: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"), " Crackdown"),
    # Hypothesis 3b.
    str_c("Interaction: ", c("Environment", "Human rights", "Refugee relief"), " Criticized"),
    str_c("Interaction: ", c("Environment", "Human rights", "Refugee relief"), " Crackdown"),
    # Hypothesis 4b.
    str_c("Interaction: ", c("Handful of wealthy private donors", "Government grants"), " Criticized"),
    str_c("Interaction: ", c("Handful of wealthy private donors", "Government grants"), " Crackdown"),
    # Hypothesis 4c.
    str_c("Interaction: Emergency ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    str_c("Interaction: Environment ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    str_c("Interaction: Human rights ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    str_c("Interaction: Refugee relief ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    # Hypothesis 5b.
    str_c("Interaction: Financial transparency", c(" Criticized", " Crackdown")),
    # Hypothesis 5c.
    str_c("Interaction: Financial transparency ", c("Small private donors", "Wealthy private donors", "Government grants"), " Criticized"),
    str_c("Interaction: Financial transparency ", c("Small private donors", "Wealthy private donors", "Government grants"), " Crackdown"),
    # Hypothesis 5d.
    str_c("Interaction: Financial transparency Emergency", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Environment", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Human rights", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Refugee relief", c(" Criticized", " Crackdown")),
    # Hypothesis 5e.
    str_c("Interaction: Financial transparency Small private donoros Emergency", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Small private donoros Environment", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Small private donoros Human rights", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Small private donoros Refugee relief", c(" Criticized", " Crackdown"))
  )
  level_labels <- c(level_labels, interaction_labels)
}

level_lookup <- tibble(levels = level_labels) %>% 
  separate(levels, c("attribute", "level"), sep = ": ") %>% 
  mutate(levels_long = paste0(attribute, ": ", level),
         i = 1:n()) %>% 
  mutate_at(vars(attribute, level, levels_long), list(fct_inorder))

if (intercept != 1) {
  # Extract covariate labels
  covariate_lookup <- covariates %>% 
    pivot_longer(cols = starts_with("Q"), names_to = "question", values_to = "labels") %>% 
    bind_cols(
      covariates %>% 
        mutate_all(as.integer) %>% 
        rename_all(paste0, "_temp") %>% 
        pivot_longer(cols = starts_with("Q"), names_to = "question_temp", values_to = "level")
    ) %>% 
    distinct() %>% 
    filter(level != 1) %>% 
    separate(labels, c("covariate", "label"), sep = ": ") %>% 
    mutate(covariates_long = paste0(covariate, ": ", label), j = 2:(n() + 1)) %>% 
    select(covariate, label, covariates_long, j) %>% 
    bind_rows(
      tibble(
        covariate = "Intercept",
        label = "Intercept",
        covariates_long = "Intercept",
        j = 1
      )
    ) %>% 
    arrange(j) %>% 
    mutate_at(vars(covariate, label, covariates_long), list(fct_inorder))
}

# Join level and covariate labels to draws
draws <- draws %>% 
  left_join(level_lookup, by = "i")

if (intercept != 1) {
  draws <- draws %>% 
    left_join(covariate_lookup, by = "j")
}
```

```{r}
diff_cols <- function(col1, col2) {
  posterior_diffs <- draws %>% 
    filter(attribute %in% c({{col1}}, {{col2}})) %>% 
    select(.chain, .iteration, Theta, attribute) %>% 
    pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
    mutate(diff_theta = .[[4]] - .[[3]],
           diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

  diff_theta <- mean(posterior_diffs$diff_theta_e)
  p_greater0 <- mean(posterior_diffs$diff_theta_e > 0)
  p_less0 <- mean(posterior_diffs$diff_theta_e < 0)
  p_not0 <- ifelse(median(posterior_diffs$diff_theta_e) > 0, p_greater0, p_less0)

  phrase <- paste0("∆θ = ", round(diff_theta, 3), "; p(∆ ≠ 0) = ", round(p_not0, 3))

  return(list(posterior_diffs = posterior_diffs, diff_theta = diff_theta,
              p_greater0 = p_greater0, p_less0 = p_less0, p_not0 = p_not0,
              phrase = phrase))
}
```

# Causal pathway

Our theory and hypotheses are laid out in the causal pathway below. Organizational factors (**O**, or organizational practices like financial disclosure and accountability practices; **I**, or issue area; and **F**, or funding) all influence the decision to donate (**D**). Structural factors like an NGO's relationship with its host government (**G**) also influence donation behavior, but the relationship itself is shaped by both funding and issue area.

```{r causal-dag, fig.width=13/3, fig.height=2}
node_colors <- tribble(
  ~name, ~type,
  "D", "Outcome",
  "I", "Organizational factor",
  "F", "Organizational factor",
  "O", "Organizational factor",
  "G", "Structural factor"
) %>% 
  mutate(type = fct_inorder(type))

# Set up DAG structure
theory_dag <- dagify(D ~ I + F + O + G,
                     G ~ I + F,
                     outcome = "D",
                     exposure = "C") %>% 
  tidy_dagitty(layout = "dh", seed = 1234)

# Add node types/colors to DAG data
theory_dag$data <- theory_dag$data %>% 
  left_join(node_colors, by = "name")

# Make DAG plot
plot_dag <- ggplot(theory_dag, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(size = 6, mapping = aes(color = type)) +
  geom_dag_edges(start_cap = circle(4, "mm"),
                 end_cap = circle(4, "mm")) +
  geom_dag_text(size = pts(6), family = "Roboto Condensed", fontface = "bold") +
  scale_color_manual(values = c(clrs$blue, clrs$red, clrs$orange), name = NULL) +
  theme_ngo() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank())

# Show and save plot
plot_dag
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.pdf"),
       width = 13/3, height = 2, units = "in", device = cairo_pdf)
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.png"),
       width = 13/3, height = 2, units = "in", type = "cairo", dpi = 300)
```

# Sample details

```{r}
vars_to_summarize <- tribble(
  ~category, ~variable, ~clean_name,
  "Demographics", "Q5.12", "Gender",
  "Demographics", "Q5.17", "Age",
  "Demographics", "Q5.13", "Marital status",
  "Demographics", "Q5.14", "Education",
  "Demographics", "Q5.15", "Income",
  "Attitudes toward charity", "Q2.5", "Frequency of donating to charity",
  "Attitudes toward charity", "Q2.6", "Amount of donations to charity last year",
  "Attitudes toward charity", "Q2.7_f", "Importance of trusting charities",
  "Attitudes toward charity", "Q2.8_f", "Level of trust in charities",
  "Attitudes toward charity", "Q2.10", "Frequency of volunteering",
  "Politics, ideology, and religion", "Q2.1", "Frequency of following national news",
  "Politics, ideology, and religion", "Q5.7", "Traveled to a developing country",
  "Politics, ideology, and religion", "Q5.1", "Voted in last election",
  "Politics, ideology, and religion", "Q5.6_f", "Trust in political institutions and the state",
  "Politics, ideology, and religion", "Q5.2_f", "Political ideology",
  "Politics, ideology, and religion", "Q5.4", "Involvement in activist causes",
  "Politics, ideology, and religion", "Q5.8", "Frequency of attending religious services",
  "Politics, ideology, and religion", "Q5.9", "Importance of religion"
)

summarize_factor <- function(x) {
  output <- table(x) %>% 
    as_tibble() %>% 
    magrittr::set_colnames(., c("level", "count")) %>% 
    mutate(level = factor(level, levels = levels(x), ordered = TRUE)) %>%
    mutate(prop = count / sum(count),
           nice_prop = scales::percent(prop))

  return(list(output))
}

participant_summary <- results %>% 
  select(one_of(vars_to_summarize$variable)) %>% 
  summarize_all(summarize_factor) %>% 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "details") %>% 
  left_join(vars_to_summarize, by = "variable") %>% 
  unnest(details) %>% 
  mutate(level = as.character(level)) %>% 
  mutate(level = case_when(
    variable == "Q2.7_f" & level == "1" ~ "1 (not important)",
    variable == "Q2.7_f" & level == "7" ~ "7 (important)",
    variable == "Q2.8_f" & level == "1" ~ "1 (no trust)",
    variable == "Q2.8_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.6_f" & level == "1" ~ "1 (no trust)",
    variable == "Q5.6_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.2_f" & level == "1" ~ "1 (extremely liberal)",
    variable == "Q5.2_f" & level == "7" ~ "7 (extremely conservative)",
    variable == "Q5.15" & level == "Less than median" ~ "Less than 2017 national median ($61,372)",
    variable == "Q5.17" & level == "Less than median" ~ "Less than 2017 national median (36)",
    TRUE ~ level
  )) %>% 
  mutate(clean_name_shrunk = ifelse(clean_name == lag(clean_name), "", clean_name),
         clean_name_shrunk = ifelse(is.na(clean_name_shrunk), 
                                    clean_name[1], 
                                    clean_name_shrunk),
         category_shrunk = ifelse(category == lag(category), "", category),
         category_shrunk = ifelse(is.na(category_shrunk), 
                                    category[1], 
                                    category_shrunk))
```

```{r results="asis"}
participant_summary %>% 
  select(" " = category_shrunk, "  " = clean_name_shrunk, 
         "Response" = level, "N" = count, "%" = nice_prop) %>% 
  pandoc.table.return(caption = 'Summary of individual respondent characteristics {#tbl:sample-details}',
                      justify = "lllcc") %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-sample-details.md"))
```

# Results

## Plot marginals

```{r plot-marginals}
# Plot marginals
if (intercept == 1) {
  draws %>%
    ggplot(aes(x = Gamma, y = levels_long)) +
    stat_halfeye(.width = .95) +
    geom_vline(xintercept = 0, color = "grey") +
    labs(y = "Attribute Levels")
}
if (intercept != 1) {
  draws %>%
    filter(j != 1) %>%
    ggplot(aes(x = Gamma, y = covariates_long)) +
    stat_halfeye(.width = .95) +
    geom_vline(xintercept = 0, color = "white") +
    labs(y = "Covariates") +
    facet_wrap(~ levels_long, ncol = 5)
}

# Save plot
if (intercept == 1) file_name <- paste("marginal-posteriors_intercept.png")
if (intercept == 1 & interactions == 1) file_name <- paste("marginal-posteriors_interactions.png")
if (public_affairs == 1) file_name <- paste("marginal-posteriors_public-affairs.png")
if (political_ideology == 1) file_name <- paste("marginal-posteriors_political-ideology.png")
if (social_views == 1) file_name <- paste("marginal-posteriors_social-views.png")
if (charity_voluntarism == 1) file_name <- paste("marginal-posteriors_charity-voluntarism.png")
if (demographics == 1) file_name <- paste("marginal-posteriors_demographics.png")
if (public_political == 1) file_name <- paste("marginal-posteriors_public-political.png")
if (public_political_social == 1) file_name <- paste("marginal-posteriors_public-political-social.png")
if (public_political_social_charity == 1) file_name <- paste("marginal-posteriors_public-political-social-charity.png")
if (public_political_social_charity_demo == 1) file_name <- paste("marginal-posteriors_public-political-social-charity-demos.png")
if (intercept == 1) width_length <- c(10, 10)
if (intercept != 1) width_length <- c(20, 20)
if (public_political_social_charity == 1) width_length <- c(20, 30)
if (public_political_social_charity_demo == 1) width_length <- c(20, 30)
ggsave(
  file_name,
  path = here::here("analysis", "output", "figures"),
  width = width_length[1],
  height = width_length[2],
  units = "in"
)
```

## Plot heatmaps

Instead of plotting marginals, we can use a heatmap to identify where there are significant differences in the non-intercept models, which can then be narrowed down using the marginal plots.

```{r plot-heatmaps}
if (public_affairs == 1) plot_title <- paste("Public Affairs")
if (political_ideology == 1) plot_title <- paste("Political Ideology")
if (social_views == 1) plot_title <- paste("Social Views")
if (charity_voluntarism == 1) plot_title <- paste("Covariates related to charity and voluntarism")
if (demographics == 1) plot_title <- paste("Demographics")
if (public_political == 1) plot_title <- paste("Public Affairs and Political Ideology")
if (public_political_social == 1) plot_title <- paste("Public Affairs, Political Ideology, and Social Views")
if (public_political_social_charity == 1) plot_title <- paste("Public Affairs, Political Ideology, Social Views, and Charity and Voluntarism")
if (public_political_social_charity_demo == 1) plot_title <- paste("Public Affairs, Political Ideology, Social Views, Charity and Voluntarism, and Demographics")

if (charity_voluntarism == 1) {
  j_labels <- tribble(
    ~j, ~j_label,
    2, "Importance of trusting charities", 
    3, "Trust charities", 
    4, "Frequency of donation", 
    5, "Amount of donation", 
    6, "Volunteered", 
    7, "Volunteer frequency", 
    8, "Personal involvement in activism", 
    9, "Family involvement in activism", 
    10, "Membership in church", 
    11, "Membership in sports", 
    12, "Membership in art, music, or educational organization", 
    13, "Membership in labor union", 
    14, "Membership in political party", 
    15, "Membership in envrionmental organization", 
    16, "Membership in professional association", 
    17, "Membership in humanitarian organization", 
    18, "Membership in consumer organization", 
    19, "Membership in other organization"
  ) %>% 
    mutate(j_label = fct_inorder(j_label))
}

if (intercept != 1) {
  # Posterior means.
  posterior_means <- draws %>% 
    filter(j != 1) %>% 
    left_join(j_labels, by = "j") %>% 
    left_join(level_lookup, by = "i") %>% 
    mutate(
      levels = factor(i, labels = level_labels), 
      covariates = factor(j, labels = 1:length(covariate_labels))
    ) %>% 
    group_by(category, attribute, j_label) %>% 
    summarize(
      mean = mean(Gamma),
      ci_lower = quantile(Gamma, .025),
      ci_upper = quantile(Gamma, .975),
      gt_zero = if_else(ci_lower > 0 & ci_upper > 0, 1, 0),
      lt_zero = if_else(ci_lower < 0 & ci_upper < 0, -1, 0)
    )

  # Heatmap of Gamma matrix.
  plot_heatmap <- posterior_means %>% 
    mutate(
      gt_lt_zero = gt_zero + lt_zero,
      Coefficients = factor(
        gt_lt_zero, 
        levels = -1:1,
        labels = c("Negative", "Not Significant", "Positive"))
    ) %>% 
    ggplot(aes(x = j_label, y = attribute, fill = Coefficients)) +
    geom_tile(color = "white", size = 0.25) +
    # scale_fill_brewer(palette = "Blues") +
    scale_fill_manual(values = c(clrs$orange, "grey80", clrs$olive)) +
    scale_x_discrete(expand = c(0.001, 0.001)) +
    scale_y_discrete(expand = c(0.001, 0.001)) +
    # guides(fill = guide_legend(override.aes = list(key.height = 0.5, key.width = 0.5))) +
    guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
    theme_ngo(base_size = 14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
          legend.position = "bottom",
          strip.placement = "outside") +
    labs(
      title = plot_title,
      x = "Covariates",
      y = "Attribute Levels"
    ) + 
    # facet_wrap(~ category, ncol = 1, scales = "free_y")
    facet_grid(vars(category), scales = "free_y", space = "free_y")
}

plot_heatmap
ggsave(plot_heatmap, filename = here("analysis", "output", "figures", "q2-heatmap.pdf"),
       width = 26/3, height = 6.5, units = "in", device = cairo_pdf)
ggsave(plot_heatmap, filename = here("analysis", "output", "figures", "q2-heatmap.png"),
       width = 26/3, height = 6.5, units = "in", type = "cairo", dpi = 300)

# Save plot
if (public_affairs == 1) file_name <- paste("public-affairs_coefficient-matrix.png")
if (political_ideology == 1) file_name <- paste("political-ideology_coefficient-matrix.png")
if (social_views == 1) file_name <- paste("social-views_coefficient-matrix.png")
if (charity_voluntarism == 1) file_name <- paste("charity-voluntarism_coefficient-matrix.png")
if (demographics == 1) file_name <- paste("demographics_coefficient-matrix.png")
if (public_political == 1) file_name <- paste("public-political_coefficient-matrix.png")
if (public_political_social == 1) file_name <- paste("public-political-social_coefficient-matrix.png")
if (public_political_social_charity == 1) file_name <- paste("public-political-social-charity_coefficient-matrix.png")
if (public_political_social_charity_demo == 1) file_name <- paste("public-political-social-charity-demos_coefficient-matrix.png")
ggsave(
  file_name,
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

## Plot contrasts

Another approach to address the hypotheses is to plot contrasts.

```{r}
#storing intercept draws for easier reference
draws_intercepts <- draws
```

### Hypothesis 1, Branding

> Donors will be more likely to donate to Oxfam and Red Cross compared to Amnesty International and Greenpeace [Mechanism: awareness of need and contentiousness of issue area]

```{r branding-contrasts}
# Pull the posterior draws for the organizations and compute contrasts
contrasts <- tibble(
  amnesty = draws %>% filter(level == "Amnesty International") %>% select(Gamma) %>% pull(),
  greenpeace = draws %>% filter(level == "Greenpeace") %>% select(Gamma) %>% pull(),
  oxfam = draws %>% filter(level == "Oxfam") %>% select(Gamma) %>% pull(),
  red_cross = draws %>% filter(level == "Red Cross") %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    oxfam_amnesty = oxfam - amnesty,
    oxfam_greenpeace = oxfam - greenpeace,
    red_cross_amnesty = red_cross - amnesty,
    red_cross_greenpeace = red_cross - greenpeace,
  )

# Plot the contrasts
contrasts %>%
  pivot_longer(oxfam_amnesty:red_cross_greenpeace, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis01.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors are indeed more likely to donate to Red Cross compared to Amnesty International and Greenpeace. However, compared to Amnesty International and Greenpeace, donors are less likely to donate to Oxfam. In fact, there is a negligible difference between donors likelihood to donate to Oxfam and Amnesty International.

### Hypothesis 2, Government Crackdown

#### Hypothesis 2a, Relationship with Host Government

> Donors will show increased willingness to donate to NGOs that are facing government crackdown or criticism [Mechanism: Governments wouldn't be cracking down on them if they didn't perceive a threat from them which means organizations implementing their missions effectively. This perception of efficacy leads to increased donations.]

```{r crackdown-contrasts-a}
# Pull the posterior draws for government crackdown
# Reference Level = Friendly relationship with government

contrasts <- tibble(
  under_crackdown = draws %>% filter(level == "Under crackdown") %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(level == "Criticized") %>% select(Gamma) %>% pull()
  )

# Plot the contrasts
contrasts %>%
  pivot_longer(under_crackdown:criticized, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis02a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Contrary to our hypothesis, donors showed decreased willingness to donate to NGOs that are facing government crackdown or criticism. This decreased likelihood to donate is also greater for NGOs under crackdown compared to those under criticism.

#### Hypothesis 2b, Organization and Relationship with Host Government

> Donors will show increased willingness to donate to Oxfam and Red Cross when they are facing government crackdown or criticism compared to when Amnesty or Greenpeace is facing crackdown.

```{r crackdown-contrasts-b}
# Overwrite draws with the appropriate interaction data

draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_2b.rds"))

# Pull the posterior draws for organization*government crackdown interaction
# Reference Level = Friendly relationship with government

contrasts_int <- tibble(
  amnesty = draws %>% filter(i == 1) %>% select(Gamma) %>% pull(),
  greenpeace = draws %>% filter(i == 2) %>% select(Gamma) %>% pull(),
  oxfam = draws %>% filter(i == 3) %>% select(Gamma) %>% pull(),
  red_cross = draws %>% filter(i == 4)%>% select(Gamma) %>% pull(),
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  accountability = draws %>% filter(i == 9) %>% select(Gamma) %>% pull(),
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  amnesty_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  greenpeace_criticized = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  oxfam_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  red_cross_criticized = draws %>% filter(i == 17)%>% select(Gamma) %>% pull(),
  amnesty_crackdown = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  greenpeace_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  oxfam_crackdown = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  red_cross_crackdown = draws %>% filter(i == 21)%>% select(Gamma) %>% pull()
) %>% 
  mutate(
    oxfam_amnesty_criticized = (oxfam + criticized + oxfam_criticized) - (amnesty + criticized + amnesty_criticized),

    oxfam_greenpeace_criticized = (oxfam + criticized + oxfam_criticized) - (greenpeace + criticized + greenpeace_criticized),

    red_cross_amnesty_criticized = (red_cross + criticized + red_cross_criticized) - (amnesty + criticized + amnesty_criticized),

    red_cross_greenpeace_criticized = (red_cross + criticized + red_cross_criticized) - (greenpeace + criticized + greenpeace_criticized),

    oxfam_amnesty_crackdown = (oxfam + crackdown + oxfam_crackdown) - (amnesty + crackdown +  amnesty_crackdown),

    oxfam_greenpeace_crackdown = (oxfam + crackdown + oxfam_crackdown) - (greenpeace + crackdown +  greenpeace_crackdown),

    red_cross_amnesty_crackdown = (red_cross + crackdown + red_cross_crackdown) - (amnesty + crackdown +  amnesty_crackdown),

    red_cross_greenpeace_crackdown = (red_cross + crackdown + red_cross_crackdown) - (greenpeace + crackdown + greenpeace_crackdown)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(oxfam_amnesty_criticized:red_cross_greenpeace_crackdown, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis02b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
 )
```

These data don't support our hypothesis; the Red Cross brand has an insignificant impact on likelihood to donate compared to both Greenpeace and Amnesty. While the contrasts for Oxfam are closer to being statistically significant, the effect is working in the other direction--Greenpeace and Amnesty both induce increased likelihood to donate on behalf of donors.

### Hypothesis 3, Issue Area

#### Hypothesis 3a, Issue Area

> Donors will show increased willingness to donate to NGOs working in less contentious issue areas (emergency response and refugee relief) over more contentious issue areas (environment and human rights)

```{r issue-contrasts-a}
# Pull the posterior draws for issue area
# Reference Level = Emergency Response
draws <- draws_intercepts
contrasts <- tibble(
  environment = draws %>% filter(level == "Environment") %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(level == "Human rights") %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(level == "Refugee relief") %>% select(Gamma) %>% pull()
  )

# Plot the contrasts
contrasts %>%
  pivot_longer(environment:refugee_relief, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis03a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors appear to be sensitive to NGO issue area when they are considering to donate. Emergency response was the most favored issue area by donors, and experienced the highest likelihood of receiving donations. While this partially matches the claim of the hypothesis, donors showed the least likelihood to donate to refugee relief NGOs, which was contrary to our expectations. It is worth noting that there is some overlap between the intervals of the issue areas, which would suggest that their effect isn't as significant as other factors.

#### Hypothesis 3b, Relationship with Host Government and Issue Area

> Donors will show increased willingness to donate to NGOs facing government crackdown/criticism working in less contentious issue areas (emergency response and refugee relief) over more contentious issue areas (environment and human rights) [Mechanisms: Perceptions of deservingness of NGOs dealing with emergency response and refugee relief. Donors are also more likely to donate to programs that are compatible with government preferences and have easily measurable outputs, which environment and human rights programs often lack. NGOs working on more contentious issue areas may be expelled or shut down, which would be a waste of donor resources, make it less likely that they donate to these groups.]

```{r issue-contrasts-b}
# Overwrite draws with the appropriate interaction data
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_3b.rds"))

# Pull the posterior draws for government crackdown*issue area interaction
# Reference Level = Friendly relationship with government
# Reference Level = Emergency Response

contrasts_int <- tibble(
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  criticized_environment = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  criticized_humanrights = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  criticized_refugee = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  crackdown_environment = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  crackdown_humanrights = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  crackdown_refugee = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
) %>% 
  mutate(
  criticized_refugee_environment = (criticized + refugee_relief + criticized_refugee) - (criticized + environment + criticized_environment),

  criticized_refugee_humanrights = (criticized + refugee_relief + criticized_refugee) - (criticized + human_rights + criticized_humanrights),

  crackdown_refugee_environment = (crackdown + refugee_relief + crackdown_refugee - crackdown + environment + crackdown_environment),

  crackdown_refugee_humanrights = (crackdown + refugee_relief + crackdown_refugee - crackdown + human_rights + crackdown_humanrights)
)

# Plot the contrasts
contrasts_int %>%
  pivot_longer(criticized_refugee_environment:crackdown_refugee_humanrights, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
 "contrasts_hypothesis03b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

The generated contrasts don't provides results which are statistically significant. 0 is found within all intervals.

### Hypothesis 4, Funding Sources

#### Hypothesis 4a, Funding Sources

> Donors will show increased willingness to donate to NGOs that are funded primarily by numerous small private donors compared to NGOs that are funded by a handful of wealthy private donors and government grants [Mechanism: Perception of efficacy - your contribution matters as a small donor. Government funding may also imply lack of independence of government which can reduce the efficiency of an organization.]

```{r funding-contrasts-a}
# Pull the posterior draws for funding sources
# Reference Level = Funded primarily by many small private donations
draws <- draws_intercepts
contrasts <- tibble(
  few_wealthy_private_donors = draws %>% filter(level == "Handful of wealthy private donors") %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(level == "Government grants") %>% select(Gamma) %>% pull()
  )

# Plot the contrasts
contrasts %>%
  pivot_longer(few_wealthy_private_donors:government_grants, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis04a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)

```

As stated in the hypothesis, donors show increased willingness to donate to NGOs funded by numerous small private donors, as opposed to those funded by a handful of wealthy private donors or government grants.

#### Hypothesis 4b, Relationship with Host Government and Funding Sources

> Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors

```{r funding-contrasts-b}
# Overwrite draws with the appropriate interaction data
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_4b.rds"))

# Pull the posterior draws for government crackdown*funding sources interaction
# Reference Level = Friendly relationship with government
# Reference Level = Funded primarily by many small private donations

contrasts_int <- tibble(
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  criticized_wealthydonors = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  criticized_grants = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  crackdown_wealthydonors = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  crackdown_grants = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
) %>% mutate(
  criticized_wealthydonors = (criticized + few_wealthy_private_donors + criticized_wealthydonors) - criticized,

  criticized_grants = (criticized + government_grants + criticized_grants) - criticized,

  crackdown_wealthydonors = (crackdown + few_wealthy_private_donors + crackdown_wealthydonors) - crackdown,

  crackdown_grants = (crackdown + government_grants + crackdown_grants) - crackdown
)

# Plot the contrasts
contrasts_int %>%
  pivot_longer(criticized_wealthydonors:crackdown_grants, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis04b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)

```

The marginal posteriors for these contrasts are strange, but still tell us that the interaction between funding sources and government relationship produce no significant effects.

#### Hypothesis 4c, Relationship with Host Government, Funding Sources, and Issue Area

> Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors and work in less contentious areas (emergency response and refugee relief)

```{r funding-contrasts-c}
# Overwrite draws with the appropriate interaction data
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_4c.rds"))

# Pull the posterior draws for issue area*government crackdown*funding sources interaction
# Reference Level = Funded primarily by many small private donations
# Reference Level = Friendly relationship with government

contrasts_int <- tibble(
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  emergency_crackdown_wealthydonors = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  emergency_crackdown_grants = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  environment_crackdown_wealthydonors = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  environment_crackdown_grants = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  humanrights_crackdown_wealthydonors = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  humanrights_crackdown_grants = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  refugee_crackdown_wealthydonors = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  refugee_crackdown_grants = draws %>% filter(i == 21) %>% select(Gamma) %>% pull(),
) %>% 
  mutate(
  emergency_humanrights_wealthydonors = (crackdown + emergency_crackdown_wealthydonors) - (crackdown + few_wealthy_private_donors + human_rights + humanrights_crackdown_wealthydonors),

  emergency_humanrights_grants = (crackdown + emergency_crackdown_grants) - (crackdown + government_grants + human_rights + humanrights_crackdown_grants),

  emergency_environment_wealthydonors = (crackdown + emergency_crackdown_wealthydonors) - (crackdown + few_wealthy_private_donors + environment + environment_crackdown_wealthydonors),

  emergency_environment_grants = (crackdown + emergency_crackdown_grants) - (crackdown + government_grants + environment + environment_crackdown_wealthydonors),

  refugee_humanrights_wealthydonors = (crackdown + refugee_relief + refugee_crackdown_wealthydonors) - (crackdown + few_wealthy_private_donors + human_rights + humanrights_crackdown_wealthydonors),

  refugee_humanrights_grants = (crackdown + refugee_relief + refugee_crackdown_grants) - (crackdown + government_grants +  human_rights + humanrights_crackdown_grants),

  refugee_environment_wealthydonors = (crackdown + refugee_relief + refugee_crackdown_wealthydonors) - (crackdown + few_wealthy_private_donors + environment + environment_crackdown_wealthydonors),

  refugee_environment_grants = (crackdown + refugee_relief + refugee_crackdown_grants) - (crackdown + government_grants + environment + environment_crackdown_grants),

  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(emergency_humanrights_wealthydonors:refugee_environment_grants, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis04c.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)

```

Once again, we see some interesting distributions, but no results significant enough to confirm our hypothesis.

### Hypothesis 5, Organizational Practices

#### Hypothesis 5a, Organizational Practices

> Donors will show increased willingness to donate to NGOs that are financially transparent [Mechanism: Perception of efficacy]

```{r practices-contrasts-a}
# Pull the posterior draws for organizational practices
draws <- draws_intercepts
contrasts <- tibble(
  financial_transparency = draws %>% filter(level == "Financial transparency") %>% select(Gamma) %>% pull(),
  accountability = draws %>% filter(level == "Accountability") %>% select(Gamma) %>% pull()
)

# Plot the contrasts
contrasts %>%
  pivot_longer(financial_transparency, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)

```

Donors show an increased likelihood to donate to NGOs that are financially transparent, as hypothesized.

#### Hypothesis 5b, Relationship with Host Government, Organizational Practices

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent

```{r practices-contrasts-b}
# Overwrite draws with the appropriate interaction data
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5b.rds"))

# Pull the posterior draws for organizational practices*government crackdown interaction
# Reference Level = Friendly relationship with government

contrasts_int <- tibble(
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  financial_transparency_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  financial_transparency_crackdown = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  ) %>% mutate(
    financial_transparency_criticized = fin_transparency + criticized + financial_transparency_criticized,
    financial_transparency_crackdown = fin_transparency + crackdown + financial_transparency_crackdown
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(financial_transparency_criticized:financial_transparency_crackdown, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)

```

Donors show a similar increase in willingness to donate to accountable NGOs in comparison to NGOs that are financially transparent. However, this is contrary to what was hypothesized--we predicted that donor's willingness to donate wouldn't be affected by accountability practices.

#### Hypothesis 5c, Relationship with Host Government, Organizational Practices, and Funding Sources

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and are funded primarily by numerous small private donors

```{r practices-contrasts-c}
# Overwrite draws with the appropriate interaction data
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5c.rds"))

# Pull the posterior draws for organizational practices*funding sources*government crackdown interaction
# Reference Level = Friendly relationship with government
# Reference Level = Funded primarily by many small private donations
# All interactions include financial transparency

contrasts_int <- tibble(
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  fin_smalldonors_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  fin_wealthydonors_criticized = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  fin_grants_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  fin_smalldonors_crackdown = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  fin_wealthydonors_crackdown = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  fin_grants_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  ) %>%
  mutate(
  fin_smalldonors_wealthydonors_criticized = (fin_transparency + criticized + fin_smalldonors_criticized) - (fin_transparency + criticized + few_wealthy_private_donors + fin_wealthydonors_criticized),

  fin_smalldonors_grants_criticized = (fin_transparency + criticized + fin_smalldonors_criticized) -  (fin_transparency + criticized + government_grants + fin_grants_criticized),

  fin_smalldonors_wealthydonors_crackdown = (fin_transparency + crackdown + fin_smalldonors_crackdown) - (fin_transparency + crackdown + few_wealthy_private_donors + fin_wealthydonors_crackdown),

  fin_smalldonors_grants_crackdown = (fin_transparency + crackdown + fin_smalldonors_crackdown) - (fin_transparency + crackdown + government_grants + fin_grants_crackdown)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(fin_smalldonors_wealthydonors_criticized:fin_smalldonors_grants_crackdown, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05c.png",#
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

These intervals are massive; 0 is well within them, leaving us with insufficient evidence to confirm the hypothesis.

#### Hypothesis 5d, Relationship with Host Government, Organizational Practices, and Issue Area

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief)

```{r practices-contrasts-d}
# Overwrite draws with the appropriate interaction data
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5d.rds"))

# Pull the posterior draws for organizational practices*issue area*government crackdown interaction
# Reference Level = Friendly relationship with government
# Reference Level = Emergency Response
# All interactions include financial transparency

contrasts_int <- tibble(
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  fin_emergency_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  fin_emergency_crackdown = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  fin_environment_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  fin_environment_crackdown = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  fin_humanrights_criticized = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  fin_humanrights_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  fin_refugee_criticized = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  fin_refugee_crackdown = draws %>% filter(i == 21) %>% select(Gamma) %>% pull()
  ) %>%
  mutate(
  fin_emergency_environment_criticized = (fin_transparency + criticized + fin_emergency_criticized) - (fin_transparency + criticized + environment + fin_environment_criticized),

  fin_emergency_humanrights_criticized = (fin_transparency + criticized + fin_emergency_criticized) - fin_transparency + criticized + human_rights + fin_humanrights_criticized,

  fin_emergency_environment_crackdown = (fin_transparency + crackdown + fin_emergency_crackdown) - (fin_transparency + crackdown + environment + fin_environment_crackdown),

  fin_emergency_humanright_crackdown = (fin_transparency + crackdown + fin_emergency_crackdown) - (fin_transparency + crackdown + human_rights + fin_humanrights_crackdown),

  fin_refugee_environment_criticized = (fin_transparency + criticized + fin_refugee_criticized) - (fin_transparency + criticized + environment + fin_environment_criticized),

  fin_refugee_humanright_criticized = (fin_transparency + criticized + fin_refugee_criticized) - (fin_transparency + criticized + human_rights + fin_humanrights_criticized),

  fin_refugee_environment_crackdown = (fin_transparency + crackdown + fin_refugee_crackdown) - (fin_transparency + crackdown + environment + fin_environment_crackdown),

  fin_refugee_humanright_crackdown = (fin_transparency + crackdown + fin_refugee_crackdown) - (fin_transparency + crackdown + human_rights + fin_humanrights_crackdown)
)

# Plot the contrasts
contrasts_int %>%
  pivot_longer(fin_emergency_environment_criticized:fin_refugee_humanright_crackdown, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05d.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

The contrasts don't support the hypothesis, however, there is an interesting difference between emergency_environment_crackdown and emergency_environment_criticized.

#### Hypothesis 5e, Relationship with Host Government, Organizational Practices, Issue Area, and Funding Sources

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief) and are funded by numerous small donors

```{r practices-contrasts-e}
# Overwrite draws with the appropriate interaction data
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5e.rds"))

# Pull the posterior draws for organizational practices*funding*issue area*government crackdown interaction
# Reference Level = Friendly relationship with government
# Reference Level = Emergency Response
# All interactions include small donors
# All interactions include financial transparency

contrasts_int <- tibble(
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  fin_sd_emergency_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  fin_sd_emergency_crackdown = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  fin_sd_environment_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  fin_sd_environment_crackdown = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  fin_sd_humanrights_criticized = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  fin_sd_humanrights_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  fin_sd_refugee_criticized = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  fin_sd_refugee_crackdown = draws %>% filter(i == 21) %>% select(Gamma) %>% pull()
  ) %>% mutate(
  fin_sd_emergency_environment_criticized = (fin_transparency + criticized + fin_sd_emergency_criticized) - (fin_transparency + criticized + environment +  fin_sd_environment_criticized),

  fin_sd_emergency_humanrights_criticized = (fin_transparency + criticized + fin_sd_emergency_criticized) - (fin_transparency + criticized + human_rights +  fin_sd_humanrights_criticized),

  fin_sd_emergency_environment_crackdown = (fin_transparency + crackdown + fin_sd_emergency_crackdown) - (fin_transparency + crackdown + environment +  fin_sd_environment_crackdown),

  fin_sd_emergency_humanright_crackdown = (fin_transparency + crackdown + fin_sd_emergency_crackdown) - (fin_transparency + crackdown + human_rights + fin_sd_humanrights_crackdown),

  fin_sd_refugee_environment_criticized = (fin_transparency + criticized + refugee_relief +  fin_sd_refugee_criticized) - (fin_transparency + criticized + environment +  fin_sd_environment_criticized),

  fin_sd_refugee_humanright_criticized = (fin_transparency + criticized + refugee_relief +  fin_sd_refugee_criticized) - (fin_transparency + criticized + human_rights +  fin_sd_humanrights_criticized),

  fin_sd_refugee_environment_crackdown = (fin_transparency + crackdown + refugee_relief + fin_sd_refugee_crackdown) - (fin_transparency + crackdown + environment + fin_sd_environment_crackdown),

  fin_sd_refugee_humanright_crackdown = (fin_transparency + crackdown + refugee_relief + fin_sd_refugee_crackdown) - (fin_transparency + crackdown + human_rights + fin_sd_humanrights_crackdown
))

# Plot the contrasts
contrasts_int %>%
  pivot_longer(fin_sd_emergency_environment_criticized:fin_sd_refugee_humanright_crackdown, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05e.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

These marginal posteriors don't support our hypothesis, but again show an interesting difference between emergency_environment_criticized and emergency_environment_crackdown.

#### Hypothesis 5f, Organizational Practices

> Donors should be no more or less likely to donate to NGOs that are accountable and hold regular third party audits [Mechanism: Donors don't necessarily seek assurance through third-party programs/audits and charity watchdogs, but rather through word of mouth, personal scrutiny and local networks]

```{r practices-contrasts-b}
# Pull the posterior draws for organizational practices
draws <- draws_intercepts
contrasts <- tibble(
  financial_transparency = draws %>% filter(level == "Financial transparency") %>% select(Gamma) %>% pull(),
  accountability = draws %>% filter(level == "Accountability") %>% select(Gamma) %>% pull()
)

# Plot the contrasts
contrasts %>%
  pivot_longer(accountability, names_to = "level", values_to = "Gamma") %>% 
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Difference in Attribute Levels", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05f.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors show a similar increase in willingness to donate to accountable NGOs in comparison to NGOs that are financially transparent. However, this is contrary to what was hypothesized--we predicted that donor's willingness to donate wouldn't be affected by accountability practices.

## Model fit table

```{r model-fit}
# Load model fit table
read_rds(model_fit, here::here("analysis", "output", "model_fit.rds"))
```

The intercept-only model works best in terms of PSIS-LOO, followed by models with political ideology, public affairs and demographics covariates. While PSIS-LOO doesn't properly account for the specific benefits of a hierarchical model, it suggests that political ideology above all other covariate sets specified, has the most explanatory power.

## Baseline probabilities

```{r probs-baseline, fig.width=4.5, fig.height=4}
orgs <- draws %>% 
  filter(category == "Organization") %>% 
  mutate(prob = plogis(Theta))

fig_baseline <- ggplot(orgs, aes(x = prob, y = attribute)) +
  geom_halfeyeh(.width = 0.95, aes(fill = attribute)) +
  scale_fill_manual(values = c(clrs$yellow, clrs$olive, clrs$lime, clrs$red),
                    guide = FALSE) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Baseline probability of selecting organization", y = NULL) +
  theme_ngo()

fig_baseline
ggsave(fig_baseline, filename = here("analysis", "output", "figures", "probs-org.pdf"),
       width = 13/3, height = 3, units = "in", device = cairo_pdf)
ggsave(fig_baseline, filename = here("analysis", "output", "figures", "probs-org.png"),
       width = 13/3, height = 3, units = "in", type = "cairo", dpi = 300)
```

```{r}
orgs %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(tidy = data %>% map(~ median_hdci(.$prob))) %>% 
  unnest(tidy)
```

## Coefficients and marginal effects

```{r first-hypotheses-coefficients, fig.width=26/3, fig.height=5.5}
diff_refugee_hr <- diff_cols("Refugee relief", "Human rights")
diff_hr_env <- diff_cols("Human rights", "Environment")
diff_env_refugee <- diff_cols("Environment", "Refugee relief")

plot_issue <- ggplot(filter(draws, category == "Issue area"), 
                     aes(x = exp(Theta), y = attribute)) + 
  stat_halfeyeh(.width = 0.95, fill = clrs$blue) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.7, ymin = 2.1, ymax = 2.9, size = 0.25) +
  # annotate(geom = "label", x = 0.7, y = 2.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_refugee_hr$phrase) +
  # annotate(geom = "linerange", x = 0.8, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.8, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_hr_env$phrase) +
  # annotate(geom = "linerange", x = 0.585, ymin = 1.1, ymax = 2.9, size = 0.25) +
  # annotate(geom = "label", x = 0.585, y = 2.05, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_env_refugee$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Issue area",
       subtitle = "Reference group = emergency response") +
  theme_ngo()

diff_govt_richpeople <- diff_cols("Government grants", 
                                  "Handful of wealthy private donors")

plot_funding <- ggplot(filter(draws, category == "Funding sources"), 
                       aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$maroon) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.825, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.825, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_govt_richpeople$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Funding sources",
       subtitle = "Reference group = many small private donations") +
  theme_ngo()

diff_account_trans <- diff_cols("Accountability", "Financial transparency")

plot_org <- ggplot(filter(draws, category == "Organizational practices"), 
                   aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$orange) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 1.85, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 1.85, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_account_trans$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Organizational practices",
       subtitle = "Reference groups = no accountability; no transparency") +
  theme_ngo()

diff_govt <- diff_cols("Under crackdown", "Criticized")

plot_government <- ggplot(filter(draws, category == "Relationship with host government"), 
                          aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$teal) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.75, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.75, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_govt$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Relationship with host government",
       subtitle = "Reference group = friendly relationship with government") +
  theme_ngo()

plot_first_hypotheses <- (plot_org + plot_issue + plot_funding + plot_government)

plot_first_hypotheses
ggsave(plot_first_hypotheses, filename = here("analysis", "output", "figures", "results-h1-4.pdf"),
       width = 26/3, height = 5.5, units = "in", device = cairo_pdf)
ggsave(plot_first_hypotheses, filename = here("analysis", "output", "figures", "results-h1-4.png"),
       width = 26/3, height = 5.5, units = "in", type = "cairo", dpi = 300)
```

```{r}
draws %>% 
  filter(category != "Organization") %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(tidy = data %>% map(~ median_hdci(.$Theta))) %>% 
  unnest(tidy) %>% 
  mutate(exp_y = exp(y))
```

## Differences in means

According to Kruschke 8.4, we can look at the posterior distribution of the difference in thetas, then calculate the proportion of those differences that are greater/less than 0

$$
\operatorname{mean}(\theta_\text{Government grants} - \theta_\text{Private donors}) = 
$$

$$
\Delta \theta = 4; p(\Delta > 0) = 5
$$

```{r eval=FALSE}
col1 <- "Government grants"
col2 <- "Handful of wealthy private donors"

posterior_diffs <- draws %>% 
  filter(attribute %in% c({{col1}}, {{col2}})) %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]],
         diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

mean(posterior_diffs$diff_theta_e)

asdf <- draws %>% 
  filter(category == "Funding sources") %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]]) %>% 
  mutate(diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  summarize(bloop = exp(mean(Theta)),
            bloop1 = mean(exp(Theta)))

mean(asdf$diff_theta_e)
tidybayes::hdi(asdf$diff_theta_e)

mean(asdf$diff_theta > 0)
mean(asdf$diff_theta_e > 0)

asdf <- draws %>% 
  filter(category == "Organizational practices") %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]]) %>% 
  mutate(diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

mean(asdf$diff_theta > 0)
mean(asdf$diff_theta_e > 0)

ggplot(asdf, aes(x = diff_theta_e, y = 1)) +
  geom_halfeyeh(.width = 0.95)

draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  summarize(bloop = mean(Theta),
            bloop1 = exp(mean(Theta)))

qwer <- draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(hdi = data %>% map(~ exp(tidybayes::hdi(.$Theta, .width = 0.95))))
asdf %>% 
  unnest(hdi)
mean(asdf$Theta)
tidybayes::hdi(asdf$Theta, .width = 0.95)
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=100}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>