### 2021-12-14 analysis/04_figs-tables.Rmd
Finished remaining contrasts through 5f

---
title: "Figures and tables for paper"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%F')`"
output: 
  html_document:
    code_folding: hide
    pandoc_args:
      - "--default-image-extension=png"
editor_options: 
  chunk_output_type: console
---

# Load packages, data, and draws

```{r setup, warning=FALSE, message=FALSE}
library(tidyverse)
library(magrittr)
library(tidybayes)
library(broom)
library(ggridges)
library(ggraph)
library(ggdag)
library(pander)
library(patchwork)
library(scales)
library(here)

# Load custom plot functions
source(here("R", "graphics.R"))

# General settings
source(here("analysis", "options.R"))

# Make all the randomness reproducible
set.seed(1234)

# Load data
results <- read_rds(here("data", "derived_data", "final_data.rds"))

# Load draws
intercept <- 1                            # Intercept-only
interactions <- 0                         # Include interactions with the intercept-only model
public_affairs <- 0                       # Public affairs
political_ideology <- 0                   # Political ideology
social_views <- 0                         # Social views
charity_voluntarism <- 0                  # Charity and voluntarism
demographics <- 0                         # Demographics
public_political <- 0                     # Public affairs + Political ideology
public_political_social <- 0              # + Social views
public_political_social_charity <- 0      # + Charity and voluntarism
public_political_social_charity_demo <- 0 # + Demographics

if (intercept == 1 & interactions == 0) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "intercept.rds")
  ) %>% 
    mutate(model = "Intercept only")
}
if (intercept == 1 & interactions == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "interactions.rds")
  ) %>% 
    mutate(model = "Intercept and interactions")
}
if (public_affairs == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "public_affairs.rds")
  ) %>% 
    mutate(model = "Public affairs")
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7 # Public affairs
    )
}
if (political_ideology == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "political_ideology.rds")
  ) %>% 
    mutate(model = "Political ideology")
  covariates <- results %>%
    select(
      Q5.2 # Political ideology
    )
}
if (social_views == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "social_views.rds")
  ) %>% 
    mutate(model = "Social views")
  covariates <- results %>%
    select(
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10 # Social views
    )
}
if (charity_voluntarism == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "charity_voluntarism.rds")
  ) %>% 
    mutate(model = "Charity and voluntarism")
  covariates <- results %>%
    select(
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10 # Charity and voluntarism
    )
}
if (demographics == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "demographics.rds")
  ) %>% 
    mutate(model = "Demographics")
  covariates <- results %>%
    select(
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17 # Demographics
    )
}
if (public_political == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "public_political.rds")
  ) %>% 
    mutate(model = "Public affairs and political ideology")
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2                                   # Political ideology
    )
}
if (public_political_social == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "public_political_social.rds")
  ) %>% 
    mutate(model = "Public affairs, political ideology, and social views")
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7, # Public affairs
      Q5.2,                                  # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10         # Social views
    )
}
if (public_political_social_charity == 1) {
  draws <- read_rds(
    here::here("data", "raw_data", "posterior_draws", "public_political_social_charity.rds")
  ) %>% 
    mutate(
      model = "Public affairs, political ideology, social views, charity and voluntarism"
    )
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10  # Charity and voluntarism
    )
}
if (public_political_social_charity_demo == 1) {
  draws <- read_rds(
    here::here(
      "analysis", 
      "output", 
      "posterior_draws", 
      "public_political_social_charity_demo.rds"
    )
  ) %>% 
    mutate(
      model = "Public affairs, 
      political ideology, 
      social views, 
      charity and voluntarism, 
      and demographics"
    )
  covariates <- results %>%
    select(
      Q2.1, Q2.2, Q2.3_1:Q2.3_7, Q2.4, Q5.7,                           # Public affairs
      Q5.2,                                                            # Political ideology
      Q5.6, Q5.11, Q5.8, Q5.9, Q5.10,                                  # Social views
      Q2.7, Q2.8, Q2.5, Q2.6, Q2.9, Q2.10, Q5.4, Q5.5, Q5.3_1:Q5.3_10, # Charity and voluntarism
      Q5.12, Q5.13, Q5.14, Q5.15, Q5.16_1:Q5.16_6, Q5.17               # Demographics
    )
}

# Name levels labels
level_labels <- c(
  str_c("Organization: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross")),
  str_c("Issue area: ", c("Environment", "Human rights", "Refugee relief")),
  str_c("Organizational practices: ", c("Financial transparency", "Accountability")),
  str_c("Funding sources: ", c("Handful of wealthy private donors", "Government grants")),
  str_c("Relationship with host government: ", c("Criticized", "Under crackdown"))
)

if (intercept == 1 & interactions == 1) {
  interaction_labels <- c(
    # Hypothesis 2b.
    str_c("Interaction: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"), " Criticized"),
    str_c("Interaction: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross"), " Crackdown"),
    # Hypothesis 3b.
    str_c("Interaction: ", c("Environment", "Human rights", "Refugee relief"), " Criticized"),
    str_c("Interaction: ", c("Environment", "Human rights", "Refugee relief"), " Crackdown"),
    # Hypothesis 4b.
    str_c("Interaction: ", c("Handful of wealthy private donors", "Government grants"), " Criticized"),
    str_c("Interaction: ", c("Handful of wealthy private donors", "Government grants"), " Crackdown"),
    # Hypothesis 4c.
    str_c("Interaction: Emergency ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    str_c("Interaction: Environment ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    str_c("Interaction: Human rights ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    str_c("Interaction: Refugee relief ", c("Wealthy private donors", "Government grants"), " Crackdown"),
    # Hypothesis 5b.
    str_c("Interaction: Financial transparency", c(" Criticized", " Crackdown")),
    # Hypothesis 5c.
    str_c("Interaction: Financial transparency ", c("Small private donors", "Wealthy private donors", "Government grants"), " Criticized"),
    str_c("Interaction: Financial transparency ", c("Small private donors", "Wealthy private donors", "Government grants"), " Crackdown"),
    # Hypothesis 5d.
    str_c("Interaction: Financial transparency Emergency", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Environment", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Human rights", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Refugee relief", c(" Criticized", " Crackdown")),
    # Hypothesis 5e.
    str_c("Interaction: Financial transparency Small private donoros Emergency", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Small private donoros Environment", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Small private donoros Human rights", c(" Criticized", " Crackdown")),
    str_c("Interaction: Financial transparency Small private donoros Refugee relief", c(" Criticized", " Crackdown"))
  )
  level_labels <- c(level_labels, interaction_labels)
}

level_lookup <- tibble(levels = level_labels) %>% 
  separate(levels, c("attribute", "level"), sep = ": ") %>% 
  mutate(levels_long = paste0(attribute, ": ", level),
         i = 1:n()) %>% 
  mutate_at(vars(attribute, level, levels_long), list(fct_inorder))

if (intercept != 1) {
  # Extract covariate labels
  covariate_lookup <- covariates %>% 
    pivot_longer(cols = starts_with("Q"), names_to = "question", values_to = "labels") %>% 
    bind_cols(
      covariates %>% 
        mutate_all(as.integer) %>% 
        rename_all(paste0, "_temp") %>% 
        pivot_longer(cols = starts_with("Q"), names_to = "question_temp", values_to = "level")
    ) %>% 
    distinct() %>% 
    filter(level != 1) %>% 
    separate(labels, c("covariate", "label"), sep = ": ") %>% 
    mutate(covariates_long = paste0(covariate, ": ", label), j = 2:(n() + 1)) %>% 
    select(covariate, label, covariates_long, j) %>% 
    bind_rows(
      tibble(
        covariate = "Intercept",
        label = "Intercept",
        covariates_long = "Intercept",
        j = 1
      )
    ) %>% 
    arrange(j) %>% 
    mutate_at(vars(covariate, label, covariates_long), list(fct_inorder))
}

# Join level and covariate labels to draws
draws <- draws %>% 
  left_join(level_lookup, by = "i")

if (intercept != 1) {
  draws <- draws %>% 
    left_join(covariate_lookup, by = "j")
}
```

```{r}
diff_cols <- function(col1, col2) {
  posterior_diffs <- draws %>% 
    filter(attribute %in% c({{col1}}, {{col2}})) %>% 
    select(.chain, .iteration, Theta, attribute) %>% 
    pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
    mutate(diff_theta = .[[4]] - .[[3]],
           diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

  diff_theta <- mean(posterior_diffs$diff_theta_e)
  p_greater0 <- mean(posterior_diffs$diff_theta_e > 0)
  p_less0 <- mean(posterior_diffs$diff_theta_e < 0)
  p_not0 <- ifelse(median(posterior_diffs$diff_theta_e) > 0, p_greater0, p_less0)

  phrase <- paste0("∆θ = ", round(diff_theta, 3), "; p(∆ ≠ 0) = ", round(p_not0, 3))

  return(list(posterior_diffs = posterior_diffs, diff_theta = diff_theta,
              p_greater0 = p_greater0, p_less0 = p_less0, p_not0 = p_not0,
              phrase = phrase))
}
```

# Causal pathway

Our theory and hypotheses are laid out in the causal pathway below. Organizational factors (**O**, or organizational practices like financial disclosure and accountability practices; **I**, or issue area; and **F**, or funding) all influence the decision to donate (**D**). Structural factors like an NGO's relationship with its host government (**G**) also influence donation behavior, but the relationship itself is shaped by both funding and issue area.

```{r causal-dag, fig.width=13/3, fig.height=2}
node_colors <- tribble(
  ~name, ~type,
  "D", "Outcome",
  "I", "Organizational factor",
  "F", "Organizational factor",
  "O", "Organizational factor",
  "G", "Structural factor"
) %>% 
  mutate(type = fct_inorder(type))

# Set up DAG structure
theory_dag <- dagify(D ~ I + F + O + G,
                     G ~ I + F,
                     outcome = "D",
                     exposure = "C") %>% 
  tidy_dagitty(layout = "dh", seed = 1234)

# Add node types/colors to DAG data
theory_dag$data <- theory_dag$data %>% 
  left_join(node_colors, by = "name")

# Make DAG plot
plot_dag <- ggplot(theory_dag, aes(x = x, y = y, xend = xend, yend = yend)) +
  geom_dag_point(size = 6, mapping = aes(color = type)) +
  geom_dag_edges(start_cap = circle(4, "mm"),
                 end_cap = circle(4, "mm")) +
  geom_dag_text(size = pts(6), family = "Roboto Condensed", fontface = "bold") +
  scale_color_manual(values = c(clrs$blue, clrs$red, clrs$orange), name = NULL) +
  theme_ngo() +
  theme(panel.grid = element_blank(),
        panel.border = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank())

# Show and save plot
plot_dag
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.pdf"),
       width = 13/3, height = 2, units = "in", device = cairo_pdf)
ggsave(plot_dag, filename = here("analysis", "output", "figures", "causal-path.png"),
       width = 13/3, height = 2, units = "in", type = "cairo", dpi = 300)
```

# Sample details

```{r}
vars_to_summarize <- tribble(
  ~category, ~variable, ~clean_name,
  "Demographics", "Q5.12", "Gender",
  "Demographics", "Q5.17", "Age",
  "Demographics", "Q5.13", "Marital status",
  "Demographics", "Q5.14", "Education",
  "Demographics", "Q5.15", "Income",
  "Attitudes toward charity", "Q2.5", "Frequency of donating to charity",
  "Attitudes toward charity", "Q2.6", "Amount of donations to charity last year",
  "Attitudes toward charity", "Q2.7_f", "Importance of trusting charities",
  "Attitudes toward charity", "Q2.8_f", "Level of trust in charities",
  "Attitudes toward charity", "Q2.10", "Frequency of volunteering",
  "Politics, ideology, and religion", "Q2.1", "Frequency of following national news",
  "Politics, ideology, and religion", "Q5.7", "Traveled to a developing country",
  "Politics, ideology, and religion", "Q5.1", "Voted in last election",
  "Politics, ideology, and religion", "Q5.6_f", "Trust in political institutions and the state",
  "Politics, ideology, and religion", "Q5.2_f", "Political ideology",
  "Politics, ideology, and religion", "Q5.4", "Involvement in activist causes",
  "Politics, ideology, and religion", "Q5.8", "Frequency of attending religious services",
  "Politics, ideology, and religion", "Q5.9", "Importance of religion"
)

summarize_factor <- function(x) {
  output <- table(x) %>% 
    as_tibble() %>% 
    magrittr::set_colnames(., c("level", "count")) %>% 
    mutate(level = factor(level, levels = levels(x), ordered = TRUE)) %>%
    mutate(prop = count / sum(count),
           nice_prop = scales::percent(prop))

  return(list(output))
}

participant_summary <- results %>% 
  select(one_of(vars_to_summarize$variable)) %>% 
  summarize_all(summarize_factor) %>% 
  pivot_longer(cols = everything(), names_to = "variable", values_to = "details") %>% 
  left_join(vars_to_summarize, by = "variable") %>% 
  unnest(details) %>% 
  mutate(level = as.character(level)) %>% 
  mutate(level = case_when(
    variable == "Q2.7_f" & level == "1" ~ "1 (not important)",
    variable == "Q2.7_f" & level == "7" ~ "7 (important)",
    variable == "Q2.8_f" & level == "1" ~ "1 (no trust)",
    variable == "Q2.8_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.6_f" & level == "1" ~ "1 (no trust)",
    variable == "Q5.6_f" & level == "7" ~ "7 (complete trust)",
    variable == "Q5.2_f" & level == "1" ~ "1 (extremely liberal)",
    variable == "Q5.2_f" & level == "7" ~ "7 (extremely conservative)",
    variable == "Q5.15" & level == "Less than median" ~ "Less than 2017 national median ($61,372)",
    variable == "Q5.17" & level == "Less than median" ~ "Less than 2017 national median (36)",
    TRUE ~ level
  )) %>% 
  mutate(clean_name_shrunk = ifelse(clean_name == lag(clean_name), "", clean_name),
         clean_name_shrunk = ifelse(is.na(clean_name_shrunk), 
                                    clean_name[1], 
                                    clean_name_shrunk),
         category_shrunk = ifelse(category == lag(category), "", category),
         category_shrunk = ifelse(is.na(category_shrunk), 
                                    category[1], 
                                    category_shrunk))
```

```{r results="asis"}
participant_summary %>% 
  select(" " = category_shrunk, "  " = clean_name_shrunk, 
         "Response" = level, "N" = count, "%" = nice_prop) %>% 
  pandoc.table.return(caption = 'Summary of individual respondent characteristics {#tbl:sample-details}',
                      justify = "lllcc") %T>%
  cat() %>%
  cat(file = here("analysis", "output", "tables", "tbl-sample-details.md"))
```

# Results

## Plot marginals

```{r plot-marginals}
# Plot marginals
if (intercept == 1) {
  draws %>%
    ggplot(aes(x = Gamma, y = levels_long)) +
    stat_halfeye(.width = .95) +
    geom_vline(xintercept = 0, color = "grey") +
    labs(y = "Attribute Levels")
}
if (intercept != 1) {
  draws %>%
    filter(j != 1) %>%
    ggplot(aes(x = Gamma, y = covariates_long)) +
    stat_halfeye(.width = .95) +
    geom_vline(xintercept = 0, color = "white") +
    labs(y = "Covariates") +
    facet_wrap(~ levels_long, ncol = 5)
}

# Save plot
if (intercept == 1) file_name <- paste("marginal-posteriors_intercept.png")
if (intercept == 1 & interactions == 1) file_name <- paste("marginal-posteriors_interactions.png")
if (public_affairs == 1) file_name <- paste("marginal-posteriors_public-affairs.png")
if (political_ideology == 1) file_name <- paste("marginal-posteriors_political-ideology.png")
if (social_views == 1) file_name <- paste("marginal-posteriors_social-views.png")
if (charity_voluntarism == 1) file_name <- paste("marginal-posteriors_charity-voluntarism.png")
if (demographics == 1) file_name <- paste("marginal-posteriors_demographics.png")
if (public_political == 1) file_name <- paste("marginal-posteriors_public-political.png")
if (public_political_social == 1) file_name <- paste("marginal-posteriors_public-political-social.png")
if (public_political_social_charity == 1) file_name <- paste("marginal-posteriors_public-political-social-charity.png")
if (public_political_social_charity_demo == 1) file_name <- paste("marginal-posteriors_public-political-social-charity-demos.png")
if (intercept == 1) width_length <- c(10, 10)
if (intercept != 1) width_length <- c(20, 20)
if (public_political_social_charity == 1) width_length <- c(20, 30)
if (public_political_social_charity_demo == 1) width_length <- c(20, 30)
ggsave(
  file_name,
  path = here::here("analysis", "output", "figures"),
  width = width_length[1],
  height = width_length[2],
  units = "in"
)
```

## Plot heatmaps

Instead of plotting marginals, we can use a heatmap to identify where there are significant differences in the non-intercept models, which can then be narrowed down using the marginal plots.

```{r plot-heatmaps}
if (public_affairs == 1) plot_title <- paste("Public Affairs")
if (political_ideology == 1) plot_title <- paste("Political Ideology")
if (social_views == 1) plot_title <- paste("Social Views")
if (charity_voluntarism == 1) plot_title <- paste("Covariates related to charity and voluntarism")
if (demographics == 1) plot_title <- paste("Demographics")
if (public_political == 1) plot_title <- paste("Public Affairs and Political Ideology")
if (public_political_social == 1) plot_title <- paste("Public Affairs, Political Ideology, and Social Views")
if (public_political_social_charity == 1) plot_title <- paste("Public Affairs, Political Ideology, Social Views, and Charity and Voluntarism")
if (public_political_social_charity_demo == 1) plot_title <- paste("Public Affairs, Political Ideology, Social Views, Charity and Voluntarism, and Demographics")

if (charity_voluntarism == 1) {
  j_labels <- tribble(
    ~j, ~j_label,
    2, "Importance of trusting charities", 
    3, "Trust charities", 
    4, "Frequency of donation", 
    5, "Amount of donation", 
    6, "Volunteered", 
    7, "Volunteer frequency", 
    8, "Personal involvement in activism", 
    9, "Family involvement in activism", 
    10, "Membership in church", 
    11, "Membership in sports", 
    12, "Membership in art, music, or educational organization", 
    13, "Membership in labor union", 
    14, "Membership in political party", 
    15, "Membership in envrionmental organization", 
    16, "Membership in professional association", 
    17, "Membership in humanitarian organization", 
    18, "Membership in consumer organization", 
    19, "Membership in other organization"
  ) %>% 
    mutate(j_label = fct_inorder(j_label))
}

if (intercept != 1) {
  # Posterior means.
  posterior_means <- draws %>% 
    filter(j != 1) %>% 
    left_join(j_labels, by = "j") %>% 
    left_join(level_lookup, by = "i") %>% 
    mutate(
      levels = factor(i, labels = level_labels), 
      covariates = factor(j, labels = 1:length(covariate_labels))
    ) %>% 
    group_by(category, attribute, j_label) %>% 
    summarize(
      mean = mean(Gamma),
      ci_lower = quantile(Gamma, .025),
      ci_upper = quantile(Gamma, .975),
      gt_zero = if_else(ci_lower > 0 & ci_upper > 0, 1, 0),
      lt_zero = if_else(ci_lower < 0 & ci_upper < 0, -1, 0)
    )

  # Heatmap of Gamma matrix.
  plot_heatmap <- posterior_means %>% 
    mutate(
      gt_lt_zero = gt_zero + lt_zero,
      Coefficients = factor(
        gt_lt_zero, 
        levels = -1:1,
        labels = c("Negative", "Not Significant", "Positive"))
    ) %>% 
    ggplot(aes(x = j_label, y = attribute, fill = Coefficients)) +
    geom_tile(color = "white", size = 0.25) +
    # scale_fill_brewer(palette = "Blues") +
    scale_fill_manual(values = c(clrs$orange, "grey80", clrs$olive)) +
    scale_x_discrete(expand = c(0.001, 0.001)) +
    scale_y_discrete(expand = c(0.001, 0.001)) +
    # guides(fill = guide_legend(override.aes = list(key.height = 0.5, key.width = 0.5))) +
    guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
    theme_ngo(base_size = 14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
          legend.position = "bottom",
          strip.placement = "outside") +
    labs(
      title = plot_title,
      x = "Covariates",
      y = "Attribute Levels"
    ) + 
    # facet_wrap(~ category, ncol = 1, scales = "free_y")
    facet_grid(vars(category), scales = "free_y", space = "free_y")
}

plot_heatmap
ggsave(plot_heatmap, filename = here("analysis", "output", "figures", "q2-heatmap.pdf"),
       width = 26/3, height = 6.5, units = "in", device = cairo_pdf)
ggsave(plot_heatmap, filename = here("analysis", "output", "figures", "q2-heatmap.png"),
       width = 26/3, height = 6.5, units = "in", type = "cairo", dpi = 300)

# Save plot
if (public_affairs == 1) file_name <- paste("public-affairs_coefficient-matrix.png")
if (political_ideology == 1) file_name <- paste("political-ideology_coefficient-matrix.png")
if (social_views == 1) file_name <- paste("social-views_coefficient-matrix.png")
if (charity_voluntarism == 1) file_name <- paste("charity-voluntarism_coefficient-matrix.png")
if (demographics == 1) file_name <- paste("demographics_coefficient-matrix.png")
if (public_political == 1) file_name <- paste("public-political_coefficient-matrix.png")
if (public_political_social == 1) file_name <- paste("public-political-social_coefficient-matrix.png")
if (public_political_social_charity == 1) file_name <- paste("public-political-social-charity_coefficient-matrix.png")
if (public_political_social_charity_demo == 1) file_name <- paste("public-political-social-charity-demos_coefficient-matrix.png")
ggsave(
  file_name,
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

## Plot contrasts

To formally answer the hypotheses, we compute contrasts of marginal posteriors. First, we need the intercept-only draws and associated levels.

```{r intercept-draws}
# Import draws
draws_intercept <- read_rds(here::here("data", "raw_data", "posterior_draws", "intercept.rds")) %>% 
  mutate(model = "Intercept only")

# Name levels labels
level_labels <- c(
  str_c("Organization: ", c("Amnesty International", "Greenpeace", "Oxfam", "Red Cross")),
  str_c("Issue area: ", c("Environment", "Human rights", "Refugee relief")),
  str_c("Organizational practices: ", c("Financial transparency", "Accountability")),
  str_c("Funding sources: ", c("Handful of wealthy private donors", "Government grants")),
  str_c("Relationship with host government: ", c("Criticized", "Under crackdown"))
)

level_lookup <- tibble(levels = level_labels) %>% 
  separate(levels, c("attribute", "level"), sep = ": ") %>% 
  mutate(levels_long = paste0(attribute, ": ", level),
         i = 1:n()) %>% 
  mutate_at(vars(attribute, level, levels_long), list(fct_inorder))

# Join levels and labels to draws
draws_intercept <- draws_intercept %>% 
  left_join(level_lookup, by = "i")
```

### Hypothesis 1, Branding

> Donors will be more likely to donate to Oxfam and Red Cross compared to Amnesty International and Greenpeace [Mechanism: awareness of need and contentiousness of issue area]

```{r contrasts-hypothesis01}
# Compute the contrasts
contrasts <- tibble(
  amnesty = draws_intercept %>% filter(level == "Amnesty International") %>% select(Gamma) %>% pull(),
  greenpeace = draws_intercept %>% filter(level == "Greenpeace") %>% select(Gamma) %>% pull(),
  oxfam = draws_intercept %>% filter(level == "Oxfam") %>% select(Gamma) %>% pull(),
  red_cross = draws_intercept %>% filter(level == "Red Cross") %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "oxfam - amnesty" = oxfam - amnesty,
    "oxfam - greenpeace" = oxfam - greenpeace,
    "red cross - amnesty" = red_cross - amnesty,
    "red cross - greenpeace" = red_cross - greenpeace
  )

# Plot the contrasts
contrasts %>%
  pivot_longer("oxfam - amnesty":"red cross - greenpeace", names_to = "level", values_to = "Gamma") %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis01.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors are indeed more likely to donate to Red Cross compared to Amnesty International and Greenpeace, holding every other attribute constant. However, compared to Amnesty International and Greenpeace, donors are less likely to donate to Oxfam. In fact, there is a negligible difference between donors likelihood to donate to Oxfam and Amnesty International. This might be because the population we sampled from was specifically in the United States, where Oxfam and Amnesty may be less well known.

### Hypothesis 2, Government Crackdown

#### Hypothesis 2a, Relationship with Host Government

> Donors will show increased willingness to donate to NGOs that are facing government crackdown or criticism [Mechanism: Governments wouldn't be cracking down on them if they didn't perceive a threat from them which means organizations implementing their missions effectively. This perception of efficacy leads to increased donations.]

```{r contrasts-hypothesis02a}
# Reference Level = Friendly relationship with government
# Compute the contrasts
contrasts <- tibble(
  "criticized - friendly" = draws_intercept %>% filter(level == "Criticized") %>% select(Gamma) %>% pull(),
  "crackdown - friendly" = draws_intercept %>% filter(level == "Under crackdown") %>% select(Gamma) %>% pull()
)

# Plot the contrasts
contrasts %>%
  pivot_longer("crackdown - friendly":"criticized - friendly", names_to = "level", values_to = "Gamma") %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis02a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Contrary to our hypothesis, donors showed decreased willingness to donate to NGOs that are facing government crackdown or criticism relative to NGOs that have a friendly relationship with the government, holding every other attribute constant, with this decreased likelihood to donate greater for NGOs under crackdown compared to those being criticized.

#### Hypothesis 2b, Organization and Relationship with Host Government

> Donors will show increased willingness to donate to Oxfam and Red Cross when they are facing government crackdown or criticism compared to when Amnesty or Greenpeace is facing crackdown.

```{r contrasts-hypothesis02b}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_2b.rds"))

# Reference Level = Friendly relationship with government
# Compute the contrasts
contrasts_int <- tibble(
  # Attribute levels
  amnesty = draws %>% filter(i == 1) %>% select(Gamma) %>% pull(),
  greenpeace = draws %>% filter(i == 2) %>% select(Gamma) %>% pull(),
  oxfam = draws %>% filter(i == 3) %>% select(Gamma) %>% pull(),
  red_cross = draws %>% filter(i == 4)%>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  # Hypothesis 2b
  amnesty_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  greenpeace_criticized = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  oxfam_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  red_cross_criticized = draws %>% filter(i == 17)%>% select(Gamma) %>% pull(),
  amnesty_crackdown = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  greenpeace_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  oxfam_crackdown = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  red_cross_crackdown = draws %>% filter(i == 21)%>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "criticized oxfam - criticized amnesty" = 
      (oxfam + criticized + oxfam_criticized) - (amnesty + criticized + amnesty_criticized),
    "criticized oxfam - criticized greenpeace" = 
      (oxfam + criticized + oxfam_criticized) - (greenpeace + criticized + greenpeace_criticized),
    "crackdown oxfam - crackdown amnesty" = 
      (oxfam + crackdown + oxfam_crackdown) - (amnesty + crackdown +  amnesty_crackdown),
    "crackdown oxfam - crackdown greenpeace" = 
      (oxfam + crackdown + oxfam_crackdown) - (greenpeace + crackdown +  greenpeace_crackdown),
    "criticized red cross - criticized amnesty" = 
      (red_cross + criticized + red_cross_criticized) - (amnesty + criticized + amnesty_criticized),
    "criticized red cross - criticized greenpeace" = 
      (red_cross + criticized + red_cross_criticized) - (greenpeace + criticized + greenpeace_criticized),
    "crackdown red cross - crackdown amnesty" = 
      (red_cross + crackdown + red_cross_crackdown) - (amnesty + crackdown +  amnesty_crackdown),
    "crackdown red cross - crackdown greenpeace" = 
      (red_cross + crackdown + red_cross_crackdown) - (greenpeace + crackdown + greenpeace_crackdown)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "criticized oxfam - criticized amnesty":"crackdown red cross - crackdown greenpeace", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis02b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
 )
```

Although donors show significantly more willingness to donate to Red Cross when they are criticized or under crackdown from the host government, donors are less likely to donate to Oxfam under the same conditions, holding every other attribute constant. The only contrast which isn't statistically significant is the difference between Oxfam and Amnesty when both are under crackdown.

### Hypothesis 3, Issue Area

#### Hypothesis 3a, Issue Area

> Donors will show increased willingness to donate to NGOs working in less contentious issue areas (emergency response and refugee relief) over more contentious issue areas (environment and human rights)

```{r contrasts-hypothesis03a}
# Reference Level = Emergency Response
# Compute the contrasts
contrasts <- tibble(
  "environment - emergency" = draws_intercept %>% filter(level == "Environment") %>% select(Gamma) %>% pull(),
  "human rights - emergency" = draws_intercept %>% filter(level == "Human rights") %>% select(Gamma) %>% pull(),
  "refugee relief - emergency" = draws_intercept %>% filter(level == "Refugee relief") %>% select(Gamma) %>% pull()
)

# Plot the contrasts
contrasts %>%
  pivot_longer(
    "environment - emergency":"refugee relief - emergency", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis03a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors appear to be sensitive to NGO issue area when they are considering to donate. Emergency response was the most favored issue area by donors, and experienced the highest likelihood of receiving donations, holding every other attribute constant. While this partially matches the claim of the hypothesis, donors showed the least likelihood to donate to refugee relief NGOs relative to emergency response, although this is not significantly different from their likelihood of donating for environmental causes with the same reference. In other words, refugee relief appears to be as contentious as environmental issues.

#### Hypothesis 3b, Relationship with Host Government and Issue Area

> Donors will show increased willingness to donate to NGOs facing government crackdown/criticism working in less contentious issue areas (emergency response and refugee relief) over more contentious issue areas (environment and human rights) [Mechanisms: Perceptions of deservingness of NGOs dealing with emergency response and refugee relief. Donors are also more likely to donate to programs that are compatible with government preferences and have easily measurable outputs, which environment and human rights programs often lack. NGOs working on more contentious issue areas may be expelled or shut down, which would be a waste of donor resources, make it less likely that they donate to these groups.]

```{r contrasts-hypothesis03b}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_3b.rds"))

# Reference Level = Emergency Response
# Reference Level = Friendly relationship with government
# Compute the contrasts
contrasts_int <- tibble(
  # Attribute levels
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  # Hypothesis 3b
  criticized_environment = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  criticized_human_rights = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  criticized_refugee = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  crackdown_environment = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  crackdown_human_rights = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  crackdown_refugee = draws %>% filter(i == 19) %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "environment - emergency" = environment,
    "human rights - emergency" = human_rights,
    "refugee relief - emergency" = refugee_relief,
    "criticized environment - criticized emergency" = 
      (criticized + environment + criticized_environment) - (criticized),
    "criticized human rights - criticized emergency" = 
      (criticized + human_rights + criticized_human_rights) - (criticized),
    "criticized refugee relief - criticized emergency" = 
      (criticized + refugee_relief + criticized_refugee) - (criticized),
    "crackdown environment - crackdown emergency" = 
      (crackdown + environment + crackdown_environment) - (crackdown),
    "crackdown human rights - crackdown emergency" = 
      (crackdown + human_rights + crackdown_human_rights) - (crackdown),
    "crackdown refugee relief - crackdown emergency" = 
      (crackdown + human_rights + crackdown_human_rights) - (crackdown)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "environment - emergency":"crackdown refugee relief - crackdown emergency", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
 "contrasts_hypothesis03b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

When NGOs are criticized by the government, we see the same pattern as before, with emergency response being most preferred followed by human rights, environment, and refugee relief, holding every other attribute constant. However, we do not see an increase in likelihood to donate relative to an emergency response NGO being criticized. However, when NGOs are under crackdown by the government, we see that human rights, environment, and refugee relief all increase in their likelihood relative to an emergency response NGO under crackdown. In fact, both human rights and refugee relief NGOs under crackdown are indistinguishable from an emergency response NGO under crackdown in terms of propensity to donate.

### Hypothesis 4, Funding Sources

#### Hypothesis 4a, Funding Sources

> Donors will show increased willingness to donate to NGOs that are funded primarily by numerous small private donors compared to NGOs that are funded by a handful of wealthy private donors and government grants [Mechanism: Perception of efficacy - your contribution matters as a small donor. Government funding may also imply lack of independence of government which can reduce the efficiency of an organization.]

```{r contrasts-hypothesis04a}
# Reference Level = Funded primarily by many small private donations
# Compute the contrasts
contrasts <- tibble(
  "wealthy donors - small donors" = draws_intercept %>% filter(level == "Handful of wealthy private donors") %>% select(Gamma) %>% pull(),
  "government grants - small donors" = draws_intercept %>% filter(level == "Government grants") %>% select(Gamma) %>% pull()
)

# Plot the contrasts
contrasts %>%
  pivot_longer(
    "wealthy donors - small donors":"government grants - small donors", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis04a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors indeed show increased willingness to donate to NGOs funded by numerous small private donors, as opposed to those funded by a handful of wealthy private donors or government grants, holding every other attribute constant, with government grants preferred to a small handful of wealthy donors relative to numerous small private donors.

#### Hypothesis 4b, Relationship with Host Government and Funding Sources

> Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors

```{r contrasts-hypothesis04b}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_4b.rds"))

# Reference Level = Funded primarily by many small private donations
# Reference Level = Friendly relationship with government
# Compute the contrasts
contrasts_int <- tibble(
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  criticized_wealthydonors = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  criticized_grants = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  crackdown_wealthydonors = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  crackdown_grants = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
) %>% 
  mutate(
    "wealthy donors - small donors" = few_wealthy_private_donors,
    "government grants - small donors" = government_grants,
    "criticized wealthy donors - criticized small donors" = 
      (few_wealthy_private_donors + criticized + criticized_wealthydonors) - (criticized),
    "criticized government grants - criticized small donors" = 
      (government_grants + criticized + criticized_grants) - (criticized),
    "crackdown wealthy donors - crackdown small donors" = 
      (few_wealthy_private_donors + crackdown + crackdown_wealthydonors) - (crackdown),
    "crackdown government grants - crackdown small donors" = 
      (government_grants + crackdown + crackdown_grants) - (crackdown)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "wealthy donors - small donors":"crackdown government grants - crackdown small donors", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis04b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

There is no significant increase in willingness to donate to NGOs facing government criticism or crackdown that are funded primarily by small private donors, holding every other attribute constant. The same pattern holds with small private donors most preferred, followed by government grants and wealthy donors, with their relative importance held constant.

#### Hypothesis 4c, Relationship with Host Government, Funding Sources, and Issue Area

> Donors will show increased willingness to donate to NGOs that are facing government crackdown and are funded primarily by numerous small private donors and work in less contentious areas (emergency response and refugee relief)

```{r contrasts-hypothesis04c}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_4c.rds"))

# Reference Level = Emergency Response
# Reference Level = Funded primarily by many small private donations
# Reference Level = Friendly relationship with government
# Compute the contrasts
contrasts_int <- tibble(
  # Attribute levels
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  # Hypothesis 4c
  emergency_crackdown_wealthydonors = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  emergency_crackdown_grants = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  environment_crackdown_wealthydonors = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  environment_crackdown_grants = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  humanrights_crackdown_wealthydonors = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  humanrights_crackdown_grants = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  refugee_crackdown_wealthydonors = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  refugee_crackdown_grants = draws %>% filter(i == 21) %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "wealthy donors environment - small donors emergency" = 
      few_wealthy_private_donors + environment,
    "wealthy donors human rights - small donors emergency" = 
      few_wealthy_private_donors + human_rights,
    "wealthy donors refugee relief - small donors emergency" = 
      few_wealthy_private_donors + refugee_relief,
    "crackdown wealthy donors environment - crackdown small donors emergency" = 
      (few_wealthy_private_donors + environment + crackdown + environment_crackdown_wealthydonors) - (crackdown + emergency_crackdown_wealthydonors),
    "crackdown wealthy donors human rights - crackdown small donors emergency" = 
      (few_wealthy_private_donors + human_rights + crackdown + humanrights_crackdown_wealthydonors) - (crackdown + emergency_crackdown_wealthydonors),
    "crackdown wealthy donors refugee relief - crackdown small donors emergency" = 
      (few_wealthy_private_donors + refugee_relief + crackdown + refugee_crackdown_wealthydonors) - (crackdown + emergency_crackdown_wealthydonors),
    "government grants environment - small donors emergency" = 
      government_grants + environment,
    "government grants human rights - small donors emergency" = 
      government_grants + human_rights,
    "government grants refugee relief - small donors emergency" = 
      government_grants + refugee_relief,
    "crackdown government grants environment - crackdown small donors emergency" = 
      (government_grants + environment + crackdown + environment_crackdown_grants) - (crackdown + emergency_crackdown_grants),
    "crackdown government grants human rights - crackdown small donors emergency" = 
      (government_grants + human_rights + crackdown + humanrights_crackdown_grants) - (crackdown + emergency_crackdown_grants),
    "crackdown government grants refugee relief - crackdown small donors emergency" = 
      (government_grants + refugee_relief + crackdown + refugee_crackdown_grants) - (crackdown + emergency_crackdown_grants)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "wealthy donors environment - small donors emergency":"crackdown government grants refugee relief - crackdown small donors emergency", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis04c.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

There an increase in likelihood to donate in general when NGOs are funded by government grants, which we've seen separately, though the difference isn't significant when accounting for the NGOs being under crackdown, holding every other attribute constant. We also see that refugee relief continues to be seen as contentious as the environment. Interestingly, the median likelihood to donate is somewhat normalized across issue areas when the NGOs are under crackdown and all funded by wealthy donors.

### Hypothesis 5, Organizational Practices

#### Hypothesis 5a, Financial Transparency

> Donors will show increased willingness to donate to NGOs that are financially transparent [Mechanism: Perception of efficacy]

```{r contrasts-hypothesis05a}
# Reference Level = No financial transparency
# Compute the contrasts
contrasts <- tibble(
  "financial transparency - no financial transparency" = draws_intercept %>% filter(level == "Financial transparency") %>% select(Gamma) %>% pull()
)

# Plot the contrasts
contrasts %>%
  pivot_longer("financial transparency - no financial transparency", names_to = "level", values_to = "Gamma") %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05a.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors show an increased likelihood to donate to NGOs that are financially transparent, holding every other attribute constant.

#### Hypothesis 5b, Relationship with Host Government and Financial Transparency

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent

```{r contrasts-hypothesis05b}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5b.rds"))

# Reference Level = No financial transparency
# Reference Level = Friendly relationship with government
contrasts_int <- tibble(
  # Attribute levels
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  # Hypothesis 5b
  fin_transparency_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  fin_transparency_crackdown = draws %>% filter(i == 15) %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "financial transparency - no financial transparency" = fin_transparency,
    "criticized financial transparency - criticized no financial transparency" = 
      (fin_transparency + criticized + fin_transparency_criticized) - (criticized),
    "crackdown financial transparency - crackdown no financial transparency" = 
      (fin_transparency + crackdown + fin_transparency_crackdown) - (crackdown)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "financial transparency - no financial transparency":"crackdown financial transparency - crackdown no financial transparency", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05b.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

There is no increased willingness to donate to NGOs that are criticized or under government crackdown that are also financially transparent, holding every other attribute constant.

#### Hypothesis 5c, Relationship with Host Government, Financial Transparency, and Funding Sources

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and are funded primarily by numerous small private donors

```{r contrasts-hypothesis05c}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5c.rds"))

# Reference Level = No financial transparency
# Reference Level = Friendly relationship with government
# Reference Level = Funded primarily by many small private donations
contrasts_int <- tibble(
  # Attribute levels
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  # Hypothesis 5c
  fin_smalldonors_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  fin_wealthydonors_criticized = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  fin_grants_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  fin_smalldonors_crackdown = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  fin_wealthydonors_crackdown = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  fin_grants_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "financially transparent small donors - not financially transparent small donors" = 
      (fin_transparency),
    "financially transparent wealthy donors - not financially transparent wealthy donors" = 
      (fin_transparency + few_wealthy_private_donors) - (few_wealthy_private_donors),
    "financially transparent government grants - not financially transparent government grants" = 
      (fin_transparency + government_grants) - (government_grants),
    "criticized financially transparent small donors - criticized not financially transparent small donors" = 
      (fin_transparency + criticized + fin_smalldonors_criticized) - (criticized),
    "criticized financially transparent wealthy donors - criticized not financially transparent wealthy donors" =
      (fin_transparency + few_wealthy_private_donors + criticized + fin_wealthydonors_criticized) - (few_wealthy_private_donors + criticized),
    "criticized financially transparent government grants - criticized not financially transparent government grants" = 
      (fin_transparency + government_grants + criticized + fin_grants_criticized) - (government_grants + criticized),
    "crackdown financially transparent small donors - crackdown not financially transparent small donors" = 
      (fin_transparency + crackdown + fin_smalldonors_crackdown) - (crackdown),
    "crackdown financially transparent wealthy donors - crackdown not financially transparent wealthy donors" =
      (fin_transparency + few_wealthy_private_donors + crackdown + fin_wealthydonors_crackdown) - (few_wealthy_private_donors + crackdown),
    "crackdown financially transparent government grants - crackdown not financially transparent government grants" = 
      (fin_transparency + government_grants + crackdown + fin_grants_crackdown) - (government_grants + crackdown),
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "financially transparent small donors - not financially transparent small donors":"crackdown financially transparent government grants - crackdown not financially transparent government grants", 
    names_to = "level", values_to = "Gamma"
  ) %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05c.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

There is no increased willingness to donate to NGOs that are criticized or under government crackdown when they are also financially transparent and funded by numerous small private donors, holding every other attribute constant.

#### Hypothesis 5d, Relationship with Host Government, Financial Transparency, and Issue Area

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief)

```{r contrasts-hypothesis05d}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5d.rds"))

# Reference Level = No financial transparency
# Reference Level = Friendly relationship with government
# Reference Level = Emergency Response
contrasts_int <- tibble(
  # Attribute levels
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  # Hypothesis 5d
  fin_emergency_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  fin_emergency_crackdown = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  fin_environment_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  fin_environment_crackdown = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  fin_humanrights_criticized = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  fin_humanrights_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  fin_refugee_criticized = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  fin_refugee_crackdown = draws %>% filter(i == 21) %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "financially transparent environment - not financially transparent emergency" = 
      (fin_transparency + environment),
    "financially transparent human rights - not financially transparent emergency" = 
      (fin_transparency + human_rights),
    "financially transparent refugee relief - not financially transparent emergency" = 
      (fin_transparency + refugee_relief),
    "criticized financially transparent environment - criticized not financially transparent emergency" = 
      (fin_transparency + environment + criticized + fin_environment_criticized) - (criticized),
    "criticized financially transparent human rights - criticized not financially transparent emergency" = 
      (fin_transparency + human_rights + criticized + fin_humanrights_criticized) - (criticized),
    "criticized financially transparent refugee relief - criticized not financially transparent emergency" = 
      (fin_transparency + refugee_relief + criticized + fin_refugee_criticized) - (criticized),
    "crackdown financially transparent environment - crackdown not financially transparent emergency" = 
      (fin_transparency + environment + crackdown + fin_environment_crackdown) - (crackdown),
    "crackdown financially transparent human rights - crackdown not financially transparent emergency" = 
      (fin_transparency + human_rights + crackdown + fin_humanrights_crackdown) - (crackdown),
    "crackdown financially transparent refugee relief - crackdown not financially transparent emergency" = 
      (fin_transparency + refugee_relief + crackdown + fin_refugee_crackdown) - (crackdown)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "financially transparent environment - not financially transparent emergency":"crackdown financially transparent refugee relief - crackdown not financially transparent emergency", 
    names_to = "level", values_to = "Gamma") %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05d.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

There is no increased willingness to donate to NGOs that are criticized or under government crackdown when they are also financially transparent and work in less contentious areas, holding every other attribute constant.

#### Hypothesis 5e, Relationship with Host Government, Financial Transparency, Issue Area, and Funding Sources

> Donors will show increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief) and are funded by numerous small donors

```{r contrasts-hypothesis05e}
# Import the appropriate interaction draws
draws <- read_rds(here::here("data", "raw_data", "posterior_draws", "interactions_5e.rds"))

# Reference Level = No financial transparency
# Reference Level = Friendly relationship with government
# Reference Level = Emergency Response
# Reference Level = Funded primarily by many small private donations
contrasts_int <- tibble(
  # Attribute levels
  environment = draws %>% filter(i == 5) %>% select(Gamma) %>% pull(),
  human_rights = draws %>% filter(i == 6) %>% select(Gamma) %>% pull(),
  refugee_relief = draws %>% filter(i == 7) %>% select(Gamma) %>% pull(),
  fin_transparency = draws %>% filter(i == 8)%>% select(Gamma) %>% pull(),
  few_wealthy_private_donors = draws %>% filter(i == 10) %>% select(Gamma) %>% pull(),
  government_grants = draws %>% filter(i == 11) %>% select(Gamma) %>% pull(),
  criticized = draws %>% filter(i == 12)%>% select(Gamma) %>% pull(),
  crackdown = draws %>% filter(i == 13) %>% select(Gamma) %>% pull(),
  # Hypothesis 5e
  fin_emergency_criticized = draws %>% filter(i == 14) %>% select(Gamma) %>% pull(),
  fin_emergency_crackdown = draws %>% filter(i == 15) %>% select(Gamma) %>% pull(),
  fin_environment_criticized = draws %>% filter(i == 16) %>% select(Gamma) %>% pull(),
  fin_environment_crackdown = draws %>% filter(i == 17) %>% select(Gamma) %>% pull(),
  fin_humanrights_criticized = draws %>% filter(i == 18) %>% select(Gamma) %>% pull(),
  fin_humanrights_crackdown = draws %>% filter(i == 19) %>% select(Gamma) %>% pull(),
  fin_refugee_criticized = draws %>% filter(i == 20) %>% select(Gamma) %>% pull(),
  fin_refugee_crackdown = draws %>% filter(i == 21) %>% select(Gamma) %>% pull()
) %>% 
  mutate(
    "financially transparent environment small donors - not financially transparent emergency small donors" = 
      (fin_transparency + environment),
    "financially transparent human rights small donors - not financially transparent emergency small donors" = 
      (fin_transparency + human_rights),
    "financially transparent refugee relief small donors - not financially transparent emergency small donors" = 
      (fin_transparency + refugee_relief),
    "criticized financially transparent environment small donors - criticized not financially transparent emergency small donors" = 
      (fin_transparency + environment + criticized + fin_environment_criticized) - (criticized),
    "criticized financially transparent human rights small donors - criticized not financially transparent emergency small donors" = 
      (fin_transparency + human_rights + criticized + fin_humanrights_criticized) - (criticized),
    "criticized financially transparent refugee relief small donors - criticized not financially transparent emergency small donors" = 
      (fin_transparency + refugee_relief + criticized + fin_refugee_criticized) - (criticized),
    "crackdown financially transparent environment small donors - crackdown not financially transparent emergency small donors" = 
      (fin_transparency + environment + crackdown + fin_environment_crackdown) - (crackdown),
    "crackdown financially transparent human rights small donors - crackdown not financially transparent emergency small donors" = 
      (fin_transparency + human_rights + crackdown + fin_humanrights_crackdown) - (crackdown),
    "crackdown financially transparent refugee relief small donors - crackdown not financially transparent emergency small donors" = 
      (fin_transparency + refugee_relief + crackdown + fin_refugee_crackdown) - (crackdown),
    "financially transparent environment wealthy donors - not financially transparent emergency wealthy donors" = 
      (fin_transparency + environment + few_wealthy_private_donors) - (few_wealthy_private_donors),
    "financially transparent human rights wealthy donors - not financially transparent emergency wealthy donors" = 
      (fin_transparency + human_rights + few_wealthy_private_donors) - (few_wealthy_private_donors),
    "financially transparent refugee relief wealthy donors - not financially transparent emergency wealthy donors" = 
      (fin_transparency + refugee_relief + few_wealthy_private_donors) - (few_wealthy_private_donors),
    "criticized financially transparent environment wealthy donors - criticized not financially transparent emergency wealthy donors" = 
      (fin_transparency + environment + criticized + fin_environment_criticized + few_wealthy_private_donors) - (criticized + few_wealthy_private_donors),
    "criticized financially transparent human rights wealthy donors - criticized not financially transparent emergency wealthy donors" = 
      (fin_transparency + human_rights + criticized + fin_humanrights_criticized + few_wealthy_private_donors) - (criticized + few_wealthy_private_donors),
    "criticized financially transparent refugee relief wealthy donors - criticized not financially transparent emergency wealthy donors" = 
      (fin_transparency + refugee_relief + criticized + fin_refugee_criticized + few_wealthy_private_donors) - (criticized + few_wealthy_private_donors),
    "crackdown financially transparent environment wealthy donors - crackdown not financially transparent emergency wealthy donors" = 
      (fin_transparency + environment + crackdown + fin_environment_crackdown + few_wealthy_private_donors) - (crackdown + few_wealthy_private_donors),
    "crackdown financially transparent human rights wealthy donors - crackdown not financially transparent emergency wealthy donors" = 
      (fin_transparency + human_rights + crackdown + fin_humanrights_crackdown + few_wealthy_private_donors) - (crackdown + few_wealthy_private_donors),
    "crackdown financially transparent refugee relief wealthy donors - crackdown not financially transparent emergency wealthy donors" = 
      (fin_transparency + refugee_relief + crackdown + fin_refugee_crackdown + few_wealthy_private_donors) - (crackdown + few_wealthy_private_donors),
    "financially transparent environment government grants - not financially transparent emergency government grants" = 
      (fin_transparency + environment + government_grants) - (government_grants),
    "financially transparent human rights government grants - not financially transparent emergency government grants" = 
      (fin_transparency + human_rights + government_grants) - (government_grants),
    "financially transparent refugee relief government grants - not financially transparent emergency government grants" = 
      (fin_transparency + refugee_relief + government_grants) - (government_grants),
    "criticized financially transparent environment government grants - criticized not financially transparent emergency government grants" = 
      (fin_transparency + environment + criticized + fin_environment_criticized + government_grants) - (criticized + government_grants),
    "criticized financially transparent human rights government grants - criticized not financially transparent emergency government grants" = 
      (fin_transparency + human_rights + criticized + fin_humanrights_criticized + government_grants) - (criticized + government_grants),
    "criticized financially transparent refugee relief government grants - criticized not financially transparent emergency government grants" = 
      (fin_transparency + refugee_relief + criticized + fin_refugee_criticized + government_grants) - (criticized + government_grants),
    "crackdown financially transparent environment government grants - crackdown not financially transparent emergency government grants" = 
      (fin_transparency + environment + crackdown + fin_environment_crackdown + government_grants) - (crackdown + government_grants),
    "crackdown financially transparent human rights government grants - crackdown not financially transparent emergency government grants" = 
      (fin_transparency + human_rights + crackdown + fin_humanrights_crackdown + government_grants) - (crackdown + government_grants),
    "crackdown financially transparent refugee relief government grants - crackdown not financially transparent emergency government grants" = 
      (fin_transparency + refugee_relief + crackdown + fin_refugee_crackdown + government_grants) - (crackdown + government_grants)
  )

# Plot the contrasts
contrasts_int %>%
  pivot_longer(
    "financially transparent environment small donors - not financially transparent emergency small donors":"crackdown financially transparent refugee relief government grants - crackdown not financially transparent emergency government grants", 
    names_to = "level", values_to = "Gamma") %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05e.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

There is no increased willingness to donate to NGOs that are criticized by the government/under government crackdown when they are also financially transparent and work in less contentious areas (emergency response and refugee relief) and are funded by numerous small donors, holding every other attribute constant.

#### Hypothesis 5f, Accountability

> Donors should be no more or less likely to donate to NGOs that are accountable and hold regular third party audits [Mechanism: Donors don't necessarily seek assurance through third-party programs/audits and charity watchdogs, but rather through word of mouth, personal scrutiny and local networks]

```{r contrasts-hypothesis05f}
# Reference Level = No accountability
# Compute the contrasts
contrasts <- tibble(
  "accountability - no accountability" = draws_intercept %>% filter(level == "Accountability") %>% select(Gamma) %>% pull()
)

# Plot the contrasts
contrasts %>%
  pivot_longer("accountability - no accountability", names_to = "level", values_to = "Gamma") %>% 
  mutate(level = fct_rev(fct_inorder(level))) %>%
  ggplot(aes(x = Gamma, y = level)) +
  stat_halfeye(.width = .95) +
  geom_vline(xintercept = 0, color = "grey") +
  labs(y = "Contrasts", x = "Difference in Gammas")

# Save plot
ggsave(
  "contrasts_hypothesis05f.png",
  path = here::here("analysis", "output", "figures"),
  width = 12,
  height = 6,
  units = "in"
)
```

Donors show a similar increase in willingness to donate to accountable NGOs in comparison to NGOs that are financially transparent, holding every other attribute constant.

## Model fit table

```{r model-fit}
# Load model fit table
read_rds(model_fit, here::here("analysis", "output", "model_fit.rds"))
```

The intercept-only model works best in terms of PSIS-LOO, followed by models with political ideology, public affairs and demographics covariates. While PSIS-LOO doesn't properly account for the specific benefits of a hierarchical model, it suggests that political ideology above all other covariate sets specified, has the most explanatory power.

## Baseline probabilities

```{r probs-baseline, fig.width=4.5, fig.height=4}
orgs <- draws %>% 
  filter(category == "Organization") %>% 
  mutate(prob = plogis(Theta))

fig_baseline <- ggplot(orgs, aes(x = prob, y = attribute)) +
  geom_halfeyeh(.width = 0.95, aes(fill = attribute)) +
  scale_fill_manual(values = c(clrs$yellow, clrs$olive, clrs$lime, clrs$red),
                    guide = FALSE) +
  scale_x_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = "Baseline probability of selecting organization", y = NULL) +
  theme_ngo()

fig_baseline
ggsave(fig_baseline, filename = here("analysis", "output", "figures", "probs-org.pdf"),
       width = 13/3, height = 3, units = "in", device = cairo_pdf)
ggsave(fig_baseline, filename = here("analysis", "output", "figures", "probs-org.png"),
       width = 13/3, height = 3, units = "in", type = "cairo", dpi = 300)
```

```{r}
orgs %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(tidy = data %>% map(~ median_hdci(.$prob))) %>% 
  unnest(tidy)
```

## Coefficients and marginal effects

```{r first-hypotheses-coefficients, fig.width=26/3, fig.height=5.5}
diff_refugee_hr <- diff_cols("Refugee relief", "Human rights")
diff_hr_env <- diff_cols("Human rights", "Environment")
diff_env_refugee <- diff_cols("Environment", "Refugee relief")

plot_issue <- ggplot(filter(draws, category == "Issue area"), 
                     aes(x = exp(Theta), y = attribute)) + 
  stat_halfeyeh(.width = 0.95, fill = clrs$blue) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.7, ymin = 2.1, ymax = 2.9, size = 0.25) +
  # annotate(geom = "label", x = 0.7, y = 2.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_refugee_hr$phrase) +
  # annotate(geom = "linerange", x = 0.8, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.8, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_hr_env$phrase) +
  # annotate(geom = "linerange", x = 0.585, ymin = 1.1, ymax = 2.9, size = 0.25) +
  # annotate(geom = "label", x = 0.585, y = 2.05, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_env_refugee$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Issue area",
       subtitle = "Reference group = emergency response") +
  theme_ngo()

diff_govt_richpeople <- diff_cols("Government grants", 
                                  "Handful of wealthy private donors")

plot_funding <- ggplot(filter(draws, category == "Funding sources"), 
                       aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$maroon) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.825, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.825, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_govt_richpeople$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Funding sources",
       subtitle = "Reference group = many small private donations") +
  theme_ngo()

diff_account_trans <- diff_cols("Accountability", "Financial transparency")

plot_org <- ggplot(filter(draws, category == "Organizational practices"), 
                   aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$orange) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 1.85, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 1.85, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_account_trans$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Organizational practices",
       subtitle = "Reference groups = no accountability; no transparency") +
  theme_ngo()

diff_govt <- diff_cols("Under crackdown", "Criticized")

plot_government <- ggplot(filter(draws, category == "Relationship with host government"), 
                          aes(x = exp(Theta), y = attribute)) + 
  geom_halfeyeh(.width = 0.95, fill = clrs$teal) +
  geom_vline(xintercept = 1, color = clrs$red, size = 0.5, linetype = "dotted") +
  # annotate(geom = "linerange", x = 0.75, ymin = 1.1, ymax = 1.9, size = 0.25) +
  # annotate(geom = "label", x = 0.75, y = 1.5, hjust = 0.5, size = pts(7),
  #          family = "Roboto Condensed Light",
  #          label = diff_govt$phrase) +
  labs(x = "exp(θ)", y = NULL,
       title = "Relationship with host government",
       subtitle = "Reference group = friendly relationship with government") +
  theme_ngo()

plot_first_hypotheses <- (plot_org + plot_issue + plot_funding + plot_government)

plot_first_hypotheses
ggsave(plot_first_hypotheses, filename = here("analysis", "output", "figures", "results-h1-4.pdf"),
       width = 26/3, height = 5.5, units = "in", device = cairo_pdf)
ggsave(plot_first_hypotheses, filename = here("analysis", "output", "figures", "results-h1-4.png"),
       width = 26/3, height = 5.5, units = "in", type = "cairo", dpi = 300)
```

```{r}
draws %>% 
  filter(category != "Organization") %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(tidy = data %>% map(~ median_hdci(.$Theta))) %>% 
  unnest(tidy) %>% 
  mutate(exp_y = exp(y))
```

## Differences in means

According to Kruschke 8.4, we can look at the posterior distribution of the difference in thetas, then calculate the proportion of those differences that are greater/less than 0

$$
\operatorname{mean}(\theta_\text{Government grants} - \theta_\text{Private donors}) = 
$$

$$
\Delta \theta = 4; p(\Delta > 0) = 5
$$

```{r eval=FALSE}
col1 <- "Government grants"
col2 <- "Handful of wealthy private donors"

posterior_diffs <- draws %>% 
  filter(attribute %in% c({{col1}}, {{col2}})) %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]],
         diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

mean(posterior_diffs$diff_theta_e)

asdf <- draws %>% 
  filter(category == "Funding sources") %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]]) %>% 
  mutate(diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  summarize(bloop = exp(mean(Theta)),
            bloop1 = mean(exp(Theta)))

mean(asdf$diff_theta_e)
tidybayes::hdi(asdf$diff_theta_e)

mean(asdf$diff_theta > 0)
mean(asdf$diff_theta_e > 0)

asdf <- draws %>% 
  filter(category == "Organizational practices") %>% 
  select(.chain, .iteration, Theta, attribute) %>% 
  pivot_wider(names_from = "attribute", values_from = "Theta") %>% 
  mutate(diff_theta = .[[4]] - .[[3]]) %>% 
  mutate(diff_theta_e = exp(.[[4]]) - exp(.[[3]]))

mean(asdf$diff_theta > 0)
mean(asdf$diff_theta_e > 0)

ggplot(asdf, aes(x = diff_theta_e, y = 1)) +
  geom_halfeyeh(.width = 0.95)

draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  summarize(bloop = mean(Theta),
            bloop1 = exp(mean(Theta)))

qwer <- draws %>% 
  filter(category == "Funding sources") %>% 
  group_by(attribute) %>% 
  nest() %>% 
  mutate(hdi = data %>% map(~ exp(tidybayes::hdi(.$Theta, .width = 0.95))))
asdf %>% 
  unnest(hdi)
mean(asdf$Theta)
tidybayes::hdi(asdf$Theta, .width = 0.95)
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=100}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>