### 2023-08-26 analysis/01_experimental-design.Rmd

---
title: "Check experimental design"
author: "Andrew Heiss, Marc Dotson, and Suparna Chaudhry"
date: "Last run: `r format(Sys.time(), '%B %e, %Y')`"
output: 
  html_document:
    code_folding: show
editor_options: 
  chunk_output_type: console
---

# Load the design and check level balance

The experimental design was created as part of the conjoint study in Qualtrics. Here we check to make sure that there is a balance across attribute levels (i.e., levels show up a roughly equal number of times for each attribute).

```{r load-data, warning=FALSE, message=FALSE}
library(tidyverse)
library(fastDummies)
library(here)

# Load design
design <- read_csv(here("data", "derived_data", "design.csv"))

# Check to make sure there is a balance across levels
as_tibble(design) %>%
  select(Organization:`Relationship with host government`) %>%
  gather(key = atts, value = lvls) %>%
  ggplot(aes(x = lvls)) +
  geom_bar() +
  facet_wrap(~ atts, scales = "free", nrow = 2)
```

# Dummy-code the design

Now we prepare the experimental design for use in estimating the model by dropping the first level of each attribute, which are all discrete, as the reference level.

```{r dummy-design, warning=FALSE, message=FALSE}
# Convert the attributes in design into factors
dummy_design <- as.data.frame(design)
for(i in 4:ncol(dummy_design)) {
  dummy_design[,i] <- as.factor(dummy_design[,i])
}

# Rename the variables.
colnames(dummy_design)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get a dummy-coded version of all the attribute levels
dummy_design <- dummy_cols(
  dummy_design,
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = TRUE,
  remove_selected_columns = TRUE
)

# Rename the index variables
colnames(dummy_design)[4:15] <- c(
  str_c("org", 2:4), str_c("issue", 2:4), str_c("transp", 2),
  str_c("account", 2), str_c("fund", 2:3), str_c("gov_rel", 2:3)
)

# Save dummy-coded design output
write.csv(dummy_design, here("data", "derived_data", "dummy_design.csv"))
```

# Code all possible interactions

We have a number of hypotheses referencing specific interactions. To answer these hypotheses, we need to specify each of the needed interactions.

```{r dummy-interactions, warning=FALSE, message=FALSE}
# Convert the attributes in design into factors
interactions <- as.data.frame(design)
for(i in 4:ncol(interactions)) {
  interactions[,i] <- as.factor(interactions[,i])
}

# Rename the variables.
colnames(interactions)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get a dummy-coded version of all the attribute levels
interactions <- dummy_cols(
  interactions, 
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = FALSE,
  remove_selected_columns = TRUE
)

# Rename the dummy-coded variables
colnames(interactions)[4:21] <- c(
  str_c("org", 1:4), str_c("issue", 1:4), str_c("transp", 1:2), 
  str_c("account", 1:2), str_c("fund", 1:3), str_c("gov_rel", 1:3)
)

# Build out all the necessary interactions
interactions <- interactions %>% 
  as_tibble() %>% 
  mutate(
    # Hypothesis 2b.
    amnesty_criticized = org1 * gov_rel2,
    greenpeace_criticized = org2 * gov_rel2,
    oxfam_criticized = org3 * gov_rel2,
    redcross_criticized = org4 * gov_rel2,
    amnesty_crackdown = org1 * gov_rel3,
    greenpeace_crackdown = org2 * gov_rel3,
    oxfam_crackdown = org3 * gov_rel3,
    redcross_crackdown = org4 * gov_rel3,
    # Hypothesis 3b.
    # emergency_criticized = issue1 * gov_rel2,
    environment_criticized = issue2 * gov_rel2,
    humanrights_criticized = issue3 * gov_rel2,
    refugees_criticized = issue4 * gov_rel2,
    # emergency_crackdown = issue1 * gov_rel3,
    environment_crackdown = issue2 * gov_rel3,
    humanrights_crackdown = issue3 * gov_rel3,
    refugees_crackdown = issue4 * gov_rel3,
    # Hypothesis 4b.
    # smalldonors_criticized = fund1 * gov_rel2,
    wealthydonors_criticized = fund2 * gov_rel2,
    govgrants_criticized = fund3 * gov_rel2,
    # smalldonors_crackdown = fund1 * gov_rel3,
    wealthydonors_crackdown = fund2 * gov_rel3,
    govgrants_crackdown = fund3 * gov_rel3,
    # Hypothesis 4c.
    # emergency_smalldonors_crackdown = issue1 * fund1 * gov_rel3,
    emergency_wealthydonors_crackdown = issue1 * fund2 * gov_rel3, # ??????
    emergency_govgrants_crackdown = issue1 * fund3 * gov_rel3,     # ??????
    # environment_smalldonors_crackdown = issue2 * fund1 * gov_rel3,
    environment_wealthydonors_crackdown = issue2 * fund2 * gov_rel3,
    environment_govgrants_crackdown = issue2 * fund3 * gov_rel3,
    # humanrights_smalldonors_crackdown = issue3 * fund1 * gov_rel3,
    humanrights_wealthydonors_crackdown = issue3 * fund2 * gov_rel3,
    humanrights_govgrants_crackdown = issue3 * fund3 * gov_rel3,
    # refugees_smalldonors_crackdown = issue4 * fund1 * gov_rel3,
    refugees_wealthydonors_crackdown = issue4 * fund2 * gov_rel3,
    refugees_govgrants_crackdown = issue4 * fund3 * gov_rel3,
    # Hypothesis 5b.
    # nottransp_criticized = transp1 * gov_rel2,
    transp_criticized = transp2 * gov_rel2,
    # nottransp_crackdown = transp1 * gov_rel3,
    transp_crackdown = transp2 * gov_rel3,
    # Hypothesis 5c.
    # nottransp_smalldonors_criticized = transp1 * fund1 * gov_rel2,
    transp_smalldonors_criticized = transp2 * fund1 * gov_rel2,
    # nottransp_wealthydonors_criticized = transp1 * fund2 * gov_rel2,
    transp_wealthydonors_criticized = transp2 * fund2 * gov_rel2,
    # nottransp_govgrants_criticized = transp1 * fund3 * gov_rel2,
    transp_govgrants_criticized = transp2 * fund3 * gov_rel2,
    # nottransp_smalldonors_crackdown = transp1 * fund1 * gov_rel3,
    transp_smalldonors_crackdown = transp2 * fund1 * gov_rel3,
    # nottransp_wealthydonors_crackdown = transp1 * fund2 * gov_rel3,
    transp_wealthydonors_crackdown = transp2 * fund2 * gov_rel3,
    # nottransp_govgrants_crackdown = transp1 * fund3 * gov_rel3,
    transp_govgrants_crackdown = transp2 * fund3 * gov_rel3,
    # Hypothesis 5d.
    # nottransp_emergency_criticized = transp1 * issue1 * gov_rel2,
    transp_emergency_criticized = transp2 * issue1 * gov_rel2,
    # nottransp_emergency_crackdown = transp1 * issue1 * gov_rel3,
    transp_emergency_crackdown = transp2 * issue1 * gov_rel3,
    # nottransp_environment_criticized = transp1 * issue2 * gov_rel2,
    transp_environment_criticized = transp2 * issue2 * gov_rel2,
    # nottransp_environment_crackdown = transp1 * issue2 * gov_rel3,
    transp_environment_crackdown = transp2 * issue2 * gov_rel3,
    # nottransp_humanrights_criticized = transp1 * issue3 * gov_rel2,
    transp_humanrights_criticized = transp2 * issue3 * gov_rel2,
    # nottransp_humanrights_crackdown = transp1 * issue3 * gov_rel3,
    transp_humanrights_crackdown = transp2 * issue3 * gov_rel3,
    # nottransp_refugees_criticized = transp1 * issue4 * gov_rel2,
    transp_refugees_criticized = transp2 * issue4 * gov_rel2,
    # nottransp_refugees_crackdown = transp1 * issue4 * gov_rel3,
    transp_refugees_crackdown = transp2 * issue4 * gov_rel3,
    # Hypothesis 5e.
    # nottransp_smalldonors_emergency_criticized = transp1 * fund1 * issue1 * gov_rel2,
    transp_smalldonors_emergency_criticized = transp2 * fund1 * issue1 * gov_rel2,
    # nottransp_smalldonors_emergency_crackdown = transp1 * fund1 * issue1 * gov_rel3,
    transp_smalldonors_emergency_crackdown = transp2 * fund1 * issue1 * gov_rel3,
    # nottransp_smalldonors_environment_criticized = transp1 * fund1 * issue2 * gov_rel2,
    transp_smalldonors_environment_criticized = transp2 * fund1 * issue2 * gov_rel2,
    # nottransp_smalldonors_environment_crackdown = transp1 * fund1 * issue2 * gov_rel3,
    transp_smalldonors_environment_crackdown = transp2 * fund1 * issue2 * gov_rel3,
    # nottransp_smalldonors_humanrights_criticized = transp1 * fund1 * issue3 * gov_rel2,
    transp_smalldonors_humanrights_criticized = transp2 * fund1 * issue3 * gov_rel2,
    # nottransp_smalldonors_humanrights_crackdown = transp1 * fund1 * issue3 * gov_rel3,
    transp_smalldonors_humanrights_crackdown = transp2 * fund1 * issue3 * gov_rel3,
    # nottransp_smalldonors_refugees_criticized = transp1 * fund1 * issue4 * gov_rel2,
    transp_smalldonors_refugees_criticized = transp2 * fund1 * issue4 * gov_rel2,
    # nottransp_smalldonors_refugees_crackdown = transp1 * fund1 * issue4 * gov_rel3,
    transp_smalldonors_refugees_crackdown = transp2 * fund1 * issue4 * gov_rel3
  ) %>% 
  select(-contains("1"))

# Save interact-coded design output
write.csv(interactions, here("data", "derived_data", "dummy_interactions.csv"))
```

# Index-coded experimental alternative

It's possible an index-coded design will serve us better, allowing us to compute whatever contrasts are needed from a single model. An index-coded design has no reference levels.

```{r index-design, warning=FALSE, message=FALSE}
################################################################################
# This is an experimental alternative to the dummy-coded design.
################################################################################

# Convert the attributes in design into factors
index_design <- as.data.frame(design)
for(i in 4:ncol(index_design)) {
  index_design[,i] <- as.factor(index_design[,i])
}

# Rename the variables.
colnames(index_design)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get an index-coded version of all the attribute levels
index_design <- dummy_cols(
  index_design,
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = FALSE,
  remove_selected_columns = TRUE
)

# Rename the index-coded variables
colnames(index_design)[4:21] <- c(
  str_c("org", 1:4), str_c("issue", 1:4), str_c("transp", 1:2), 
  str_c("account", 1:2), str_c("fund", 1:3), str_c("gov_rel", 1:3)
)

# Save index-coded design output
write.csv(index_design, here("data", "derived_data", "index_design.csv"))
```

We have a number of hypotheses referencing specific interactions. To answer these hypotheses, we need to specify each of the needed interactions.

```{r index-interactions, warning=FALSE, message=FALSE}
################################################################################
# This is an experimental alternative to the interactions based on the 
# dummy-coded design.
################################################################################

# Convert the attributes in design into factors
interactions <- as.data.frame(design)
for(i in 4:ncol(interactions)) {
  interactions[,i] <- as.factor(interactions[,i])
}

# Rename the variables
colnames(interactions)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get an index-coded version of all the attribute levels
interactions <- dummy_cols(
  interactions, 
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = FALSE,
  remove_selected_columns = TRUE
)

# Rename the index-coded variables
colnames(interactions)[4:21] <- c(
  str_c("org", 1:4), str_c("issue", 1:4), str_c("transp", 1:2), 
  str_c("account", 1:2), str_c("fund", 1:3), str_c("gov_rel", 1:3)
)

# Build out all the necessary interactions
interactions <- interactions %>% 
  as_tibble() %>% 
  mutate(
    # Hypothesis 2b.
    amnesty_criticized = org1 * gov_rel2,
    greenpeace_criticized = org2 * gov_rel2,
    oxfam_criticized = org3 * gov_rel2,
    redcross_criticized = org4 * gov_rel2,
    amnesty_crackdown = org1 * gov_rel3,
    greenpeace_crackdown = org2 * gov_rel3,
    oxfam_crackdown = org3 * gov_rel3,
    redcross_crackdown = org4 * gov_rel3,
    # Hypothesis 3b.
    emergency_criticized = issue1 * gov_rel2,
    environment_criticized = issue2 * gov_rel2,
    humanrights_criticized = issue3 * gov_rel2,
    refugees_criticized = issue4 * gov_rel2,
    emergency_crackdown = issue1 * gov_rel3,
    environment_crackdown = issue2 * gov_rel3,
    humanrights_crackdown = issue3 * gov_rel3,
    refugees_crackdown = issue4 * gov_rel3,
    # Hypothesis 4b.
    smalldonors_criticized = fund1 * gov_rel2,
    wealthydonors_criticized = fund2 * gov_rel2,
    govgrants_criticized = fund3 * gov_rel2,
    smalldonors_crackdown = fund1 * gov_rel3,
    wealthydonors_crackdown = fund2 * gov_rel3,
    govgrants_crackdown = fund3 * gov_rel3,
    # Hypothesis 4c.
    emergency_smalldonors_crackdown = issue1 * fund1 * gov_rel3,
    emergency_wealthydonors_crackdown = issue1 * fund2 * gov_rel3, # ??????
    emergency_govgrants_crackdown = issue1 * fund3 * gov_rel3,     # ??????
    environment_smalldonors_crackdown = issue2 * fund1 * gov_rel3,
    environment_wealthydonors_crackdown = issue2 * fund2 * gov_rel3,
    environment_govgrants_crackdown = issue2 * fund3 * gov_rel3,
    humanrights_smalldonors_crackdown = issue3 * fund1 * gov_rel3,
    humanrights_wealthydonors_crackdown = issue3 * fund2 * gov_rel3,
    humanrights_govgrants_crackdown = issue3 * fund3 * gov_rel3,
    refugees_smalldonors_crackdown = issue4 * fund1 * gov_rel3,
    refugees_wealthydonors_crackdown = issue4 * fund2 * gov_rel3,
    refugees_govgrants_crackdown = issue4 * fund3 * gov_rel3,
    # Hypothesis 5b.
    nottransp_criticized = transp1 * gov_rel2,
    transp_criticized = transp2 * gov_rel2,
    nottransp_crackdown = transp1 * gov_rel3,
    transp_crackdown = transp2 * gov_rel3,
    # Hypothesis 5c.
    nottransp_smalldonors_criticized = transp1 * fund1 * gov_rel2,
    transp_smalldonors_criticized = transp2 * fund1 * gov_rel2,
    nottransp_wealthydonors_criticized = transp1 * fund2 * gov_rel2,
    transp_wealthydonors_criticized = transp2 * fund2 * gov_rel2,
    nottransp_govgrants_criticized = transp1 * fund3 * gov_rel2,
    transp_govgrants_criticized = transp2 * fund3 * gov_rel2,
    nottransp_smalldonors_crackdown = transp1 * fund1 * gov_rel3,
    transp_smalldonors_crackdown = transp2 * fund1 * gov_rel3,
    nottransp_wealthydonors_crackdown = transp1 * fund2 * gov_rel3,
    transp_wealthydonors_crackdown = transp2 * fund2 * gov_rel3,
    nottransp_govgrants_crackdown = transp1 * fund3 * gov_rel3,
    transp_govgrants_crackdown = transp2 * fund3 * gov_rel3,
    # Hypothesis 5d.
    nottransp_emergency_criticized = transp1 * issue1 * gov_rel2,
    transp_emergency_criticized = transp2 * issue1 * gov_rel2,
    nottransp_emergency_crackdown = transp1 * issue1 * gov_rel3,
    transp_emergency_crackdown = transp2 * issue1 * gov_rel3,
    nottransp_environment_criticized = transp1 * issue2 * gov_rel2,
    transp_environment_criticized = transp2 * issue2 * gov_rel2,
    nottransp_environment_crackdown = transp1 * issue2 * gov_rel3,
    transp_environment_crackdown = transp2 * issue2 * gov_rel3,
    nottransp_humanrights_criticized = transp1 * issue3 * gov_rel2,
    transp_humanrights_criticized = transp2 * issue3 * gov_rel2,
    nottransp_humanrights_crackdown = transp1 * issue3 * gov_rel3,
    transp_humanrights_crackdown = transp2 * issue3 * gov_rel3,
    nottransp_refugees_criticized = transp1 * issue4 * gov_rel2,
    transp_refugees_criticized = transp2 * issue4 * gov_rel2,
    nottransp_refugees_crackdown = transp1 * issue4 * gov_rel3,
    transp_refugees_crackdown = transp2 * issue4 * gov_rel3,
    # Hypothesis 5e.
    nottransp_smalldonors_emergency_criticized = transp1 * fund1 * issue1 * gov_rel2,
    transp_smalldonors_emergency_criticized = transp2 * fund1 * issue1 * gov_rel2,
    nottransp_smalldonors_emergency_crackdown = transp1 * fund1 * issue1 * gov_rel3,
    transp_smalldonors_emergency_crackdown = transp2 * fund1 * issue1 * gov_rel3,
    nottransp_smalldonors_environment_criticized = transp1 * fund1 * issue2 * gov_rel2,
    transp_smalldonors_environment_criticized = transp2 * fund1 * issue2 * gov_rel2,
    nottransp_smalldonors_environment_crackdown = transp1 * fund1 * issue2 * gov_rel3,
    transp_smalldonors_environment_crackdown = transp2 * fund1 * issue2 * gov_rel3,
    nottransp_smalldonors_humanrights_criticized = transp1 * fund1 * issue3 * gov_rel2,
    transp_smalldonors_humanrights_criticized = transp2 * fund1 * issue3 * gov_rel2,
    nottransp_smalldonors_humanrights_crackdown = transp1 * fund1 * issue3 * gov_rel3,
    transp_smalldonors_humanrights_crackdown = transp2 * fund1 * issue3 * gov_rel3,
    nottransp_smalldonors_refugees_criticized = transp1 * fund1 * issue4 * gov_rel2,
    transp_smalldonors_refugees_criticized = transp2 * fund1 * issue4 * gov_rel2,
    nottransp_smalldonors_refugees_crackdown = transp1 * fund1 * issue4 * gov_rel3,
    transp_smalldonors_refugees_crackdown = transp2 * fund1 * issue4 * gov_rel3
  )

# Save interact-coded design output
write.csv(interactions, here("data", "derived_data", "index_interactions.csv"))
```

## Aggregate-coded interaction alternative

```{r aggregate-interactions, warning=FALSE, message=FALSE}
################################################################################
# This is an experimental alternative where specific contrasts are estimated
# as part of the model (hence aggregate-coded) rather than computed using
# marginal posteriors after estimation.
################################################################################

# Convert the attributes in design into factors
interactions <- as.data.frame(design)
for(i in 4:ncol(interactions)) {
  interactions[,i] <- as.factor(interactions[,i])
}

# Rename the variables
colnames(interactions)[4:9] <- c("org", "issue", "transp", "account", "fund", "gov_rel")

# Use dummy_cols() to get an index-coded version of all the attribute levels
interactions <- dummy_cols(
  interactions,
  select_columns = c("org", "issue", "transp", "account", "fund", "gov_rel"),
  remove_first_dummy = FALSE,
  remove_selected_columns = TRUE
)

# Rename the index-coded variables
colnames(interactions)[4:21] <- c(
  str_c("org", 1:4), str_c("issue", 1:4), str_c("transp", 1:2),
  str_c("account", 1:2), str_c("fund", 1:3), str_c("gov_rel", 1:3)
)

################################################################################
# Are we sure this is encoding the needed contrasts exactly??
################################################################################

# # Build out all the necessary interactions...
# # ...by multiplying together the needed index-coded levels.
# org1 = amnesty
# org2 = greenpeace
# org3 = oxfam
# org4 = redcross
# gov_rel1 = friendly
# gov_rel2 = criticized
# gov_rel3 = crackdown
# issue1 = emergency
# issue2 = environment
# issue3 = human rights
# issue4 = refugee
# transp1 = not transparent
# transp2 = transparent
# account1 = not accountable
# account2 = accountable
# fund1 = smalldonors
# fund2 = wealthydonors
# fund3 = gov grants

# Sticking exactly to the hypotheses where each contrasts will 
# consist of h*_high - h*_low
interactions <- interactions %>% 
  as_tibble() %>% 
  rowwise() %>% 
  mutate(
    # Main effects
    h1_high = sum(org3, org4), # oxfam and red cross
    h1_low = sum(org1, org2), # amnesty and greenpeace
    h2a_high = sum(gov_rel2,gov_rel3), # criticized and crackdown
    h2a_low = gov_rel1, # friendly
    h3a_high = sum(issue1, issue4), # emergency and refugee
    h3a_low = sum(issue2, issue3), # environment and human rights
    h4a_high = fund1, # small donors
    h4a_low = sum(fund2, fund3), # wealthy donors and gov grants
    h5a_high = transp2, # transparent
    h5a_low = transp1, # not transparent
    # First-order Interactions
    h2b_high = h1_high*h2a_high, #  oxfam and red cross * criticized and crackdown
    h2b_low = h1_low*gov_rel3, # amnesty and greenpeace * criticized
    h3b_high = h2a_high*h3a_high, # criticized and crackdown * emergency and refugee
    h3b_low = h2a_high*h3a_low, # criticized and crackdown * environment and human rights
    h4b_high = gov_rel3*fund1, # crackdown * small donors
    h4b_low = sum(gov_rel1, gov_rel2)*h4a_high, # not crackdown * not small donors
    h5b_high = h2a_high*h5a_high, # criticized and crackdown * transparent
    h5b_low = h5a_low + h2a_low*h5a_high, # not transparent + friendly * transparent + 
    # Second-order interactions
    h4c_high = gov_rel3*h4a_high*h3a_high, # crackdown * small donors * emergency and refugee
    h4c_low = ifelse(h4c_high==1, 0, 1), # everything else
    h5c_high = h2a_high*h5a_high*h4a_high, # criticized and crackdown * transparent * small donors
    h5c_low = ifelse(h5c_high==1, 0, 1), # everything else
    h5d_high = h2a_high*h5a_high*h3a_high, # criticized and crackdown * transparent * emergency and refugee
    h5d_low = ifelse(h5d_high==1, 0, 1), # everything else
    # Third-order interaction
    h5e_high = h2a_high*h5a_high*h3a_high*h4a_high, # criticized and crackdown * transparent * emergency and refugee * small donors
    h5e_low = ifelse(h5e_high==1, 0, 1)) %>% 
  ungroup()

# Save interact-coded design output
write.csv(interactions, here("data", "derived_data", "aggregate_interactions.csv"))
```

# Original computing environment

<button data-toggle="collapse" data-target="#sessioninfo" class="btn btn-primary btn-md btn-info">Here's what we used the last time we built this page</button>

<div id="sessioninfo" class="collapse">

```{r show-session-info, echo=TRUE, width=90}
writeLines(readLines(file.path(Sys.getenv("HOME"), ".R/Makevars")))

devtools::session_info()
```

</div>